### 2017-02-06 Paper/Paper_v1.tex
Arbeit und Bib-Datei

\documentclass[a4paper]{scrreprt}
\usepackage[backend=bibtex,style=numeric]{biblatex}
\usepackage{ngerman}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[style=numeric]{biblatex}
\bibliography{../PaperBib}
\usepackage{graphicx}
\usepackage{epstopdf}

\begin{document}

\title{Improving the modelling of tempo and phoneme duration}
\author{Alexandra Krah}
\date{06. April ~2017}
\maketitle
\chapter*{}
\tableofcontents

\chapter{Introduction}
The attempts to improve the quality of speech synthesis when tempo changes include linguistic processing, and phoneme length manipulation before the actual synthesis takes place. As mentioned in a patent documentation of Fujitsu Limited, phoneme length manipulation plays an important role in this context and one cannot rely on a linear adjustment function \cite{nishiike2008}. Phoneme length adjustment is the aspect we wish to adress in this work. In particular, we want to find an algorithm to modify the phoneme length which takes into consideration the local speech rate. The result should be an approximation of the phoneme length change which is close to the actual change that occurs when the speaking rate alters. Consequently, the modified speech unit would sound more natural at a faster or a slower speech rate than being modified with a linear function.

Hoequist and Kohler observed already back in 1986 that the change in speaking tempo does not produce a linear change of the acoustic segments of an utterance (i. e. phonemes) \cite{Hoequist1986}.

Syllable and foot structure of the utternace belong to the important factors influencing the segment length modification  \cite{Hoequist1986}.

More important as the segmentation of the utterance in words is its segmentation in prosodical segments. In spontaneous speech, some word parts may fall, others may be uttered together with other words so that the resulting phoneme duration tends to relate more to the structure of the resulting prosodic segment as to the original word \cite{Kohler1986}. This is also an alternative for calculating speech rate.
\section{Goals}
\section{Outline}

\chapter{Fundamentals}
\section{Data Mining}
\section{Machine Learning}
\section{Error Evaluation}
\section{Preparing the Data}

\chapter{Corpora}
\section{Verbmobil}
The Verbmobil corpus is a database of spontaneous speech containing a collection of appointment making dialogs, fully transliterated and annotated.

\chapter{Phoneme durations and the dynamics of speech}
We omitted pauses at the beginning and at the end of the turns.

\section{Glottal stop}
\section{Vowels}
\section{Consonants}

\chapter{Defining speech rate}

Local speech rate (check window size) vs. global speech rate vs. relative speech rate.

Pfitzinger \cite{Pfitzinger1998} showed that syllable count per second is a better approximation of the real speech rate than calculating the number of phones per second. He also proposed a formula combining the two rates, which should further improve this result for the calculation of the local speech rate, which was confirmed in later studies \cite{Pfitzinger1999}.

We decided to calculate global speech rate as a ratio of number of syllables per second. As syllables were not annotated in our database, we decided to use the number of (realized?) vowels as an approximation for the number of syllables \cite{Yishan_Jiao_2015} \cite{Kohler1995}.

We did compare speech rate considering only realized vowels/syllables and considering both realized and not realized syllables, as Koreman \cite{Koreman_2006} also did and...
Calculate also local speech rate.

\chapter{Finding an algorithm for modelling phoneme durations based on speech rate}

\chapter{Conclusion and Outlook}
Analyze duration changes of vowels in interrupted words.

\printbibliography
\end{document}

### 2017-02-09 Paper/Paper_v1.tex
Kleine Änderungen an der PDF

@@
 
-We did compare speech rate considering only realized vowels/syllables and considering both realized and not realized syllables, as Koreman \cite{Koreman_2006} also did and...
-Calculate also local speech rate.
+Koreman showed \cite{Koreman_2006} that the perceived speech rate is also influenced by the listener's knowledge of the expected articulations for a particular utterance, therefore it makes a difference if all expected phones are also articulated or not.
+Consequently, we did compare speech rate considering only articulated vowels/syllables, as well as considering both realized and intended syllables,  and tested both methods in our model.
+
+Calculated also local speech rate.
+
+We obtained the best approximation when using speech rate calculated as ...
 

### 2017-02-13 py_files/arff_file_builder.py
py-files for error calculation and some minor changes to the others.

@@ -8,13 +8,13 @@ import time
 # Generates an almost ready for processing with weka text file. 
 
 # List of phonemes to search for
-#phoneme_list = ["Q"]
+phoneme_list = ["a"]
 # All diphthongs:
 #phoneme_list = ["OY", "aU", "aI"]
 # All long vowels:
 #phoneme_list = ["a:", "e:", "E:", "i:", "o:", "u:", "y:", "2:", "a~:"]
 # All vowels:
-phoneme_list = ["a", "a~", "e", "E", "I", "i", "O", "o", "U", "u", "Y", "y", "9", "2", "a:", "a~:", "e:", "E:", "i:", "o:", "u:", "y:", "2:", "OY", "aU", "aI"]
+#phoneme_list = ["a", "a~", "e", "E", "I", "i", "O", "o", "U", "u", "Y", "y", "9", "2", "a:", "a~:", "e:", "E:", "i:", "o:", "u:", "y:", "2:", "OY", "aU", "aI"]
 
 # Get the dictionary of phonemes (keys) and their attributes (values)
 phon_dict = phon_dict.phon_dict(phoneme_list)
@@ -24,15 +24,16 @@ phon_dict = phon_dict.phon_dict(phoneme_list)
 dict_list = []
 for el in phoneme_list:
 	dict_list += phon_dict[el]
-
+no_of_e = len(dict_list)/12
+print(str(no_of_e))
 # Get the labels for the attribute: word
 word_list = set(dict_list[3::12])
 
 
 #test_file = open("C:/Users/alexutza_a/Abschlussarbeit/sp_kurzes_a.txt", "w") -> gibt falsche Anzahl von speech rates
-arff_file = open("C:/Users/alexutza_a/Abschlussarbeit/wekadateien/vowels_no_schwa.txt", "w")
+arff_file = open("C:/Users/alexutza_a/Abschlussarbeit/wekadateien/short_aln.txt", "w")
 
-arff_file.write("% ARFF file for Verbmobil\n% Phoneme: vowels_no_schwa\n%\n@relation vowels_no_schwa\n\n% List of attributes:\n%\n")
+arff_file.write("% ARFF file for Verbmobil\n% Phoneme: short_aln\n%\n@relation short_aln\n\n% List of attributes:\n%\n")
 arff_file.write("% Duration is expressed in milliseconds\n")
 
 arff_file.write("@attribute phoneme { ") 

### 2017-02-13 py_files/calculate_errors.py

@@ -1,0 +1,36 @@
+import error
+import fetch_data
+import numpy as np
+
+durations_dict = fetch_data.calculate_pho_mean(fetch_data.read_trainig_files())
+st_value_means = fetch_data.standard_value_means()
+
+def get_predictions():
+	predictions = []
+	for key in durations_dict.keys():
+		for i in range(1, len(durations_dict[key])):
+			predictions.append(durations_dict[key][0])
+
+	#print(durations_dict)
+	#print(predictions)
+
+	return predictions
+
+def get_observed():
+	observed = []
+
+	for key in durations_dict.keys():
+		observed += durations_dict[key][1:]
+	#print(observed)
+	return observed
+
+def calc_rmse_for_trdata():
+	min_o = min(get_observed())
+	max_o = max(get_observed())
+	rmse = error.rmse(np.array(get_predictions()), np.array(get_observed()))
+	rmse /= (max_o - min_o)
+	mae = error.mae(np.array(get_predictions()), np.array(get_observed()))
+	
+	return rmse, mae
+
+print(calc_rmse_for_trdata())

### 2017-02-13 py_files/error.py

@@ -1,0 +1,17 @@
+import numpy as np
+
+predictions = [0.06, 0.06, 0.06, 0.06]
+actual_values = [0.04, 0.06, 0.02, 0.06]
+
+def rmse(prediction, actual):
+	return np.sqrt(((prediction - actual) ** 2).mean())
+
+rmse_val = rmse(np.array(predictions), np.array(actual_values))
+
+def mae(prediction, actual):
+	return np.absolute(prediction - actual).mean()
+
+mae_val = mae(np.array(predictions), np.array(actual_values))
+
+#print(str(rmse_val))
+#print(str(mae_val))

### 2017-02-13 py_files/fetch_data.py

@@ -1,0 +1,47 @@
+import os
+import glob
+import re
+import numpy as np
+from collections import defaultdict
+
+# Returns a dictionary of the official phoneme means for VM1+2
+def standard_value_means():
+	o_dict = defaultdict(float)
+	official_file = open("C:/Users/alexutza_a/Abschlussarbeit/DB_Verbmobil/Evaluation/Training/Basic_german_phone_list.txt")
+	for line in official_file:
+		if len(line.split()[0]) < 3:
+			o_dict[line.split()[0]] = int(round((float(line.split()[6])/0.0000625), 0))
+
+	#print(o_dict)
+	return o_dict
+
+# Returns a dictionary of phoneme occurences (keys) in the training data and their durations (values)
+def read_trainig_files():
+	t_dict = defaultdict(list)
+	os.chdir("C:/Users/alexutza_a/Abschlussarbeit/DB_Verbmobil/Evaluation/Training")
+	
+	#Iterate over the training files
+	for file in glob.glob("*.par"):
+		work_file = open(file)
+
+		for line in work_file:
+			if re.match("MAU", line):
+				t_dict[line.split()[4]].append(int(line.split()[2]))
+		work_file.close()
+	
+	# Remove breaks from the data
+	x = t_dict.pop("<p:>")
+
+	return t_dict
+
+# Calculates the mean of phoneme durations / phoneme and puts it at the beginning of the value lists
+# Returns the phoneme dict of the training data updated with the mean duration of each phoneme 
+# 	at the beginning of each value list
+def calculate_pho_mean(phon_dict):
+	for key in phon_dict.keys():
+		key_mean = np.mean(phon_dict[key])
+		phon_dict[key].insert(0, int(round((key_mean), 0)))
+	#print(phon_dict)
+	return phon_dict
+
+calculate_pho_mean(read_trainig_files())

### 2017-02-13 py_files/phon_dict.py

@@ -5,6 +5,7 @@ import re
 from collections import defaultdict
 import stress
 import speech_rate
+import math
 
 # Creates a dictionary of durations (of specific phoneme)
 dur_file = open("C:/Users/alexutza_a/Abschlussarbeit/word_durations.txt", "w")
@@ -12,7 +13,7 @@ speech_rate_list = []
 # List of all german .par files in verbmobil, where the recording was done acn (main scenario dialogue 
 	#recorded means neckband microphone)
 def verbmo_par_files():
-	pattern = 'g*acn*.par'  # Pattern to be used for filtering filenames
+	pattern = 'g*acn.*par'  # Pattern to be used for filtering filenames
 	file_list = []			# Empty list to be populated with filenames matching pattern
 	path_list = []
 	for path, subfolder, filenames in os.walk('C:/Users/alexutza_a/Abschlussarbeit/DB_Verbmobil/verbmobil_par'):
@@ -26,7 +27,6 @@ file_list, path_list = verbmo_par_files()
 # Attributes are: Filename, Position in Speech (Start, Middle, End), Containing word, Stress, Overlapping, Speech rate
 def phon_dict(phoneme_list):
 	laut_schluessel = defaultdict(list)
-
 	# Iterate over the given list of phonemes, and
 	for phoneme in phoneme_list:
 		#iterate over all .par files in g016a and open/close one file at a time
@@ -41,11 +41,10 @@ def phon_dict(phoneme_list):
 			overlapping_list = overlapped_word(datei)
 			sp_pho, sp_w, speech_rate_msyl, speech_rate_ksyl = speech_rate.speech_rate(datei)
 			zeile = 0
-			
 		
 			for line in datei:
 				zeile += 1
-				if re.match("MAU", line) and (line.split()[4] == phoneme):
+				if re.match("MAU", line) and (str(line.split()[4]) == phoneme):
 					wort = line.split()[3]
 					# Attribute: phoneme, just for easier handling
 					laut_schluessel[phoneme].append(phoneme)
@@ -75,10 +74,10 @@ def phon_dict(phoneme_list):
 					laut_schluessel[phoneme].append(speech_rate_ksyl)
 					# Attribute: word position: initial, not initial - only for Q
 					#laut_schluessel[phoneme].append(stress.find_qPosition(file, (zeile+9)))
-					# Attribute: duration (in ms)
[…]

### 2017-02-19 Paper/Paper_v1.tex
py-files for error calculation and some minor changes to the others.
Syllable counter in pyfiles and some new weka files as text-file

@@
 
+Long a shows a strong negative correlation with the local speech rate for secondary stressed and unstressed instances.
+
 \section{Glottal stop}
@@
 Analyze duration changes of vowels in interrupted words.
+Consider actual realized syllables in speech instead of syllables based on the canonical transcription.
+In spontaneous speech, the word and syllable boundaries change, and this change influences other phonetic aspects such as stress, syllable duration, phoneme duration.
 

### 2017-02-19 py_files/arff_file_builder.py

@@ -8,13 +8,13 @@ import time
 # Generates an almost ready for processing with weka text file. 
 
 # List of phonemes to search for
-phoneme_list = ["a"]
+#phoneme_list = ["a:"]
 # All diphthongs:
 #phoneme_list = ["OY", "aU", "aI"]
 # All long vowels:
 #phoneme_list = ["a:", "e:", "E:", "i:", "o:", "u:", "y:", "2:", "a~:"]
 # All vowels:
-#phoneme_list = ["a", "a~", "e", "E", "I", "i", "O", "o", "U", "u", "Y", "y", "9", "2", "a:", "a~:", "e:", "E:", "i:", "o:", "u:", "y:", "2:", "OY", "aU", "aI"]
+phoneme_list = ["a", "a~", "e", "E", "I", "i", "O", "o", "U", "u", "Y", "y", "9", "2", "a:", "a~:", "e:", "E:", "i:", "o:", "u:", "y:", "2:", "OY", "aU", "aI"]
 
 # Get the dictionary of phonemes (keys) and their attributes (values)
 phon_dict = phon_dict.phon_dict(phoneme_list)
@@ -24,17 +24,16 @@ phon_dict = phon_dict.phon_dict(phoneme_list)
 dict_list = []
 for el in phoneme_list:
 	dict_list += phon_dict[el]
-no_of_e = len(dict_list)/12
-print(str(no_of_e))
+
 # Get the labels for the attribute: word
-word_list = set(dict_list[3::12])
+word_list = set(dict_list[3::13])
 
 
 #test_file = open("C:/Users/alexutza_a/Abschlussarbeit/sp_kurzes_a.txt", "w") -> gibt falsche Anzahl von speech rates
-arff_file = open("C:/Users/alexutza_a/Abschlussarbeit/wekadateien/short_aln.txt", "w")
+arff_file = open("C:/Users/alexutza_a/Abschlussarbeit/wekadateien/all_vowels_noSchwa.txt", "w")
 
-arff_file.write("% ARFF file for Verbmobil\n% Phoneme: short_aln\n%\n@relation short_aln\n\n% List of attributes:\n%\n")
-arff_file.write("% Duration is expressed in milliseconds\n")
+arff_file.write("% ARFF file for Verbmobil\n% Phoneme: all_vowels_noSchwa\n%\n@relation all_vowels_noSchwa\n\n% List of attributes:\n%\n")
+arff_file.write("% Duration is expressed in seconds\n")
 
[…]

### 2017-02-19 py_files/calculate_errors.py

@@ -1,36 +1,19 @@
 import error
-import fetch_data
+import prepare_data
 import numpy as np
 
-durations_dict = fetch_data.calculate_pho_mean(fetch_data.read_trainig_files())
-st_value_means = fetch_data.standard_value_means()
+phon_dict, test_dict, tm_dict = prepare_data.calculate_pho_mean(prepare_data.read_trainig_files())
+#print(test_dict)
 
-def get_predictions():
-	predictions = []
-	for key in durations_dict.keys():
-		for i in range(1, len(durations_dict[key])):
-			predictions.append(durations_dict[key][0])
+testlist = prepare_data.read_testfiles()
+trainingdict = tm_dict
 
-	#print(durations_dict)
-	#print(predictions)
+predictions = prepare_data.create_prediction_list(testlist, trainingdict)
+actual = prepare_data.read_testfiles()[1::2]
 
-	return predictions
+#print(predictions)
+#print(prepare_data.read_testfiles())
+#print(actual)
 
-def get_observed():
-	observed = []
-
-	for key in durations_dict.keys():
-		observed += durations_dict[key][1:]
-	#print(observed)
-	return observed
-
-def calc_rmse_for_trdata():
-	min_o = min(get_observed())
-	max_o = max(get_observed())
-	rmse = error.rmse(np.array(get_predictions()), np.array(get_observed()))
-	rmse /= (max_o - min_o)
-	mae = error.mae(np.array(get_predictions()), np.array(get_observed()))
-	
-	return rmse, mae
-
-print(calc_rmse_for_trdata())
+rmse, mae = error.calc_error_for_data(predictions, actual)
+print(rmse, mae)

### 2017-02-19 py_files/error.py

@@ -1,17 +1,51 @@
 import numpy as np
 
-predictions = [0.06, 0.06, 0.06, 0.06]
-actual_values = [0.04, 0.06, 0.02, 0.06]
+#predictions = [0.06, 0.06, 0.06, 0.06]
+#actual_values = [0.04, 0.06, 0.02, 0.06]
 
-def rmse(prediction, actual):
+# Returns the list of predicted values, for the training set
+def get_predictions(predictions_dict):
+	predictions = []
+	for key in predictions_dict.keys():
+		for i in range(1, len(predictions_dict[key])):
+			predictions.append(predictions_dict[key][0])
+	#	if key in vowel_list:
+	#		for i in range(1, len(durations_dict[key])):
+	#			predictions.append(1149)
+	#	else:
+	#		for i in range(1, len(durations_dict[key])):
+	#			predictions.append(durations_dict[key][0])
+	#print(durations_dict)
+	#print(predictions)
+
+	return predictions
+
+# Returns the list of actual values, for the training set
+def get_observed(observed_dict):
+	observed = []
+
+	for key in observed_dict.keys():
+		observed += observed_dict[key]
+	#print(observed)
+	return observed
+# Need
+def RMSE(prediction, actual):
 	return np.sqrt(((prediction - actual) ** 2).mean())
 
-rmse_val = rmse(np.array(predictions), np.array(actual_values))
 
-def mae(prediction, actual):
+# Need
+def MAE(prediction, actual):
 	return np.absolute(prediction - actual).mean()
 
-mae_val = mae(np.array(predictions), np.array(actual_values))
+#rmse_val = rmse(np.array(predictions), np.array(actual_values))
+#mae_val = mae(np.array(predictions), np.array(actual_values))
 
-#print(str(rmse_val))
-#print(str(mae_val))
+# Need
+# Compares the two lists via rmse and mae.
+def calc_error_for_data(predictions, actuals):
+	rmse = RMSE(np.array(predictions), np.array(actuals))
+	rmse *= 0.0000625
+	mae = MAE(np.array(predictions), np.array(actuals))
+	mae *= 0.0000625
+
+	return rmse, mae

### 2017-02-19 py_files/find_syl_position.py

@@ -1,0 +1,207 @@
+import re
+from collections import defaultdict
+
+diph = ["aI", "aU", "OY"]
+vowels = ["a", "A", "a~", "e", "E", "I", "i", "O", "o", "U", "u", "Y", "y", "9", "2", "@" , "6", "a:", "a~:", "e:", "E:", "i:", "o:", "u:", "y:", "2:"]
+consonants = ["z", "S", "Z", "C", "x", "N", "b", "d", "f", "g", "h", "j", "k", "l", "m", "n", "p", "r", "s", "t", "v"]
+
+# Returns the position of the syllable containing phoneme inside word:
+# - "w_initial" = 1st syllable, "w_final" = last syllable, "w_middle" = between 1st and last syll
+# - "one_s" = if it's a one syllable word
+# The counting of syllable is done on actually realized syllables! (not the canonical transcription)
+def syl_position(datei, word_no, cursor):
+	#print("Cursor: " + str(cursor))
+	syls_word = get_sylsplit_word(datei, word_no)
+	w1, w2 = mau_split(datei, word_no, cursor)
+	mau1, mau2, sylsplit_word = prepare_strings(w1, w2, syls_word)
+	#print("MAU1: " + mau1)
+	if sylsplit_word != "":
+		syl_dict = compare_strings(mau1+mau2, sylsplit_word)
+
+		#print(syl_dict.keys())
+		first_syl = min(syl_dict.keys())
+		last_syl = max(syl_dict.keys())
+		#print("Last syl: " + str(last_syl))
+
+		if first_syl != 0:
+			x = 0
+			for i in range (first_syl, last_syl+1):
+				syl_dict[x] = syl_dict.pop(i)
+				x += 1
+			last_syl = max(syl_dict.keys())
+		#print("New last syl: " + str(last_syl))
+		#print(syl_dict)
+		syl_no = get_syl(mau1.index(mau1[-1]), syl_dict)-1
+		#print("Syl no: " + str(syl_no))
+	else:
+		first_syl, last_syl = 0, 0
+
+	inword_pos = ""
+	if first_syl != last_syl:
+		if syl_no == 0:
+			inword_pos = "w_initial"
+		elif syl_no >= last_syl:
+			inword_pos = "w_final"
+		else:
+			inword_pos = "w_middle"
+	else:
+		inword_pos = "one_s"
+
+	return inword_pos
+
+
[…]

### 2017-02-19 py_files/get_cursor_pos.py

@@ -1,0 +1,17 @@
+import re
+
+def cursor_pos():
+	work_file = open("C:/Users/alexutza_a/Abschlussarbeit/DB_Verbmobil/Evaluation/Test/g002acn1_000_AAJ.par")
+	#work_file = open("C:/LP/ad_lesbiam.txt")
+	zeile = 0
+	p_zeile = []
+
+	for line in work_file:
+		zeile += 1
+		if re.match("MAU", line) and (int(line.split()[3]) == 9) and (str(line.split()[4]) == "n"):
+			p_zeile.append(zeile)
+			print(line)
+	work_file.close()
+	return p_zeile
+#cursor_pos()
+print(cursor_pos())

### 2017-02-19 py_files/phon_counter.py

@@ -1,0 +1,18 @@
+import phon_dict
+import re
+
+file_list, path_list = phon_dict.verbmo_par_files()
+
+def phon_counter():
+	count = 0
+	for file in path_list:
+		datei = open(file)
+		for line in datei:
+				if re.match("MAU", line) and (line.split()[4] == "a"):
+					count += 1
+				elif re.match("PRO", line):
+					break
+	datei.close()
+	return count
+
+print(str(phon_counter()))

### 2017-02-19 py_files/phon_dict.py

@@ -6,6 +6,7 @@ from collections import defaultdict
 import stress
 import speech_rate
 import math
+import find_syl_position
 
 # Creates a dictionary of durations (of specific phoneme)
 dur_file = open("C:/Users/alexutza_a/Abschlussarbeit/word_durations.txt", "w")
@@ -13,7 +14,7 @@ speech_rate_list = []
 # List of all german .par files in verbmobil, where the recording was done acn (main scenario dialogue 
 	#recorded means neckband microphone)
 def verbmo_par_files():
-	pattern = 'g*acn.*par'  # Pattern to be used for filtering filenames
+	pattern = 'g*acn*.par'  # Pattern to be used for filtering filenames
 	file_list = []			# Empty list to be populated with filenames matching pattern
 	path_list = []
 	for path, subfolder, filenames in os.walk('C:/Users/alexutza_a/Abschlussarbeit/DB_Verbmobil/verbmobil_par'):
@@ -27,6 +28,7 @@ file_list, path_list = verbmo_par_files()
 # Attributes are: Filename, Position in Speech (Start, Middle, End), Containing word, Stress, Overlapping, Speech rate
 def phon_dict(phoneme_list):
 	laut_schluessel = defaultdict(list)
+
 	# Iterate over the given list of phonemes, and
 	for phoneme in phoneme_list:
 		#iterate over all .par files in g016a and open/close one file at a time
@@ -37,15 +39,17 @@ def phon_dict(phoneme_list):
 		for file in path_list:
 			datei = open(file)
 			filename = str(file)
+			#print(filename)
 			words = total_words(datei)
 			overlapping_list = overlapped_word(datei)
 			sp_pho, sp_w, speech_rate_msyl, speech_rate_ksyl = speech_rate.speech_rate(datei)
 			zeile = 0
+			
 		
 			for line in datei:
 				zeile += 1
-				if re.match("MAU", line) and (str(line.split()[4]) == phoneme):
-					wort = line.split()[3]
+				if re.match("MAU", line) and (line.split()[4] == phoneme):
+					wort = int(line.split()[3])
 					# Attribute: phoneme, just for easier handling
 					laut_schluessel[phoneme].append(phoneme)
 					# Attribute: filename 
@@ -74,9 +78,11 @@ def phon_dict(phoneme_list):
 					laut_schluessel[phoneme].append(speech_rate_ksyl)
 					# Attribute: word position: initial, not initial - only for Q
[…]

### 2017-02-19 py_files/prepare_data.py

@@ -1,0 +1,92 @@
+import os
+import glob
+import re
+import numpy as np
+from collections import defaultdict
+
+# Returns a dictionary of the official phoneme means for VM1+2
+def standard_value_means():
+	o_dict = defaultdict(float)
+	omedian_dict = defaultdict(float)
+	official_file = open("C:/Users/alexutza_a/Abschlussarbeit/DB_Verbmobil/Evaluation/Training/Basic_german_phone_list.txt")
+	for line in official_file:
+		if len(line.split()[0]) < 3:
+			o_dict[line.split()[0]] = int(round((float(line.split()[6])/0.0000625), 0))
+			omedian_dict[line.split()[0]] = int(round((float(line.split()[9])/0.0000625), 0))
+	#print(o_dict)
+	return o_dict, omedian_dict
+o_dict, omedian_dict = standard_value_means()
+
+# Returns a dictionary of phoneme occurences (keys) in the training data and their durations (values)
+def read_trainig_files():
+	training_dict = defaultdict(list)
+	os.chdir("C:/Users/alexutza_a/Abschlussarbeit/DB_Verbmobil/Evaluation/Training")
+	
+	#Iterate over the training files
+	for file in glob.glob("*.par"):
+		work_file = open(file)
+
+		for line in work_file:
+			if re.match("MAU", line):
+				training_dict[line.split()[4]].append(int(line.split()[2]))
+		work_file.close()
+	
+	# Remove breaks from the data
+	x = training_dict.pop("<p:>")
+
+	return training_dict
+
+# Returns a list of phoneme occuring in the test files, followed by their respective durations
+# Looks like: ["a", 583, "b", 12, "a", 489, ...]
+def read_testfiles():
+	compare_list = []
+	os.chdir("C:/Users/alexutza_a/Abschlussarbeit/DB_Verbmobil/Evaluation/Test")
+
+	#Iterate over the test files
+	for file in glob.glob("*.par"):
+		work_file = open(file)
+
+		for line in work_file:
+			if re.match("MAU", line):
+				compare_list.append(str(line.split()[4]))
+				compare_list.append(int(line.split()[2]))
+		work_file.close()
+	
+	# Remove breaks from the data
+	for el in compare_list:
+		if el == "<p:>":
+			x = compare_list.index(el)
+			compare_list.pop(x+1)
+			compare_list.pop(x)
+	return compare_list
+
+# Calculates the mean of phoneme durations / phoneme and puts it at the beginning of the value lists
+# Returns:
[…]

### 2017-02-19 py_files/speech_rate.py

@@ -75,7 +75,7 @@ def speech_rate(datei):
 	return speech_rate_pho, speech_rate_w, speech_rate_msyl, speech_rate_ksyl
 
 # Returns:
-# - word duration in miliseconds,
+# - word duration in seconds,
 # - phoneme count per word
 # - syllable count per word
 def word_duration(datei, word_no):

### 2017-02-20 py_files/calculate_errors.py
Added rmse per phoneme and corrected some errors

@@ -1,12 +1,13 @@
 import error
 import prepare_data
 import numpy as np
+from collections import defaultdict
 
 phon_dict, test_dict, tm_dict = prepare_data.calculate_pho_mean(prepare_data.read_trainig_files())
 #print(test_dict)
 
 testlist = prepare_data.read_testfiles()
-trainingdict = tm_dict
+trainingdict = test_dict
 
 predictions = prepare_data.create_prediction_list(testlist, trainingdict)
 actual = prepare_data.read_testfiles()[1::2]
@@ -15,5 +16,50 @@ actual = prepare_data.read_testfiles()[1::2]
 #print(prepare_data.read_testfiles())
 #print(actual)
 
-rmse, mae = error.calc_error_for_data(predictions, actual)
-print(rmse, mae)
+global_rmse, mae = error.calc_error_for_data(predictions, actual)
+print(global_rmse, mae)
+
+
+### Following is rmse calculation individually (per phoneme) ###
+
+
+
+# Set of occuring phonemes to iterate on when creating dictionary
+set_phon = set(prepare_data.read_testfiles()[::2])
+# A list of phonemes in occuring order in the actual and pred lists
+a_phone_list = prepare_data.read_testfiles()[::2]
+# Use list comprehension to inweave the phoneme list with the list of predicted durations
+#   so we can create the dictionary of predictions per phoneme
+pred_list = [x for y in zip (a_phone_list, predictions) for x in y]
+
+# Dictionary of actual values per phoneme
+a_dict = defaultdict(list)
+# Dictionary of predicted values per phoneme
+p_dict = defaultdict(list)
+
+# Populate dictionary of actual values
+for el in set_phon:
+	for y in testlist:
+		if el == y:
+			a_dict[el].append(testlist[testlist.index(y)+1])
+
+# Populate dictionary of predicted values
+for el in set_phon:
+	for z in pred_list:
+		if el == z:
+			p_dict[el].append(pred_list[pred_list.index(z)+1])
+
+# Create lists for grouping rmse results
+b_list = []
+w_list = []
+# Populate b_list with all phonemes having a better score than global rmse
+# 	and w_list with all phonemes scoring worse than global rmse.
+for el in set_phon:
+	rmse = error.calc_error_for_data(p_dict[el], a_dict[el])
+	if rmse[0] <= global_rmse:
+		b_list.append(el)
+	else:
+		print(el + ": " + str(rmse))
+		w_list.append(el)
[…]

### 2017-02-20 py_files/prepare_data.py

@@ -53,12 +53,24 @@ def read_testfiles():
 		work_file.close()
 	
 	# Remove breaks from the data
+
+	# Get list of indexes for occurences of <p:>
+	pause_index = [i for i, val in enumerate(compare_list) if val == "<p:>"]
+	#print(pause_index)
+	pause_dur = [i + 1 for i in pause_index]
+	p_list = [x for y in zip (pause_index, pause_dur) for x in y]
+	#print(p_list)
+	actual_list = []
+
+	#print(compare_list)
+	ind = 0
+	# Copy list to new list, without pauses
 	for el in compare_list:
-		if el == "<p:>":
-			x = compare_list.index(el)
-			compare_list.pop(x+1)
-			compare_list.pop(x)
-	return compare_list
+		if ind not in p_list:
+			actual_list.append(el)
+		ind += 1
+
+	return actual_list
 
 # Calculates the mean of phoneme durations / phoneme and puts it at the beginning of the value lists
 # Returns:
@@ -90,3 +102,4 @@ def create_prediction_list(testfile_list, training_dict):
 	return prediction_list
 
 #print(create_prediction_list(read_testfiles(), test_dict))
+#print(read_testfiles())

### 2017-02-24 Paper/Paper_v1.tex
Some literature facts on duration models (a).

@@
 \section{Machine Learning}
-\section{Error Evaluation}
+\section{Performance Evaluation}
 \section{Preparing the Data}
@@
 
-\chapter{Finding an algorithm for modelling phoneme durations based on speech rate}
+\chapter{Phoneme duration prediction models}
+There are several standard duration prediction methods in use: 
+\begin{itemize}
+	\item Klatt is a rather simple method developped originally for American English.
+It multiplies a so-called inherent duration of given segment with a context dependent factor value and add this to a segment specific minimal duration.
+The challenges of using this method reside in finding an appropriate inherent duration, adapting the Klatt factor values to the German language, and chosing the minimal duration to use.
+	\item CART \cite{Riley1992} is a binary classification and regression tree with questions about the influencing factors at nodes and predicted values at the leaves.
+It allows consideration of both categorical and continuous factors, is quite simple and easy to interpret, in contrast to e.g.
+neural networks, and has a good time complexity of N*logN.
+The challenge of selecting an appropriate feature set is solved statistically.
+The sole drawback is that it needs a huge amount of training data in order to perform well.
+ge is that it needs a large amount of training data.
+	\item neural networks
+\end{itemize}
+
+Brinckmann and Trouvain \cite{Brinckmann_2003} compared the Klatt and CART methods for predicting segment duration in synthetic speech using the PhonDat database.
+Their results show a significantly better performance of CART over Klatt.
+However, they also report a strong influence of  the quality of the input data on the model performance.
 

### 2017-03-04 Paper/Paper_v1.tex
Added some literature based comments, especially on duration models.

@@
 \usepackage{epstopdf}
+\usepackage{textcomp}
 
@@
 
-More important as the segmentation of the utterance in words is its segmentation in prosodical segments.
+More important as the segmentation of the utterance in words is its segmentation in prosodical segments, so-called feet.
 In spontaneous speech, some word parts may fall, others may be uttered together with other words so that the resulting phoneme duration tends to relate more to the structure of the resulting prosodic segment as to the original word \cite{Kohler1986}.
-This is also an alternative for calculating speech rate.
 \section{Goals}
@@
 \section{Performance Evaluation}
+We calculated the performance of our model in terms of root mean square error (RMSE), mean absolute error (MAE) and correlation coefficient.
 \section{Preparing the Data}
@@
 
+Following natural rules, the tipical characteristic of data in this corpus is that speech rate often varies significantly inside a speech turn.
+This results in the challenge of finding an appropriate description for the speech rate across the corpus to be used for modeling the phoneme durations.
+We handle the issue of the speech rate separately in chapter 5.
+
 \chapter{Phoneme durations and the dynamics of speech}
-We omitted pauses at the beginning and at the end of the turns.
+Pauses may be caused and/or quantitatively modified both by linguistic (e.g.
+phonological) and non-linguistic factors, like technical, psychological, dialog dynamics, etc.
+Consequently pause length variation needs to be treated separately from phoneme duration variation, as a not negligeable large set of other factors are to be considered for this, which extends beyond the purpose of this work.
+Moreover, we omitted pauses at the beginning and at the end of the turns completely from our analysis, as we supposed the influence of non-linguistic factors which we cannot evaluate whithin the limits of this work to be greater in these positions.
+
+Consistent with the study of Rosen \cite{Rosen2005} on the application of the lognormal distribution for modeling the variation of speech segment duration, we opted for a log-transformation of the observed phoneme durations, which proved to be a better description of the data in our database as well.
+(Include grafics?!)
 
 Long a shows a strong negative correlation with the local speech rate for secondary stressed and unstressed instances.
+
+There have been presented in the literature many influencing factors for the duration of phonemes, which vary not only from one author to another but also from one language to another.
+For the German language, intuition aside, we have information on such factors provided by: Kohler \cite{Kohler1992}, Riedi \cite{Riedi1995}, ...
+
+We selected ... because...
 
@@
 \chapter{Defining speech rate}
+Under speech rate one always tends to understand a fraction of something divided by total time ocupied by that something.
+The German word ``Sprechgeschwindigkeit`` involves the concept of ``speed`` in the equation, which would be space divided by time.
+In the case of speech this would be number of units divided by total time ocupied by those units.
+However, this approach doesn't consider the contribution of the single phoneme durations to the whole segment, but rather supposes an equal length for all phonemes in the numerator.
 
-Local speech rate (check window size) vs. global speech rate vs. relative speech rate.
+Speech rate may be seen as a discrete or as a continuous variable. We opted for the continuous version, as it is better suited for our purpose.
+Speech rate may be calculated as global, local or relative speech rate.
 
@@
 Consequently, we did compare speech rate considering only articulated vowels/syllables, as well as considering both realized and intended syllables, and tested both methods in our model.
+Our tests showed that ...
 
-Calculated also local speech rate.
+In our study we avoided expressing speech rate neither as a fraction, nor as a number.
+We took the segment duration (i.e.
+denominator) alone, varied it, and tried to figure out what happends in the numerator so that the proportion stays.
+Expressed mathematically, if the durations of the constituent phonemes are $d_1$, $d_2$, and $d_3$, \textit{SR} is the resulting speech rate and \textit{time} represents the total duration of the 3 elements, this corresponds to the follwoing transformation:
 
-We obtained the best approximation when using speech rate calculated as ...
+$$\frac{d_1+d_2+d_3}{time} = SR \iff (d_1+d_2+d_3)  = SR \cdot time $$
 
 \chapter{Phoneme duration prediction models}
-There are several standard duration prediction methods in use: 
+There are several duration prediction methods in use: 
 \begin{itemize}
-	\item Klatt is a rather simple method developped originally for American English.
-It multiplies a so-called inherent duration of given segment with a context dependent factor value and add this to a segment specific minimal duration.
+	\item Klatt \cite{Klatt1979} is a rather simple method developped originally for American English.
+It multiplies a so-called inherent duration of given segment with a context dependent factor value and adds this to a segment specific minimal duration.
+Then the duration is estimated by adjusting the change coeficient successively for different factors, 11 of them being suggested by Klatt.
+These account for 84\% of observed total variance in segmental durations.
 The challenges of using this method reside in finding an appropriate inherent duration, adapting the Klatt factor values to the German language, and chosing the minimal duration to use.
@@
 The challenge of selecting an appropriate feature set is solved statistically.
-The sole drawback is that it needs a huge amount of training data in order to perform well.
-ge is that it needs a large amount of training data.
-	\item neural networks
+One drawback is that it needs a huge amount of training data in order to perform well.
+Also its performance deteriorates significantly with noisy or sparse training data \cite{Moebius1996}.
+	\item Sums-of-products (SoP) proposed by van Santen \cite{Santen1994} generalizes the formula proposed by Klatt and uses a decision tree to group factors acting in the same direction, either lengthening or shortening the phonemes, so their effects add up (amplificatory interactions).
+In fact, the most important part and main challange of this model is to find an appropriate combination of factors to be used for specific phoneme groups.
+He didn't use speech rate as a factor, but observed that the performance of the model deteriorates when the speech rate changes.
+The advantage of this solution is that is copes well with noisy or missing data, and works well with much less training data than CART \cite{Moebius1996}.
+The application of this model for the German language by Möbius and van Santen using the PhontDat database resulted in an overall correlation coefficient of 0.896.
+	\item Neural networks.
+Riedi \cite{Riedi1995} implemented a feed-forward neural network with two hidden layers for the prediction of phoneme durations and obtained a slightly better correlation coefficient than CART of 0.89 vs.
+0.86.
 \end{itemize}
 
-Brinckmann and Trouvain \cite{Brinckmann_2003} compared the Klatt and CART methods for predicting segment duration in synthetic speech using the PhonDat database.
-Their results show a significantly better performance of CART over Klatt.
+Brinckmann and Trouvain \cite{Brinckmann_2003} compared Klatt and CART methods for predicting segment duration using the PhonDat database.
+Their results show a significantly better performance of CART over Klatt, with values of 0.86 vs.
+0.79 for the correlation coefficient, consistent with previous results.
 However, they also report a strong influence of  the quality of the input data on the model performance.
+
+All these models have an upper asymptote for their accuracy situated at less than 100\%.
+It has been suggested, that the influence of so-called ``macroscopic`` \cite{Cummins1999} or para-linguistic \cite{Santen1994} factors should be examined in order to move this asymptote further up.
+We focused on the speech rate.
+
+\section{Method}
+Whichever phoneme duration prediction model one takes, and whichever size of the segment one takes in the denominator, in terms of syllable, word, foot or phrase duration...
+Larger the segment size, larger the number of influencing factors for phoneme duration.
+We opted for a progressive approach, starting at the syllable level.
 
@@
 Analyze duration changes of vowels in interrupted words.
-Consider actual realized syllables in speech instead of syllables based on the canonical transcription.
+Consider actual realized syllables in speech instead of syllables based on the canonical transcription?.
 In spontaneous speech, the word and syllable boundaries change, and this change influences other phonetic aspects such as stress, syllable duration, phoneme duration.

### 2017-03-08 py_files/calculate_errors.py
Introduced the correlation coefficient

@@ -5,6 +5,7 @@ from collections import defaultdict
 import prep_data_LSR
 
 #phon_dict, test_dict, tm_dict = prepare_data.calculate_pho_mean(prepare_data.read_trainig_files())
+# A dictionary giving the values of Mean and SD in a list for each encountered phoneme.
 stat_dict = prep_data_LSR.phone_stats(prep_data_LSR.read_trainig_files())
 #print(test_dict)
 
@@ -16,12 +17,14 @@ testlist = prep_data_LSR.read_testfiles()
 predictions = prep_data_LSR.create_prediction_list(testlist, stat_dict)
 actual = prep_data_LSR.read_testfiles()[1::3]
 
+
 #print(len(predictions))
 #print(prepare_data.read_testfiles())
 #print(len(actual))
 
 global_rmse, mae = error.calc_error_for_data(predictions, actual)
-print(global_rmse, mae)
+corrCoef = error.corrCoef(predictions, actual)[0]
+print(global_rmse, mae, corrCoef)
 
 
 ### Following is rmse calculation individually (per phoneme) ###

### 2017-03-08 py_files/error.py

@@ -1,4 +1,5 @@
 import numpy as np
+from scipy.stats.stats import pearsonr
 
 #predictions = [0.06, 0.06, 0.06, 0.06]
 #actual_values = [0.04, 0.06, 0.02, 0.06]
@@ -36,6 +37,9 @@ def RMSE(prediction, actual):
 # Need
 def MAE(prediction, actual):
 	return np.absolute(prediction - actual).mean()
+
+def corrCoef(prediction, actual):
+	return pearsonr(prediction, actual)
 
 #rmse_val = rmse(np.array(predictions), np.array(actual_values))
 #mae_val = mae(np.array(predictions), np.array(actual_values))

### 2017-03-08 py_files/prep_data_LSR.py

@@ -18,7 +18,9 @@ def read_trainig_files():
 		for line in work_file:
 			if re.match("MAU", line):
 				training_dict[line.split()[4]].append(int(line.split()[2]))
-				training_dict[line.split()[4]].append(speech_rate.local_speech_rate(file, int(line.split()[3])))
+				word_duration, phon_count, syl_count = speech_rate.word_duration(file, int(line.split()[3]))
+				training_dict[line.split()[4]].append((word_duration/0.0000625)/phon_count)
+				#training_dict[line.split()[4]].append(speech_rate.local_speech_rate(file, int(line.split()[3])))
 
 		work_file.close()
 	
@@ -37,6 +39,7 @@ def phone_stats(training_dict):
 
 	return stat_dict
 
+# Not used
 def mean_of_means(stat_dict):
 	m1 = 0
 
@@ -59,7 +62,9 @@ def read_testfiles():
 			if re.match("MAU", line):
 				compare_list.append(str(line.split()[4]))
 				compare_list.append(int(line.split()[2]))
-				compare_list.append(speech_rate.local_speech_rate(file, int(line.split()[3])))
+				word_duration, phon_count, syl_count = speech_rate.word_duration(file, int(line.split()[3]))
+				compare_list.append((word_duration/0.0000625)/phon_count) # speech rate as word_duration / # phonemes
+				#compare_list.append(speech_rate.local_speech_rate(file, int(line.split()[3])))
 		work_file.close()
 	#print(compare_list)
 	# Remove breaks from the data
@@ -102,15 +107,16 @@ def official_stats():
 #  from the official mean statistics of Verbmobil, for phonemes, which don't occur in the training set.
 # @param testfile_list: the list returned by read_testfiles()
 #	Looks like: ["a", 583, 0.5, "b", 12, 0.78, "a", 489, 0.12, ...]
-# @param training_dict: the full dictionary built from the training data
-#	Looks like {a: [1452, 0.8, 799, 0.5], b : [655, 0.5, 799, 0.45]...}
+# @param stat_dict: dictionary giving the mean and the SD for each phoneme 
+#   NO: the full dictionary built from the training data
+#	NO: Looks like {a: [1452, 0.8, 799, 0.5], b : [655, 0.5, 799, 0.45]...}
[…]

### 2017-03-13 py_files/calc_errors.py
Some more code snippets for data analysis.

@@ -1,8 +1,7 @@
-import error
-import prepare_data
+import performance_measures
 import numpy as np
 from collections import defaultdict
-import prep_data_LSR
+import pd_LSR
 
 #phon_dict, test_dict, tm_dict = prepare_data.calculate_pho_mean(prepare_data.read_trainig_files())
 # A dictionary giving the values of Mean and SD in a list for each encountered phoneme.
@@ -22,8 +21,8 @@ actual = prep_data_LSR.read_testfiles()[1::3]
 #print(prepare_data.read_testfiles())
 #print(len(actual))
 
-global_rmse, mae = error.calc_error_for_data(predictions, actual)
-corrCoef = error.corrCoef(predictions, actual)[0]
+global_rmse, mae = performance_measures.calc_error_for_data(predictions, actual)
+corrCoef = performance_measures.corrCoef(predictions, actual)[0]
 print(global_rmse, mae, corrCoef)
 
 
@@ -74,7 +73,7 @@ w_list = []
 # Populate b_list with all phonemes having a better score than global rmse
 # 	and w_list with all phonemes scoring worse than global rmse.
 for el in set_phon:
-	rmse = error.calc_error_for_data(p_dict[el], a_dict[el])
+	rmse = performance_measures.calc_error_for_data(p_dict[el], a_dict[el])
 	if rmse[0] <= global_rmse:
 		b_list.append(el)
 	else:

### 2017-03-13 py_files/find_syl_position.py

@@ -31,7 +31,12 @@ def syl_position(datei, word_no, cursor):
 			last_syl = max(syl_dict.keys())
 		#print("New last syl: " + str(last_syl))
 		#print(syl_dict)
-		syl_no = get_syl(mau1.index(mau1[-1]), syl_dict)-1
+		try:
+			syl_no = get_syl(mau1.index(mau1[-1]), syl_dict)-1
+		except:
+			syl_no = 0
+		else:
+			syl_no = 0
 		#print("Syl no: " + str(syl_no))
 	else:
 		first_syl, last_syl = 0, 0
@@ -147,18 +152,19 @@ def compare_strings(mau12, sylsplit_word):
 
 	for el in mau:
 		#print(el)
-		if el == sylsplit_word[0]:
-			sylsplit_word = sylsplit_word[1:]
-		elif el in sylsplit_word:
-			if len(mau) <= len(sylsplit_word[sylsplit_word.index(el):]):
-				if "." in sylsplit_word[:sylsplit_word.index(el)]:
-					syl_count +=1
-				sylsplit_word = sylsplit_word[sylsplit_word.index(el)+1:]
-		elif (el not in sylsplit_word) and (sylsplit_word[0] == "."):
-			syl_count += 1
-			sylsplit_word = sylsplit_word[1:]
-
-		
+		try:
+			if el == sylsplit_word[0]:
+				sylsplit_word = sylsplit_word[1:]
+			elif el in sylsplit_word:
+				if len(mau) <= len(sylsplit_word[sylsplit_word.index(el):]):
+					if "." in sylsplit_word[:sylsplit_word.index(el)]:
+						syl_count +=1
+					sylsplit_word = sylsplit_word[sylsplit_word.index(el)+1:]
+			elif (el not in sylsplit_word) and (sylsplit_word[0] == "."):
+				syl_count += 1
+				sylsplit_word = sylsplit_word[1:]
+		except:
+			print(mau, sylsplit_word)
 		mau_syl_d[syl_count].append(el)
 		mau = mau[1:]
 

### 2017-03-13 py_files/pd_LSR.py

@@ -63,7 +63,7 @@ def read_testfiles():
 				compare_list.append(str(line.split()[4]))
 				compare_list.append(int(line.split()[2]))
 				word_duration, phon_count, syl_count = speech_rate.word_duration(file, int(line.split()[3]))
-				compare_list.append((word_duration/0.0000625)/phon_count) # speech rate as word_duration / # phonemes
+				compare_list.append(round((word_duration/0.0000625)/phon_count, 1)) # speech rate as word_duration / # phonemes
 				#compare_list.append(speech_rate.local_speech_rate(file, int(line.split()[3])))
 		work_file.close()
 	#print(compare_list)
@@ -102,6 +102,26 @@ def official_stats():
 	#print(o_dict)
 	return o_dict#, omedian_dict
 
+# Quote of phoneme mean duration being greater than the local speech rate (as word duration / no. of phonemes)
+def test_mean(training_dict, stat_dict):
+	test_dict = defaultdict(float)
+
+	for elem in training_dict.keys():
+		rate_list = training_dict[elem][1::2]
+		test_mean_list = [ 1 for x in rate_list if x <= stat_dict[elem][0]]
+		test_dict[elem] = len(test_mean_list)/len(rate_list)
+	return test_dict
+
+# Create a list of durations and a list of word durations from the training data
+def dur_vs_rate(training_dict):
+    duration_list = []
+    rate_list = []
+    for el in training_dict.keys():
+        duration_list += training_dict[el][::2]
+        rate_list += training_dict[el][1::2]
+    return duration_list, rate_list
+#duration_list, rate_list = dur_vs_rate(read_trainig_files())
+
 # Returns a list with predicted durations for the phonemes of the test set.
 # Predicted durations come from the observed durations in the training set (mean and SD), and
 #  from the official mean statistics of Verbmobil, for phonemes, which don't occur in the training set.
@@ -112,6 +132,7 @@ def official_stats():
 #	NO: Looks like {a: [1452, 0.8, 799, 0.5], b : [655, 0.5, 799, 0.45]...}
 def create_prediction_list(testfile_list, stat_dict):
 	phone_list = testfile_list[::3]
[…]

### 2017-03-13 py_files/phon_dict.py

@@ -21,8 +21,8 @@ def verbmo_par_files():
 		for filename in fnmatch.filter(filenames, pattern):
 			file_list.append(filename)
 			path_list.append(os.path.join(path, filename))
-	return file_list, path_list
-file_list, path_list = verbmo_par_files()
+	return path_list#,file_list
+path_list = verbmo_par_files()
 
 # Dictionary of the given phoneme instance length in ms (keys) in Verbmobil and its attributes (values in list)
 # Attributes are: Filename, Position in Speech (Start, Middle, End), Containing word, Stress, Overlapping, Speech rate
@@ -75,7 +75,9 @@ def phon_dict(phoneme_list):
 					# Attribute: speech rate as effective syllables / sec
 					laut_schluessel[phoneme].append(speech_rate_msyl)
 					# Attribute: speech rate as total syllables / sec
-					laut_schluessel[phoneme].append(speech_rate_ksyl)
+					#laut_schluessel[phoneme].append(speech_rate_ksyl)
+					word_duration, phon_count, syl_count = speech_rate.word_duration(file, int(line.split()[3]))
+					laut_schluessel[phoneme].append(round((word_duration*0.00000625)/phon_count, 4))
 					# Attribute: word position: initial, not initial - only for Q
 					#laut_schluessel[phoneme].append(stress.find_qPosition(file, (zeile+9)))
 					# Attribute: Position in word (based on syllable position): one_s, w_1, w_2, w_3
@@ -187,8 +189,8 @@ def total_words(datei):
 	words = max(no_of_words)
 	# sets back cursor position after header
 	datei.seek(91)
-
 	return words
-
+d = phon_dict(["a:"])
+print(d["a:"][:65])
 #print(p_rank_in_word("C:/Users/alexutza_a/Abschlussarbeit/DB_Verbmobil/verbmobil_par/g016a/g016acn1_016_ABE.par", 11, "a", 410))
 #print(stress_type("C:/Users/alexutza_a/Abschlussarbeit/DB_Verbmobil/verbmobil_par/g016a/g016acn1_016_ABE.par", 11, "a", 407))

### 2017-03-13 py_files/speech_rate.py

@@ -86,19 +86,21 @@ def word_duration(datei, word_no):
 
 	for line in work_file:
 		if re.match("MAU", line) and (int(line.split()[3]) == int(word_no)):
+			#print(line)
 			word_duration += int(line.split()[2])
 			phoneme_count += 1
 			if str(line.split()[4]) in vowel_list:
 				mau_syl_count += 1
 	
-	word_duration *= 0.0000625
+	#word_duration *= 0.0000625
 	work_file.close()
+	#print(word_duration)
 	return word_duration, phoneme_count, mau_syl_count
 
 # Calculates local speech rate based on Pfitzingers formula (SR = s * syl_rate + p * phoneme_rate)
 def local_speech_rate(datei, word_no):
 	w_duration, phoneme_count, syllable_count = word_duration(datei, word_no)
-	w_duration = float(w_duration)
+	w_duration = float(w_duration*0.0000625)
 	phoneme_count = float(phoneme_count)
 	syllable_count = float(syllable_count)
 

### 2017-03-14 Paper/Paper_v1.tex
Added python notebook (jupyter)
Added some info to chapter Fundamentals.

@@
 \section{Data Mining}
+Data Mining means in foreground looking for patterns in data with the help of computers.
+Patterns are helpful for humans to understand data.
+They reduce the amount of data, and of available features to the relevant ones.
+Computers help process within resonable time amounts of data that would need years of manual processing.
 \section{Machine Learning}
+Machine Learning means building a program (model) which based on data mining results from a given dataset can make predictions about an unseen dataset  (fill in missing information). 
+\section{Preparing the Data}
+A very important and time consuming step in data mining is data preparation. 
 \section{Performance Evaluation}
-We calculated the performance of our model in terms of root mean square error (RMSE), mean absolute error (MAE) and correlation coefficient.
-\section{Preparing the Data}
+Considering machine learning, it is straightforward that one needs to split the available dataset at least into two parts: a training dataset, used for discovering patterns (data mining) in data, and a test dataset, used for testing the performance of the model (result of machine learning process).
+Model performance may be evaluated in several ways. We decided for the following three:
+\begin{enumerate}
+	\item \textbf{RMSE} or \textbf{root mean squared error} measures the differences between the numeric values predicted by a model and those actually observed: 
+$RMSE = \sqrt{\frac{\sum(predicted-actual)^2}{n}}$. It is a very popular method for evaluating errors of a model, but it is very sensible to outliers.
+	\item \textbf{MAE} or \textbf{mean absolute error} is similar to RMSE, but penalises outliers less.
+It is computed using the average of the absolute errors: $MAE = \frac{\sum\left|predicted-actual\right|}{n}$
+	\item The \textbf{Correlation coefficient} is a number between -1 and 1 that quantifies the correlation between two datasets, that is how much the value change in one dataset corresponds to the value change in the other dataset.
+Values close to 1 mean that greater values in one dataset imply greater values in the second dataset, negative values mean the values of the two sets vary in opposite directions, and values close to 0 mean there is no correlation between the two datasets.
+However, one must keep in mind that correlation does not imply causality.
+\end{enumerate}
 
@@
 The Verbmobil corpus is a database of spontaneous speech containing a collection of appointment making dialogs, fully transliterated and annotated.
+We used 286 recordings from the last verbmobil phase (Verbmobil II) created by 131 speakers.
+The total VMII corpus used amounts to 11783 conversational turns representing recorded speech time of 15,5 hours.
+The ``symbolic data`` corresponding to each turn (e.g.phonetic transcription, segmentation, labeling, etc.) is captured in so-called ``par files``, representing files in the ``BAS Partitur Format`` \cite{Burger}.
 
@@
 We handle the issue of the speech rate separately in chapter 5.
+
+Another challenge resulting from the structure of this corpus as collection of dialogs is the speaker variation.
+The exact values of the phoneme duration coordinates vary across speakers, as one can see in Fig.
+???
+for the case of /a/.
 

### 2017-03-14 Paper/Paper_v1.tex
Minor changes to the pdf

@@
 It is computed using the average of the absolute errors: $MAE = \frac{\sum\left|predicted-actual\right|}{n}$
-	\item The \textbf{Correlation coefficient} is a number between -1 and 1 that quantifies the correlation between two datasets, that is how much the value change in one dataset corresponds to the value change in the other dataset.
-Values close to 1 mean that greater values in one dataset imply greater values in the second dataset, negative values mean the values of the two sets vary in opposite directions, and values close to 0 mean there is no correlation between the two datasets.
-However, one must keep in mind that correlation does not imply causality.
+	\item The \textbf{correlation coefficient} is a number between -1 and 1 that quantifies the correlation between two variables, each having a separate value set.
+We can interpret this as how much the variable movement through its value set corresponds to the movement of the other variable in its own value set.
+Values close to 1 mean that greater values of the one variable imply greater values for the second variable, negative values mean the two variables move in opposite directions, and values close to 0 mean there is no correlation between the two variables.
+However, one must keep in mind that correlation does not imply causality, so the presence of correlation between two variables does not mean that one of them has any influence on the other one.
+If we compare red Ferrari cars to blue Renault cars, one may find a correlation between car color red and car speed.
+However, this does not imply that red cars are faster.
 \end{enumerate}

### 2017-03-14 Paper/Paper_v1.tex
Some additions to section on Verbmobil.

@@
 The Verbmobil corpus is a database of spontaneous speech containing a collection of appointment making dialogs, fully transliterated and annotated.
+This was an important point in selecting this database for our purpose, as phoneme duration is influenced by many factors, and the Verbmobil textual database provides most of them.
 We used 286 recordings from the last verbmobil phase (Verbmobil II) created by 131 speakers.
@@
 
-Following natural rules, the tipical characteristic of data in this corpus is that speech rate often varies significantly inside a speech turn.
-This results in the challenge of finding an appropriate description for the speech rate across the corpus to be used for modeling the phoneme durations.
-We handle the issue of the speech rate separately in chapter 5.
-
-Another challenge resulting from the structure of this corpus as collection of dialogs is the speaker variation.
-The exact values of the phoneme duration coordinates vary across speakers, as one can see in Fig.
+Drawbacks of this database come mainly from the characteristics of spontaneous speech.
+Following natural rules, speech rate often varies significantly inside a speech turn.
+One starts his turn at a normal speech rate, then gets to a point where he/she hesitates and gets slower, like when weighting different transport means to get to the appointment they are to agree upon.
+This results in the challenge of finding an appropriate description for the speech rate to be used for modeling phoneme durations across the corpus.
+We handle this issue of speech rate separately in chapter 5.
+Another challenge resulting from the structure of this corpus as a collection of dialogs is speaker variation.
+The exact coordinates of the phoneme durations vary across speakers, as one can see in Fig.
 ???

### 2017-03-15 Paper/Paper_v1.tex
Added some remarks to pdf chapter about phonemes.

@@
 \usepackage{graphicx}
-\usepackage{epstopdf}
+\usepackage[outdir=../Graphen/]{epstopdf}
 \usepackage{textcomp}
+\usepackage{floatrow}
+\renewcommand{\figurename}{Fig.}
 
@@
 \tableofcontents
+\listoffigures
 
@@
 In spontaneous speech, some word parts may fall, others may be uttered together with other words so that the resulting phoneme duration tends to relate more to the structure of the resulting prosodic segment as to the original word \cite{Kohler1986}.
+
+Accross the paper we will use the German SAMPA notations when referring to phonemes.
 \section{Goals}
@@
 The exact coordinates of the phoneme durations vary across speakers, as one can see in Fig.
-???
-for the case of /a/.
+\ref{fig:speaker_cmp} for the case of /a:/.
+
+\begin{figure}[htbp]
+	\includegraphics[width=0.6\textwidth]{../Graphen/Duration_of_a_5speakers.eps}
+	\centering
+	\caption[Phoneme length variation between speakers]{Overview of /a:/ length on the 5 most frequent speakers.
+Phoneme length is expressed as the natural logarithmus of the sample frequency.
+While the minimum duration is technically limited to an equal value corresponding to 0.04 sec, all other coordinates vary, e.g.
+maximum value has a variation of 0.14 sec between speakers AKX and ALR.
+The greatest median variation is 0.07 sec between speakers ALR (0.15 sec) and AJZ (0.08 sec).}
+	\label{fig:speaker_cmp}
+\end{figure}
 
 \chapter{Phoneme durations and the dynamics of speech}
+The phoneme inventory used in Verbmobil includes 52 classes defined according to the German SAMPA.
+Using the classical phonetic conventions, we can group these phonemes according to their phonetic features using a tree structure into vowels, consonants and a glottal stop on a first level, although glottal stop is usually treated as a consonant.
+Further relevant groupings are:
+\begin{itemize}
+	\item Vowels
+
+	\begin{itemize}
+		\item Diphthongs: when two vowels are articulated together in a syllable. They are generally characterized by longer durations. German knows three diphthongs: /aI/, /aU/ and /OY/.
+		\item Schwa: most frequent vowel in German as well as in English.
+It is also called a reduction vowel \cite{Kohler1995}, as most of the other monophthongs tend to be articulated as schwa in non-stressed vowels.
+Verbmobil differentiates two types of schwa for the German language: /@/ like in the last syllable of \texttt{lesen} and /6/ like in \texttt{Leser}, where the reduced vowel is followed by an /r/.
+		\item Monophthongs, which may be further split according to the short - long pairs or based on the articulation point like open - closed vowels or back - front vowels.
+We shall verify the relevance of these subcategories for our purpose in the section dedicated to vowels.
+	\end{itemize}
+
+	\item Consonants
+
+	\begin{itemize}
+		\item Plosives, which are generally short, is due to their articulation characteristics.
+		\item Fricatives, which show a legth variation almost comparable to that of vowels.
+		\item Nasals: /m/, /n/, /N/.
+		\item Laterals: /l/
+		\item Specials: /h/, /C/, /j/, /r/
+	\end{itemize}
+\end{itemize}
+
 Pauses may be caused and/or quantitatively modified both by linguistic (e.g.
@@
 
-Consistent with the study of Rosen \cite{Rosen2005} on the application of the lognormal distribution for modeling the variation of speech segment duration, we opted for a log-transformation of the observed phoneme durations, which proved to be a better description of the data in our database as well.
+Consistent with the study of Rosen \cite{Rosen2005} on the application of the lognormal distribution for modeling the variation of speech segment duration, we opted for a log-transformation of the observed phoneme durations using the natural logarithm, which proved to be a better description of the data in our database as well.
 (Include grafics?!)
@@
 
+\section{Vowels}
 \section{Glottal stop}
-\section{Vowels}
 \section{Consonants}

### 2017-03-15 py_files/arff_file_builder.py
added a picture

@@ -8,13 +8,13 @@ import time
 # Generates an almost ready for processing with weka text file. 
 
 # List of phonemes to search for
-#phoneme_list = ["a:"]
+phoneme_list = ["Q"]
 # All diphthongs:
 #phoneme_list = ["OY", "aU", "aI"]
 # All long vowels:
 #phoneme_list = ["a:", "e:", "E:", "i:", "o:", "u:", "y:", "2:", "a~:"]
 # All vowels:
-phoneme_list = ["a", "a~", "e", "E", "I", "i", "O", "o", "U", "u", "Y", "y", "9", "2", "a:", "a~:", "e:", "E:", "i:", "o:", "u:", "y:", "2:", "OY", "aU", "aI"]
+#phoneme_list = ["a", "a~", "e", "E", "I", "i", "O", "o", "U", "u", "Y", "y", "9", "2", "a:", "a~:", "e:", "E:", "i:", "o:", "u:", "y:", "2:", "OY", "aU", "aI"]
 
 # Get the dictionary of phonemes (keys) and their attributes (values)
 phon_dict = phon_dict.phon_dict(phoneme_list)
@@ -30,14 +30,14 @@ word_list = set(dict_list[3::13])
 
 
 #test_file = open("C:/Users/alexutza_a/Abschlussarbeit/sp_kurzes_a.txt", "w") -> gibt falsche Anzahl von speech rates
-arff_file = open("C:/Users/alexutza_a/Abschlussarbeit/wekadateien/all_vowels_noSchwa.txt", "w")
+arff_file = open("C:/Users/alexutza_a/Abschlussarbeit/wekadateien/Q.txt", "w")
 
-arff_file.write("% ARFF file for Verbmobil\n% Phoneme: all_vowels_noSchwa\n%\n@relation all_vowels_noSchwa\n\n% List of attributes:\n%\n")
+arff_file.write("% ARFF file for Verbmobil\n% Phoneme: all_vowels_noSchwa\n%\n@relation glottal_stop\n\n% List of attributes:\n%\n")
 arff_file.write("% Duration is expressed in seconds\n")
 
-arff_file.write("@attribute phoneme { ") 
-for el in phoneme_list:
-	arff_file.write(el + ", ")
+arff_file.write("@attribute phoneme { Q") 
+#for el in phoneme_list:
+#	arff_file.write(el + ", ")
 arff_file.write( " }\n")
 
 arff_file.write("@attribute filename string\n")

### 2017-03-15 py_files/phon_dict.py

@@ -61,7 +61,7 @@ def phon_dict(phoneme_list):
 					# Attribute: pause = if a pause preceeds or follows the word in which the phoneme occurs: near_pause, no_pause
 					laut_schluessel[phoneme].append(is_pause_near(file, wort))
 					# Attribute: stress = type of stress: none, s_stress, p_stress
-					laut_schluessel[phoneme].append(stress.stress_type(file, wort, phoneme, (zeile+9)))
+					laut_schluessel[phoneme].append(stress.stress_type(file, wort, phoneme, (zeile+10))) #modified zeile from +9 to +10, just for Q
 					# Attribute: overlapping
 					if wort in overlapping_list:
 						laut_schluessel[phoneme].append("1")
@@ -79,11 +79,12 @@ def phon_dict(phoneme_list):
 					word_duration, phon_count, syl_count = speech_rate.word_duration(file, int(line.split()[3]))
 					laut_schluessel[phoneme].append(round((word_duration*0.00000625)/phon_count, 4))
 					# Attribute: word position: initial, not initial - only for Q
-					#laut_schluessel[phoneme].append(stress.find_qPosition(file, (zeile+9)))
+					laut_schluessel[phoneme].append(stress.find_qPosition(file, (zeile+9)))
 					# Attribute: Position in word (based on syllable position): one_s, w_1, w_2, w_3
 					laut_schluessel[phoneme].append(find_syl_position.syl_position(file, wort, (zeile+9)))
 					# Attribute: duration (in sec)
-					laut_schluessel[phoneme].append(round(freq_to_ms(int(line.split()[2])/1000), 4))
+					laut_schluessel[phoneme].append(int(line.split()[2]))
+					#laut_schluessel[phoneme].append(round(freq_to_ms(int(line.split()[2])/1000), 4))
 			#datei.seek(91)
 			datei.close()
 	return laut_schluessel

### 2017-03-16 Paper/Paper_v1.tex
Added section on glottal stop

@@
 \renewcommand{\figurename}{Fig.}
+\usepackage{caption}
+\usepackage{subcaption}
 
@@
 Using the classical phonetic conventions, we can group these phonemes according to their phonetic features using a tree structure into vowels, consonants and a glottal stop on a first level, although glottal stop is usually treated as a consonant.
-Further relevant groupings are:
-\begin{itemize}
-	\item Vowels
-
-	\begin{itemize}
-		\item Diphthongs: when two vowels are articulated together in a syllable. They are generally characterized by longer durations. German knows three diphthongs: /aI/, /aU/ and /OY/.
-		\item Schwa: most frequent vowel in German as well as in English.
-It is also called a reduction vowel \cite{Kohler1995}, as most of the other monophthongs tend to be articulated as schwa in non-stressed vowels.
-Verbmobil differentiates two types of schwa for the German language: /@/ like in the last syllable of \texttt{lesen} and /6/ like in \texttt{Leser}, where the reduced vowel is followed by an /r/.
-		\item Monophthongs, which may be further split according to the short - long pairs or based on the articulation point like open - closed vowels or back - front vowels.
-We shall verify the relevance of these subcategories for our purpose in the section dedicated to vowels.
-	\end{itemize}
-
-	\item Consonants
-
-	\begin{itemize}
-		\item Plosives, which are generally short, is due to their articulation characteristics.
-		\item Fricatives, which show a legth variation almost comparable to that of vowels.
-		\item Nasals: /m/, /n/, /N/.
-		\item Laterals: /l/
-		\item Specials: /h/, /C/, /j/, /r/
-	\end{itemize}
-\end{itemize}
+Further relevant groupings are being examined in the sections ``Vowels`` and ``Consonants``.
 
@@
 Moreover, we omitted pauses at the beginning and at the end of the turns completely from our analysis, as we supposed the influence of non-linguistic factors which we cannot evaluate whithin the limits of this work to be greater in these positions.
+
+\begin{figure}[htbp]
+\centering
+	\begin{subfigure}[h]{0.45\textwidth}
+	\centering
+	\includegraphics[width=\textwidth]{../Graphen/Hist_aims_no_title.eps}
+	\caption{Duration of /aI/ measured in seconds.}
+	\label{fig:ai_ms}
+	\end{subfigure}
+\hfill
+	\begin{subfigure}[h]{0.45\textwidth}
+	\centering
+	\includegraphics[width=\textwidth]{../Graphen/Hist_ailn_no_title.eps}
+	\caption{Duration of /aI/ measured in ln(sec).}
+	\label{fig:ai_ln}
+	\end{subfigure}
+\caption[Phoneme duration distribution ms vs ln(ms)]{Phoneme durations measured in seconds have a not normal distribution shape.
+However, when applying a logarithmic transformation in terms of ln(msec), the distribution is better shaped.}
+\label{fig:aI_distrib}
+\end{figure}
 
@@
 \section{Vowels}
+\begin{itemize}
+	\item Diphthongs: when two vowels are articulated together in a syllable. They are generally characterized by longer durations. German knows three diphthongs: /aI/, /aU/ and /OY/.
+	\item Schwa: most frequent vowel in German as well as in English.
+It is also called a reduction vowel \cite{Kohler1995}, as most of the other monophthongs tend to be articulated as schwa in non-stressed vowels.
+Verbmobil differentiates two types of schwa for the German language: /@/ like in the last syllable of \texttt{lesen} and /6/ like in \texttt{Leser}, where the reduced vowel is followed by an /r/.
+	\item Monophthongs, which may be further split according to the short - long pairs or based on the articulation point like open - closed vowels or back - front vowels.
+We shall verify the relevance of these subcategories for our purpose in the section dedicated to vowels.
+\end{itemize}
+
 \section{Glottal stop}
+The glottal stop occurs only word initially in our database, and was marked with \texttt{/Q/}.
+We decided to treat it separately because of its particular usage in German: it may occur only at the beginning of a word or a word stem and in front of a vowel.
+Some phonetic manuals don't event consider it a phoneme \cite{Ternes2012}.
+Furthermore, it's acoustical properties make it look like a pause in speech, as you may see in Fig.
+..., which represents a challenge for the transcription.
+As our files were segmented automatically, there is place for doubt about the accuracy of the collected data for \texttt{/Q/}.
+
+However, unlike pauses we did consider it  for analysis, as it is has an important communicative function as a delimiter for words and morphemes.
+It is in fact one of the important clues which help the hearer to understand speech \cite{Ternes2012}.
+
+The information collected from our database show a very large variation of it's duration from 0.04 sec.
+to 2.05 sec., with $\sigma = 0.06$, consistent with the official data on the complete VMII.
+However, if we remove the upper 0.8 \% of the duration data, we obtain a dataset having only half of the mentioned variation with $\sigma = 0.03$ and a maximum duration of 0.3 seconds.
 \section{Consonants}
+\begin{itemize}
+	\item Plosives, which are generally short, is due to their articulation characteristics.
+	\item Fricatives, which show a legth variation almost comparable to that of vowels.
+	\item Nasals: /m/, /n/, /N/.
+	\item Specials: /h/, /l/, /C/, /j/, /r/
+\end{itemize}
 

### 2017-03-19 Paper/Paper_v1.tex
Some more additions to paper.

@@
 \usepackage{caption}
+\setcapindent{0pt}
 \usepackage{subcaption}
@@
 The result should be an approximation of the phoneme length change which is close to the actual change that occurs when the speaking rate alters.
+This would be the same as playing an audio file faster or slower.
+The purpose is therefore not to find another duration approximation model for TTS-systems, but to improve the algorithm of phoneme duration adjustment, given a known reference phoneme and word duration, and a target word duration.
 Consequently, the modified speech unit would sound more natural at a faster or a slower speech rate than being modified with a linear function.
@@
 The total VMII corpus used amounts to 11783 conversational turns representing recorded speech time of 15,5 hours.
-The ``symbolic data`` corresponding to each turn (e.g.phonetic transcription, segmentation, labeling, etc.) is captured in so-called ``par files``, representing files in the ``BAS Partitur Format`` \cite{Burger}.
+The ``symbolic data`` corresponding to each turn (e.g.phonetic transcription, segmentation, labeling, etc.) is captured in so-called ``par files``, representing files in the ``BAS Partitur Format`` \cite{Burger2000}.
 
@@
 The exact coordinates of the phoneme durations vary across speakers, as one can see in Fig.
-\ref{fig:speaker_cmp} for the case of /a:/.
+\ref{fig:speaker_cmp} for the case of \texttt{/a:/}.
 
@@
 	\centering
-	\caption[Phoneme length variation between speakers]{Overview of /a:/ length on the 5 most frequent speakers.
+	\caption[Phoneme length variation between speakers]{Overview of \texttt{/a:/} length on the 5 most frequent speakers.
 Phoneme length is expressed as the natural logarithmus of the sample frequency.
@@
 
-Pauses may be caused and/or quantitatively modified both by linguistic (e.g.
+Pauses may be caused and/or be quantitatively modified both by linguistic (e.g.
 phonological) and non-linguistic factors, like technical, psychological, dialog dynamics, etc.
 Consequently pause length variation needs to be treated separately from phoneme duration variation, as a not negligeable large set of other factors are to be considered for this, which extends beyond the purpose of this work.
-Moreover, we omitted pauses at the beginning and at the end of the turns completely from our analysis, as we supposed the influence of non-linguistic factors which we cannot evaluate whithin the limits of this work to be greater in these positions.
+
+Consistent with the study of Rosen \cite{Rosen2005} on the application of the lognormal distribution for modeling the variation of speech segment duration, we opted for a log-transformation of the observed phoneme durations using the natural logarithm, which proved to be a better description of the data in our database as well (fig.
+\ref{fig:aI_distrib}).
 
@@
 
-Consistent with the study of Rosen \cite{Rosen2005} on the application of the lognormal distribution for modeling the variation of speech segment duration, we opted for a log-transformation of the observed phoneme durations using the natural logarithm, which proved to be a better description of the data in our database as well.
-(Include grafics?!)
-
-Long a shows a strong negative correlation with the local speech rate for secondary stressed and unstressed instances.
-
 There have been presented in the literature many influencing factors for the duration of phonemes, which vary not only from one author to another but also from one language to another.
-For the German language, intuition aside, we have information on such factors provided by: Kohler \cite{Kohler1992}, Riedi \cite{Riedi1995}, ...
+For the German language, intuition aside, we have information on such factors provided by: Kohler \cite{Kohler1992}, Riedi \cite{Riedi1995}, Moebius \cite{Moebius1996}, and Brinckmann \& Trouvain \cite{Brinckmann_2003}, who applied some of the existent segment duration models to the German language.
 
@@
 \section{Vowels}
+Vowels are an important class, because their variation in terms of duration correlates well with the speech rate.
+One can say that vowels generally tend to get proportionally longer when the words themselves are spoken slower.
+We obtained an overall correlation coefficient of 0.72 for vowels against speech rate.
+Inside the vowel group we could identify the primary stressed vowels as the actual correlation owners, with a group correlation value of 0.75, while the unstressed and secondary stressed ones showed rather modest correlation values of 0.56, resp.
+0.47.
+
+As long vowels usually carry the primary stress, as some german phonetics manuals such as the one of Kohler \cite{Kohler1995} suggest, we considered the long vowels as a standalone subclass of vowels.
+The correlation value for the long vowel group is even greather than that of the primary stressed one, amounting to 0.79.
+
 \begin{itemize}
-	\item Diphthongs: when two vowels are articulated together in a syllable. They are generally characterized by longer durations. German knows three diphthongs: /aI/, /aU/ and /OY/.
+	\item Diphthongs: when two vowels are articulated together in a syllable. They are generally characterized by longer durations. German knows three diphthongs: \texttt{/aI/, /aU/} and \texttt{/OY/}.
 	\item Schwa: most frequent vowel in German as well as in English.
 It is also called a reduction vowel \cite{Kohler1995}, as most of the other monophthongs tend to be articulated as schwa in non-stressed vowels.
-Verbmobil differentiates two types of schwa for the German language: /@/ like in the last syllable of \texttt{lesen} and /6/ like in \texttt{Leser}, where the reduced vowel is followed by an /r/.
+Verbmobil differentiates two types of schwa for the German language: \texttt{/@/} like in the last syllable of \texttt{lesen} and \texttt{/6/} like in \texttt{Leser}, where the reduced vowel is followed by an \texttt{/r/}.
 	\item Monophthongs, which may be further split according to the short - long pairs or based on the articulation point like open - closed vowels or back - front vowels.
@@
 \section{Glottal stop}
-The glottal stop occurs only word initially in our database, and was marked with \texttt{/Q/}.
+The glottal stop occurs 13 348 times and only word initially in our database, marked with \texttt{/Q/}.
 We decided to treat it separately because of its particular usage in German: it may occur only at the beginning of a word or a word stem and in front of a vowel.
 Some phonetic manuals don't event consider it a phoneme \cite{Ternes2012}.
-Furthermore, it's acoustical properties make it look like a pause in speech, as you may see in Fig.
+Furthermore, it's acoustical properties make it look like a (filled) pause in speech, as you may see in Fig.
 ..., which represents a challenge for the transcription.
@@
 However, if we remove the upper 0.8 \% of the duration data, we obtain a dataset having only half of the mentioned variation with $\sigma = 0.03$ and a maximum duration of 0.3 seconds.
+We checked manually some of the largest outliers and they proved to be a result of wrong segmentation.
+Their real duration was within the first 2 thirds of the data.
 \section{Consonants}
@@
 It has been suggested, that the influence of so-called ``macroscopic`` \cite{Cummins1999} or para-linguistic \cite{Santen1994} factors should be examined in order to move this asymptote further up.
-We focused on the speech rate.
+We focused on speech rate.
 
@@
 Larger the segment size, larger the number of influencing factors for phoneme duration.
-We opted for a progressive approach, starting at the syllable level.
+We opted for a progressive approach, starting at the syllable level ?.
+
+Our approach is based on the observation that specific phonemes, respectively phoneme groups, occupy a specific percent of the total word duration (fig.
+\ref{fig:boxplot_genau}), and that certain phonemes seem to modify their duration according to that of the segment they're in to a much higher degree than others, as you can see in fig.
+\ref{fig:w_dur_vs_pho}.
+
+\begin{figure}[htbp]
+	\includegraphics[width=\textwidth]{../Graphen/Stats_for_genau.eps}
+	\centering
+	\caption[Word duration of ``genau`` vs.
+its phoneme durations]{We ordered ascendingly the 706 occurences of the word ``genau`` based on word duration and ploted recorded durations of the component phonemes against it.
+The picture suggests clearly that the diphthong \texttt{/aU/} occupies a rather fixed steak of the total word duration, while the others show no adjustment to the increasing word total duration.
+}
+	\label{fig:w_dur_vs_pho}
+\end{figure}
+
+\begin{figure}[htbp]
+	\includegraphics[width=.55\linewidth]{../Graphen/Sound_boxplot_genau.eps}
+	\centering
+	\caption[Phoneme proportions in ``genau``]{Proportions taken by the component phonemes of ``genau`` suggest the possibility of grouping them in disjunct classes.
+Here we explored the phoneme distribution for a relatively short occurence of our reference word, of only 0.3 sec found 25 times.
+This one, as well as others, correlates well with the overall phoneme distribution on all occurences.}
+	\label{fig:boxplot_genau}
+\end{figure}
 
 \chapter{Conclusion and Outlook}
+Our model calculates an approximation for the phoneme duration when its corresponding speech segment changes in terms of duration.
+It does this no matter what was the cause of the duration change in the segment.
+However, when speech rate changes, there are also other speech characteristics, like prosody and phoneme quality which change, and which cannot be simulated only by modifying the phoneme duration.
+Therefore, making speech sound natural while modifying the speech rate needs to consider and adjust these other features too.
+
 Analyze duration changes of vowels in interrupted words.

### 2017-03-21 Paper/Paper_v1.tex
Added more code for exploring the phoneme dynamics in words. Still working on it.

@@
 Their real duration was within the first 2 thirds of the data.
+
 \section{Consonants}
+Based on the correlation between phoneme duration and speech rate we created a subgroup of consonants containing the nasals and \texttt{/x/, /C/, /h/} with a relatively good correlation value $\geq 0.7$, and considered plosives to be a subgroup because of their consistently low correlation $\leq 0.42$.
 \begin{itemize}

### 2017-03-21 py_files/arff_file_builder.py

@@ -6,9 +6,13 @@ import sys
 import time
 
 # Generates an almost ready for processing with weka text file. 
-
+time1 = time.time()
 # List of phonemes to search for
-phoneme_list = ["Q"]
+#phoneme_list = ["a", "a:", "E", "E:", "e", "e:", "O", "o", "o:", "@", "6", "9"] # needs 4,5 hrs
+#phoneme_list = ["p", "b", "t", "d", "k", "g"] # needs cca 3hrs (2hrs 45 min)
+#phoneme_list = ["f", "v", "s", "S", "z", "x", "h", "C", "Z"] # ca. 3hrs (3hrs 14 min)
+#phoneme_list = ["m", "n", "N", "l", "r", "j"] # ca. 3 hrs 40 min
+phoneme_list = ["@", "6"] # ca. 1hr 15 min
 # All diphthongs:
 #phoneme_list = ["OY", "aU", "aI"]
 # All long vowels:
@@ -30,14 +34,14 @@ word_list = set(dict_list[3::13])
 
 
 #test_file = open("C:/Users/alexutza_a/Abschlussarbeit/sp_kurzes_a.txt", "w") -> gibt falsche Anzahl von speech rates
-arff_file = open("C:/Users/alexutza_a/Abschlussarbeit/wekadateien/Q.txt", "w")
+arff_file = open("C:/Users/alexutza_a/Abschlussarbeit/wekadateien/schwa_usw", "w")
 
-arff_file.write("% ARFF file for Verbmobil\n% Phoneme: all_vowels_noSchwa\n%\n@relation glottal_stop\n\n% List of attributes:\n%\n")
+arff_file.write("% ARFF file for Verbmobil\n% Phonemes: @, 6\n%\n@relation schwa\n\n% List of attributes:\n%\n")
 arff_file.write("% Duration is expressed in seconds\n")
 
-arff_file.write("@attribute phoneme { Q") 
-#for el in phoneme_list:
-#	arff_file.write(el + ", ")
+arff_file.write("@attribute phoneme { ") 
+for el in phoneme_list:
+	arff_file.write(el + ", ")
 arff_file.write( " }\n")
 
 arff_file.write("@attribute filename string\n")
@@ -53,7 +57,7 @@ arff_file.write("@attribute overlapping { 1, 0 }\n")
 arff_file.write("@attribute speech_rate_pho numeric\n")
 arff_file.write("@attribute local_speech_rate numeric\n")
 arff_file.write("@attribute speech_rate_msyl numeric\n")
[…]

### 2017-03-21 py_files/phon_dict.py

@@ -7,6 +7,7 @@ import stress
 import speech_rate
 import math
 import find_syl_position
+import time
 
 # Creates a dictionary of durations (of specific phoneme)
 dur_file = open("C:/Users/alexutza_a/Abschlussarbeit/word_durations.txt", "w")
@@ -24,6 +25,7 @@ def verbmo_par_files():
 	return path_list#,file_list
 path_list = verbmo_par_files()
 
+t1 = time.time()
 # Dictionary of the given phoneme instance length in ms (keys) in Verbmobil and its attributes (values in list)
 # Attributes are: Filename, Position in Speech (Start, Middle, End), Containing word, Stress, Overlapping, Speech rate
 def phon_dict(phoneme_list):
@@ -61,7 +63,7 @@ def phon_dict(phoneme_list):
 					# Attribute: pause = if a pause preceeds or follows the word in which the phoneme occurs: near_pause, no_pause
 					laut_schluessel[phoneme].append(is_pause_near(file, wort))
 					# Attribute: stress = type of stress: none, s_stress, p_stress
-					laut_schluessel[phoneme].append(stress.stress_type(file, wort, phoneme, (zeile+10))) #modified zeile from +9 to +10, just for Q
+					laut_schluessel[phoneme].append(stress.stress_type(file, wort, phoneme, (zeile+9))) #modified zeile from +9 to +10, just for Q
 					# Attribute: overlapping
 					if wort in overlapping_list:
 						laut_schluessel[phoneme].append("1")
@@ -77,9 +79,9 @@ def phon_dict(phoneme_list):
 					# Attribute: speech rate as total syllables / sec
 					#laut_schluessel[phoneme].append(speech_rate_ksyl)
 					word_duration, phon_count, syl_count = speech_rate.word_duration(file, int(line.split()[3]))
-					laut_schluessel[phoneme].append(round((word_duration*0.00000625)/phon_count, 4))
+					laut_schluessel[phoneme].append(round((word_duration*0.0000625)/phon_count, 4))
 					# Attribute: word position: initial, not initial - only for Q
-					laut_schluessel[phoneme].append(stress.find_qPosition(file, (zeile+9)))
+					#laut_schluessel[phoneme].append(stress.find_qPosition(file, (zeile+9)))
[…]

### 2017-03-21 py_files/speech_rate.py

@@ -144,7 +144,9 @@ def speech_rate_list():
 		datei.close()
 	return spl
 
-#print(word_duration("C:/Users/alexutza_a/Abschlussarbeit/DB_Verbmobil/verbmobil_par/g016a/g016acn1_002_ABE.par", 0))
+#word_duration, phon_count, syl_count = word_duration("C:/Users/alexutza_a/Abschlussarbeit/DB_Verbmobil/verbmobil_par/g217a/g217acn1_004_AHP.par", 2)
+#print(word_duration("C:/Users/alexutza_a/Abschlussarbeit/DB_Verbmobil/verbmobil_par/g217a/g217acn1_004_AHP.par", 2))
+#print(round((word_duration*0.0000625)/phon_count, 4))
 #print(speech_rate("C:/Users/alexutza_a/Abschlussarbeit/DB_Verbmobil/verbmobil_par/g016a/g016acn1_000_ABE.par"))
 #print(speech_rate("C:/Users/alexutza_a/Abschlussarbeit/DB_Verbmobil/verbmobil_par/g016a/g016acn1_016_ABE.par"))
 

### 2017-03-21 py_files/stress.py

@@ -8,6 +8,14 @@ short_vowels = ["a", "e", "E", "I", "i", "O", "o", "U", "u", "Y", "y", "9", "2"]
 def stress_type(datei, word_number, phoneme, cursor):
 	work_file = open(datei)
 	word_no = int(word_number)
+	z = 0
+#	Instructions for Q to get the stress of the following phoeneme
+#	for line in work_file:
+#		z += 1
+#		if z == (cursor):
+#			phoneme = line.split()[4]
+			#print(phoneme)
+		
 	mau_rank = p_rank_in_mau(work_file, word_no, phoneme, cursor)
 	stress_type = ""
 	word = ""
@@ -235,3 +243,4 @@ def p_rank_in_mau(datei, word_number, phoneme, cursor):
 #print(p_rank_in_mau("C:/Users/alexutza_a/Abschlussarbeit/DB_Verbmobil/verbmobil_par/g419a/g419acn1_012_ALW.par", 9, "i:", 155))
 #print(set_stress("C:/Users/alexutza_a/Abschlussarbeit/DB_Verbmobil/verbmobil_par/g419a/g419acn1_012_ALW.par", "definit'i:f", "i:", 9, 0, 155))
 #print(stress_type("C:/Users/alexutza_a/Abschlussarbeit/DB_Verbmobil/verbmobil_par/g419a/g419acn1_012_ALW.par", 9,  "i:", 155))
+#print(stress_type("C:/Users/alexutza_a/Abschlussarbeit/DB_Verbmobil/verbmobil_par/g419a/g419acn1_016_ALW.par", 4, "Q", 116))

### 2017-03-27 py_files/model_ipdur.py
Some corrections and added diagram of relevant py-files.

@@ -1,0 +1,53 @@
+import numpy as np
+from collections import defaultdict
+import model_utilities
+
+# This is a relatively static phoneme duration prediction model which only considers the median (or mean) value of a specific phoneme
+#
+# Predicted value = median or mean of phon_dur in provided (training) dataset
+
+###################################################### START ##################################################################
+
+############## STATIC DATA STRUCTURES (not dataset specific) ##################################
+
+
+# Create list of filepaths to explore. This one uses entire data set, for exploration purposes.
+# path_list_training = verbmo_par_files('C:/Users/alexutza_a/Abschlussarbeit/DB_Verbmobil/verbmobil_par')
+path_list_training = model_utilities.get_path_list('C:/Users/alexutza_a/Abschlussarbeit/DB_Verbmobil/Evaluation/Training')
+
+valid_phonemes = ["a", "a~", "e", "E", "I", "i", "O", "o", "U", "u", "Y", "y", "9", "2", "a:", "a~:", "e:", "E:", "i:",
+                 "o:", "u:", "y:", "2:", "OY", "aU", "aI", "@", "6", "z", "S", "Z", "C", "x", "N", "Q", "b", "d", "f", 
+                 "g", "h", "j", "k", "l", "m", "n", "p", "r", "s", "t", "v"]
+
+#################################### DYNAMIC DATA STRUCTURES (training data) ################################
+
+# Processed version of the phon_dur_dict (defined in module model_utilities): 
+# Returns: a flat dictionary. Looks like: 
+# {"a" : [median_1, mean_1, (m_1 + m_1) /2], "b" : [median_2, mean_2, (m_2 + m_2) /2], ...}
+def phon_dur_compressed_dict(phon_dur_dict):
+	phon_dur_compressed_dict = dict( (i, []) for i in phon_dur_dict.keys())
+
+	for phon in valid_phonemes:
+		phon_dur_compressed_dict[phon] = [round(np.median(phon_dur_dict[phon]), 3)]
+		phon_dur_compressed_dict[phon].append(round(np.mean(phon_dur_dict[phon]), 3))
+		phon_dur_compressed_dict[phon].append(round((np.median(phon_dur_dict[phon]) + np.mean(phon_dur_dict[phon]))/2, 3))
+	#print(round(np.median(phon_dur_dict["v"]), 3))
+	#print(round(np.median(phon_dur_dict["aU"]), 3))
[…]

### 2017-03-28 Paper/Paper_v1.tex
Readme and paper changes.

@@
 \usepackage{subcaption}
+\usepackage{scrhack}
+\usepackage{tabularx}
+
+\newcolumntype{L}[1]{>{\raggedright\arraybackslash}m{#1}}
+\newcolumntype{C}[1]{>{\centering\arraybackslash}m{#1}}
+\newcolumntype{R}[1]{>{\raggedleft\arraybackslash}m{#1}}
 
@@
 \chapter{Introduction}
+Some people speak faster, others speak slower, and all people speak sometimes faster or slower as they usually do.
+This ``speed`` at which one speaks is what we call ``speech rate``.
+The fact that we understand people at all natural speech rates is certainly largely due to the quality features of sounds (phonemes) in the given language.
+However, if we have a perfect audio recording of speech, and play it faster or slower, we immediatly notice how naturalness deteriorates.
+Synthetic speech shows a similar behaviour when changing the speaking rate.
+In both cases phoneme quality doesn't need to change.
+It is its duration that changes, along with some other suprasegmental features like pitch, phrasal accents, etc., and that do the trick.
+
 The attempts to improve the quality of speech synthesis when tempo changes include linguistic processing, and phoneme length manipulation before the actual synthesis takes place.
 As mentioned in a patent documentation of Fujitsu Limited, phoneme length manipulation plays an important role in this context and one cannot rely on a linear adjustment function \cite{nishiike2008}.
+
+\subsection*{Scope}
 Phoneme length adjustment is the aspect we wish to adress in this work.
-In particular, we want to find an algorithm to modify the phoneme length which takes into consideration the local speech rate.
-The result should be an approximation of the phoneme length change which is close to the actual change that occurs when the speaking rate alters.
+In particular, we want to find an algorithm to adjust the phoneme duration according to the speech rate.
+The resulting phoneme duration should be close to the actual duration that occurs when the speaking rate alters.
 This would be the same as playing an audio file faster or slower.
-The purpose is therefore not to find another duration approximation model for TTS-systems, but to improve the algorithm of phoneme duration adjustment, given a known reference phoneme and word duration, and a target word duration.
+
+There are several models for phoneme duration approximation, that work pretty well for some languages, German inclusive.
+However, they consistently don't take the speech rate into consideration, their performance level being acquired within a specific speech rate.
+When this varies, model performance deteriorates, like van Santen states for his model \cite{Santen1994}.
+Moreover, all these models have an upper asymptote for their accuracy situated at less than 100\%.
+It has been suggested, that the influence of so-called ``macroscopic`` \cite{Cummins1999} or para-linguistic \cite{Santen1994} factors should be examined in order to move this asymptote further up.
+So we focused on speech rate.
+
+The purpose is therefore not to find another duration approximation model for TTS-systems, but to improve the algorithm of phoneme duration adjustment, given known reference phoneme and word durations, and a target word duration, as presented in Table \ref{tab:Question}.
 Consequently, the modified speech unit would sound more natural at a faster or a slower speech rate than being modified with a linear function.
 
-Hoequist and Kohler observed already back in 1986 that the change in speaking tempo does not produce a linear change of the acoustic segments of an utterance (i. e. phonemes) \cite{Hoequist1986}.
+\begin{table}[htbp]
+\caption {The challenge} 
+\centering
+\begin{tabular}{|L{2.5cm}||*{4}{C{.8cm}|}|*{4}{C{.8cm}|}}
+	\hline
+			& \multicolumn{4}{c||}{ \textit{genau}} & \multicolumn{4}{c|}{ \textit{genau}} \\
+	\hline
+Phonemes		& g & @ & n & aU & g & @ & n & aU \\
+	\hline
+Durations (sec) & 0.05 & 0.07 & 0.05 & 0.11 & ?  & ?  & ?  & ? \\
+	\hline
+Speech rate $dur/\#phon$ & \multicolumn{4}{c||}{0.07} & \multicolumn{4}{c|}{0.11} \\
+	\hline
+\end{tabular}
+\caption*{The first column shows our inputs: duration of each phoneme in German word ``genau``, given a speech rate of 0.07 sec, or a word duration of 0.28 sec.
+The second column presents our challenge: how long is the duration of each phoneme of the given word, when it is spoken slower, taking e.g.
+0.42 sec, speech rate of approx.
+0.11 sec?}
+\label{tab:Question} 
+\end{table}
 
+Hoequist and Kohler observed already back in 1986 that the change in speaking tempo does not produce a linear change of the acoustic segments of an utterance (i.
+e.
+phonemes) \cite{Hoequist1986}.
+Indeed, if one simply ``stretches`` all phonemes in the same manner to accomodate a new speech rate, the result is rather dissapointing, as one can see in Fig.
+\ref{fig:linear_approx}.
+
+\begin{figure}[htbp]
+	\centering
+	\includegraphics[scale=0.7]{../Graphen/Linear_approx.eps}
+	\caption[Linear phoneme duration approximation]{This is a linear solution to the issue presented in table \ref{tab:Question} for the German word ``genau``}
+	\label{fig:linear_approx}
+\end{figure}
+
+\subsubsection*{Notizen}
 Syllable and foot structure of the utternace belong to the important factors influencing the segment length modification  \cite{Hoequist1986}.
@@
 Accross the paper we will use the German SAMPA notations when referring to phonemes.
-\section{Goals}
-\section{Outline}
+
+\subsection*{Outline}
+We start our approach by defining our methods in chapter 2.
+In the following chapter we examine our database so that we can proceed with the actual task in chapter 4, where we analyze the German phonemes.
+We dedicate chapter 5 to the challenge of finding an adequate speech rate definition for our database and purpose.
+Chapter 6 finally deals with phoneme duration prediction models and the challenge of finding a solution to our above presented issue.
 
@@
 \section{Machine Learning}
-Machine Learning means building a program (model) which based on data mining results from a given dataset can make predictions about an unseen dataset  (fill in missing information). 
+Machine Learning means building a program or model which can make predictions about an unseen dataset, like filling in missing information, based on data mining results from a given dataset. 
 \section{Preparing the Data}
-A very important and time consuming step in data mining is data preparation. 
+A very important and time consuming step in data mining is data preparation, which is the next step after data collection.
+At this stage one needs to asses the quality of the data, expressed as noise ratio, and strategies for dealing with it.
+If the data used has been originally collected for another purpose, then it needs to be assessed in terms of features it contains, actually needed features, and organization resp.
+reorganization of data to serve new purpose.
+
+Considering machine learning, it is straightforward that one needs to split the available dataset at least into two parts: a training dataset, used for discovering patterns in data through data mining, and a test dataset, used for testing the performance of the model resulting from the machine learning process.
+
+We explain how we prepared our data in chapter 3, dedicated to our database.
+
 \section{Performance Evaluation}
-Considering machine learning, it is straightforward that one needs to split the available dataset at least into two parts: a training dataset, used for discovering patterns (data mining) in data, and a test dataset, used for testing the performance of the model (result of machine learning process).
-Model performance may be evaluated in several ways. We decided for the following three:
+Models produced by machine learning may have different performance levels, which may be evaluated in several ways. We decided for the following three:
 \begin{enumerate}
@@
 \chapter{Corpora}
-\section{Verbmobil}
-The Verbmobil corpus is a database of spontaneous speech containing a collection of appointment making dialogs, fully transliterated and annotated.
-This was an important point in selecting this database for our purpose, as phoneme duration is influenced by many factors, and the Verbmobil textual database provides most of them.
-We used 286 recordings from the last verbmobil phase (Verbmobil II) created by 131 speakers.
-The total VMII corpus used amounts to 11783 conversational turns representing recorded speech time of 15,5 hours.
-The ``symbolic data`` corresponding to each turn (e.g.phonetic transcription, segmentation, labeling, etc.) is captured in so-called ``par files``, representing files in the ``BAS Partitur Format`` \cite{Burger2000}.
+Segmented and annotated German speech corpora represent sets of text files, so-called ``textual data``, containing information about the speech data.
+Our minimum requirement for such corpora was that they contain speech segmented at phoneme level, allowing us to calculate the duration of each phoneme occurence.
+The unit in which this duration is expressed is not relevant.
 
-Drawbacks of this database come mainly from the characteristics of spontaneous speech.
-Following natural rules, speech rate often varies significantly inside a speech turn.
-One starts his turn at a normal speech rate, then gets to a point where he/she hesitates and gets slower, like when weighting different transport means to get to the appointment they are to agree upon.
-This results in the challenge of finding an appropriate description for the speech rate to be used for modeling phoneme durations across the corpus.
-We handle this issue of speech rate separately in chapter 5.
-Another challenge resulting from the structure of this corpus as a collection of dialogs is speaker variation.
-The exact coordinates of the phoneme durations vary across speakers, as one can see in Fig.
-\ref{fig:speaker_cmp} for the case of \texttt{/a:/}.
-
-\begin{figure}[htbp]
-	\includegraphics[width=0.6\textwidth]{../Graphen/Duration_of_a_5speakers.eps}
-	\centering
-	\caption[Phoneme length variation between speakers]{Overview of \texttt{/a:/} length on the 5 most frequent speakers.
-Phoneme length is expressed as the natural logarithmus of the sample frequency.
-While the minimum duration is technically limited to an equal value corresponding to 0.04 sec, all other coordinates vary, e.g.
-maximum value has a variation of 0.14 sec between speakers AKX and ALR.
-The greatest median variation is 0.07 sec between speakers ALR (0.15 sec) and AJZ (0.08 sec).}
-	\label{fig:speaker_cmp}
-\end{figure}
-
-\chapter{Phoneme durations and the dynamics of speech}
-The phoneme inventory used in Verbmobil includes 52 classes defined according to the German SAMPA.
-Using the classical phonetic conventions, we can group these phonemes according to their phonetic features using a tree structure into vowels, consonants and a glottal stop on a first level, although glottal stop is usually treated as a consonant.
-Further relevant groupings are being examined in the sections ``Vowels`` and ``Consonants``.
-
-Pauses may be caused and/or be quantitatively modified both by linguistic (e.g.
-phonological) and non-linguistic factors, like technical, psychological, dialog dynamics, etc.
-Consequently pause length variation needs to be treated separately from phoneme duration variation, as a not negligeable large set of other factors are to be considered for this, which extends beyond the purpose of this work.
-
-Consistent with the study of Rosen \cite{Rosen2005} on the application of the lognormal distribution for modeling the variation of speech segment duration, we opted for a log-transformation of the observed phoneme durations using the natural logarithm, which proved to be a better description of the data in our database as well (fig.
+For further processing of this information from our side, we considered the study of Rosen \cite{Rosen2005} on the application of the lognormal distribution for modeling the variation of speech segment duration.
+In accordance to this, during the data mining tasks we opted for a log-transformation of the observed phoneme durations using the natural logarithm, which proved to be a better description of the data in our database as well (fig.
 \ref{fig:aI_distrib}).
@@
 
-There have been presented in the literature many influencing factors for the duration of phonemes, which vary not only from one author to another but also from one language to another.
+\section{Verbmobil}
+The Verbmobil corpus is a database of spontaneous speech containing a collection of appointment making dialogs, fully transliterated and annotated.
+This was an important point in selecting this database for our purpose, as phoneme duration is influenced by many factors, and the Verbmobil textual database provides most of them.
+We used 286 recordings from the last verbmobil phase (Verbmobil II) created by 131 speakers.
+The total VMII corpus used amounts to 11783 conversational turns representing recorded speech time of 15,5 hours.
+The ``symbolic data`` corresponding to each turn (e.g.
+phonetic transcription, segmentation, labeling, etc.) is captured in so-called ``par files``, representing files in the ``BAS Partitur Format`` \cite{Burger2000}.
+
+Drawbacks of this database come mainly from the characteristics of spontaneous speech.
+Following natural rules, speech rate often varies significantly inside a speech turn.
+This means one starts his turn at a normal speech rate, then gets to a point where he/she hesitates and gets slower, like when weighting different transport means to get to the appointment they are to agree upon.
+This results in the challenge of finding an appropriate description for the speech rate to be used for modeling phoneme durations across the corpus.
+We handle this issue of speech rate separately in chapter 5.
+Another challenge resulting from the structure of this corpus as a collection of dialogs is speaker variation.
+The exact coordinates of phoneme durations vary across speakers, as one can see in Fig.
+\ref{fig:speaker_cmp} for the case of \texttt{/a:/}.
+
+\begin{figure}[htbp]
+	\includegraphics[width=0.6\textwidth]{../Graphen/Duration_of_a_5speakers.eps}
+	\centering
+	\caption[Phoneme length variation between speakers]{Overview of \texttt{/a:/} length on the 5 most frequent speakers.
+Phoneme length is expressed as the natural logarithmus of the sample frequency.
+While the minimum duration is technically limited to an equal value corresponding to 0.04 sec, all other coordinates vary, e.g.
+maximum value has a variation of 0.14 sec between speakers AKX and ALR.
+The greatest median variation is 0.07 sec between speakers ALR (0.15 sec) and AJZ (0.08 sec).}
+	\label{fig:speaker_cmp}
+\end{figure}
+
+\chapter{Phoneme durations and the dynamics of speech}
+The phoneme inventory used in Verbmobil includes 52 classes defined according to the German SAMPA.
+Using the classical phonetic conventions, we can group these phonemes on a first level according to their phonetic features using a tree structure into vowels, consonants and a glottal stop, although the latter one is sometimes treated as a consonant \cite{Kohler1995}.
+Further relevant groupings are being examined under sections ``Vowels`` and ``Consonants``.
+
+Pauses may be caused and/or be quantitatively modified both by linguistic (e.g.
+phonological) and non-linguistic factors, like technical, psychological, dialog dynamics, etc.
+Consequently pause length variation needs to be treated separately from phoneme duration variation, as a not negligeable large set of other factors are to be considered for this, which extends beyond the purpose of this work.
+
+The literature records on many potential influencing factors for the duration of phonemes, which vary not only from one author to another but also from one language to another.
 For the German language, intuition aside, we have information on such factors provided by: Kohler \cite{Kohler1992}, Riedi \cite{Riedi1995}, Moebius \cite{Moebius1996}, and Brinckmann \& Trouvain \cite{Brinckmann_2003}, who applied some of the existent segment duration models to the German language.
@@
 
+We identified following vowel categories as being relevant for our purpose:
+
+\begin{itemize}
+	\item Diphthongs represent two vowels that are articulated together in a syllable.
+They are generally characterized by longer durations.
+German knows three diphthongs: \texttt{/aI/, /aU/} and \texttt{/OY/} as in \textit{Haus} \texttt{/haUs/},  \textit{heiß} \texttt{/haIs/} and \textit{moin} \texttt{/mOYn/}.
+We considered them as a separate group, as they are generally longer than all other phonemes.
+	\item Schwa: most frequent vowel in German as well as in English.
+It is also called a reduction vowel \cite{Kohler1995}, as most of the other monophthongs tend to be articulated as schwa in non-stressed vowels.
+Verbmobil differentiates two types of schwa for the German language: \texttt{/@/} like in the last syllable of \textit{lesen} and \texttt{/6/} like in \textit{Leser}, where the reduced vowel is followed by an \texttt{/r/}.
+	\item Long vowels.
 As long vowels usually carry the primary stress, as some german phonetics manuals such as the one of Kohler \cite{Kohler1995} suggest, we considered the long vowels as a standalone subclass of vowels.
 The correlation value for the long vowel group is even greather than that of the primary stressed one, amounting to 0.79.
-
-\begin{itemize}
-	\item Diphthongs: when two vowels are articulated together in a syllable. They are generally characterized by longer durations. German knows three diphthongs: \texttt{/aI/, /aU/} and \texttt{/OY/}.
-	\item Schwa: most frequent vowel in German as well as in English.
-It is also called a reduction vowel \cite{Kohler1995}, as most of the other monophthongs tend to be articulated as schwa in non-stressed vowels.
-Verbmobil differentiates two types of schwa for the German language: \texttt{/@/} like in the last syllable of \texttt{lesen} and \texttt{/6/} like in \texttt{Leser}, where the reduced vowel is followed by an \texttt{/r/}.
-	\item Monophthongs, which may be further split according to the short - long pairs or based on the articulation point like open - closed vowels or back - front vowels.
-We shall verify the relevance of these subcategories for our purpose in the section dedicated to vowels.
+	\item Short vowels in opposition to the above mentioned vowel categories.
 \end{itemize}
@@
 
-However, unlike pauses we did consider it  for analysis, as it is has an important communicative function as a delimiter for words and morphemes.
-It is in fact one of the important clues which help the hearer to understand speech \cite{Ternes2012}.
+However, unlike pauses we did consider it  for analysis, as it is has an important communicative function as delimiter for words and morphemes.
+It is in fact one of the important clues which help hearer understand speech \cite{Ternes2012}.
 
@@
 Based on the correlation between phoneme duration and speech rate we created a subgroup of consonants containing the nasals and \texttt{/x/, /C/, /h/} with a relatively good correlation value $\geq 0.7$, and considered plosives to be a subgroup because of their consistently low correlation $\leq 0.42$.
+
+A big surprize in the Verbmobil dataset was the huge duration variation of the plosive /t/, which was even greater than most of the fricatives.
+Evaluation of some samples showed this to be a result of wrong segmentation in the context /ts/, giving /t/ the longer duration and /s/ the shorter one.
+Another explanation would be that the phase before the release phase may indeed be relatively long.
+
+Below an overview of the classical consonant categories.
 \begin{itemize}
@@
 
-We decided to calculate global speech rate as a ratio of number of syllables per second.
-As syllables were not annotated in our database, we decided to use the number of (realized?) vowels as an approximation for the number of syllables \cite{Yishan_Jiao_2015} \cite{Kohler1995}.
-
-Koreman showed \cite{Koreman_2006} that the perceived speech rate is also influenced by the listener's knowledge of the expected articulations for a particular utterance, therefore it makes a difference if all expected phones are also articulated or not.
-Consequently, we did compare speech rate considering only articulated vowels/syllables, as well as considering both realized and intended syllables, and tested both methods in our model.
-Our tests showed that ...
+(We decided to calculate global speech rate as a ratio of number of syllables per second.
+As syllables were not annotated in our database, we decided to use the number of (realized?) vowels as an approximation for the number of syllables \cite{Yishan_Jiao_2015} \cite{Kohler1995}.)
 
@@
 denominator) alone, varied it, and tried to figure out what happends in the numerator so that the proportion stays.
-Expressed mathematically, if the durations of the constituent phonemes are $d_1$, $d_2$, and $d_3$, \textit{SR} is the resulting speech rate and \textit{time} represents the total duration of the 3 elements, this corresponds to the follwoing transformation:
-
-$$\frac{d_1+d_2+d_3}{time} = SR \iff (d_1+d_2+d_3)  = SR \cdot time $$
+In other words, what we do is to say that when the speech rate changes, the duration of changes as well.
+Now given one has an algorithm to predict the duration change of single words in a phrase when speech rate changes, we only care about phoneme duration change inside word of given durations.
 
@@
 However, they also report a strong influence of  the quality of the input data on the model performance.
-
-All these models have an upper asymptote for their accuracy situated at less than 100\%.
-It has been suggested, that the influence of so-called ``macroscopic`` \cite{Cummins1999} or para-linguistic \cite{Santen1994} factors should be examined in order to move this asymptote further up.
-We focused on speech rate.
 
@@
 
+Koreman showed \cite{Koreman_2006} that the perceived speech rate is also influenced by the listener's knowledge of the expected articulations for a particular utterance, therefore it makes a difference if all expected phones are also articulated or not.
+Consequently, one could compare speech rate considering only articulated vowels/syllables, as well as considering both realized and intended syllables, and test both methods with a given model.
+One possible result of such an approach would be the prediction of phonemes reaching $duration = 0$ at specific speech rates, which means predict phoneme drop phenomena.
+
+\subsubsection*{Notizen}
 Analyze duration changes of vowels in interrupted words.

### 2017-03-28 py_files/phon_dict.py

@@ -7,6 +7,7 @@ import stress
 import speech_rate
 import math
 import find_syl_position
+import model_utilities
 import time
 
 # Creates a dictionary of durations (of specific phoneme)
@@ -14,16 +15,7 @@ dur_file = open("C:/Users/alexutza_a/Abschlussarbeit/word_durations.txt", "w")
 speech_rate_list = []
 # List of all german .par files in verbmobil, where the recording was done acn (main scenario dialogue 
 	#recorded means neckband microphone)
-def verbmo_par_files():
-	pattern = 'g*acn*.par'  # Pattern to be used for filtering filenames
-	file_list = []			# Empty list to be populated with filenames matching pattern
-	path_list = []
-	for path, subfolder, filenames in os.walk('C:/Users/alexutza_a/Abschlussarbeit/DB_Verbmobil/verbmobil_par'):
-		for filename in fnmatch.filter(filenames, pattern):
-			file_list.append(filename)
-			path_list.append(os.path.join(path, filename))
-	return path_list#,file_list
-path_list = verbmo_par_files()
+path_list = model_utilities.get_path_list('C:/Users/alexutza_a/Abschlussarbeit/DB_Verbmobil/verbmobil_par')
 
 
 # Dictionary of the given phoneme instance length in ms (keys) in Verbmobil and its attributes (values in list)
@@ -64,7 +56,7 @@ def phon_dict(phoneme_list):
 					laut_schluessel[phoneme].append(is_pause_near(file, wort))
 					# Attribute: stress = type of stress: none, s_stress, p_stress
 					laut_schluessel[phoneme].append(stress.stress_type(file, wort, phoneme, (zeile+9))) #modified zeile from +9 to +10, just for Q
-					# Attribute: overlapping
+					# Attribute: overlapping (1- phoneme in overlapped speech, 0- not)
 					if wort in overlapping_list:
 						laut_schluessel[phoneme].append("1")
 					else:

### 2017-03-29 py_files/eval_model.py
Slight improvement of the eval_model class.

@@ -15,14 +15,13 @@ actuals_list2 = create_eval_lists.create_modelWLIP_lists(2)[1]
 predictions_list3 = create_eval_lists.create_modelIPdur_lists(2)[0]
 actuals_list3 = create_eval_lists.create_modelIPdur_lists(2)[1]
 
-
+model_name_list = ["WL", "WL_IP", "IP_dur"]
 evaluation_lists = [(predictions_list1, actuals_list1), (predictions_list2, actuals_list2), (predictions_list3, actuals_list3)]
 
-# Evaluate model with: RMSE, MAE, and Pearson's correlation coefficient
-global_rmse, mae = performance_measures.calc_error_for_data(predictions_list2, actuals_list2)
-corrCoef = performance_measures.corrCoef(predictions_list3, actuals_list3)[0]
-print(global_rmse, mae, corrCoef)
-
-#print(predictions_list[:5])
-#print(actuals_list[:5])
-
+# Evaluate models with: RMSE, MAE, and Pearson's correlation coefficient
+for i in range(len(evaluation_lists)):
+	global_rmse, mae = performance_measures.calc_error_for_data(evaluation_lists[i][0], evaluation_lists[i][1])
+	corrCoef = performance_measures.corrCoef(evaluation_lists[i][0], evaluation_lists[i][1])[0]
+	print("RMSE for the" + model_name_list[i] + "-model: " + str(global_rmse))
+	print("MAE for the" + model_name_list[i] + "-model: " + str(mae))
+	print("Pearson corrCoef for the " + model_name_list[i] + "-model: "  + str(corrCoef))

### 2017-03-29 py_files/eval_model.py
Correction of model_wl and another improvement of eval_model

@@ -2,21 +2,17 @@ import performance_measures
 import numpy as np
 import create_eval_lists
 
-# Create the two lists of values to be compared
-predictions_list1 = create_eval_lists.create_modelWL_lists(0)[0]
-actuals_list1 = create_eval_lists.create_modelWL_lists(0)[1]
+#### Running this file prints out the results of the performance evaluation of the differen models ####
 
+# Create the two lists of values to be compared, for each model a pair
+ # - 1st numeric parameter allows chosing the unit on which one evaluates the model (0 = median, 1 = mean, 2 = (median + mean)/2 )
+ # - 2nd numeric parameter should not be changed (it's just the indices of the elem in the result vector to be used)
+evaluation_lists = [(create_eval_lists.create_modelWL_lists(1)[0], create_eval_lists.create_modelWL_lists(1)[1]), 
+					(create_eval_lists.create_modelWLIP_lists(1)[0], create_eval_lists.create_modelWLIP_lists(1)[1]), 
+					(create_eval_lists.create_modelIPdur_lists(1)[0], create_eval_lists.create_modelIPdur_lists(1)[1])]
 
-predictions_list2 = create_eval_lists.create_modelWLIP_lists(2)[0]
-actuals_list2 = create_eval_lists.create_modelWLIP_lists(2)[1]
-#print(predictions_list)
-#print(actuals_list)
-
-predictions_list3 = create_eval_lists.create_modelIPdur_lists(2)[0]
-actuals_list3 = create_eval_lists.create_modelIPdur_lists(2)[1]
-
+# A name list to use when printing the results
 model_name_list = ["WL", "WL_IP", "IP_dur"]
-evaluation_lists = [(predictions_list1, actuals_list1), (predictions_list2, actuals_list2), (predictions_list3, actuals_list3)]
 
 # Evaluate models with: RMSE, MAE, and Pearson's correlation coefficient
 for i in range(len(evaluation_lists)):
@@ -24,4 +20,4 @@ for i in range(len(evaluation_lists)):
 	corrCoef = performance_measures.corrCoef(evaluation_lists[i][0], evaluation_lists[i][1])[0]
 	print("RMSE for the" + model_name_list[i] + "-model: " + str(global_rmse))
 	print("MAE for the" + model_name_list[i] + "-model: " + str(mae))
-	print("Pearson corrCoef for the " + model_name_list[i] + "-model: "  + str(corrCoef))
[…]

### 2017-03-29 py_files/model_wl.py

@@ -41,8 +41,7 @@ valid_phonemes = ["a", "a~", "e", "E", "I", "i", "O", "o", "U", "u", "Y", "y", "
 # {"diphthongs" : { 1 : [0.5, 0.62, 0.55, ...], 2 : [0.4, 0.42, 0.45, ...] ...}, 
 #  "long_vowels" : { 1 : [...], ...}, ...}
 def pcat_dict(phon_wordleng_dic):
-	dict_keys = phon_class_dict.keys()
-	pcat_dict = dict.fromkeys(phon_class_dict.keys(), defaultdict(list))
+	pcat_dict = dict( (i, defaultdict(list)) for i in phon_class_dict.keys() )
 
 	# Incorporate all phoneme proportion lists for each phoneme class:
 	#    - the phoneme keys are grouped according to the classes of the phon_class_dict
@@ -54,27 +53,24 @@ def pcat_dict(phon_wordleng_dic):
 		p_category = ([ key for key,val in phon_class_dict.items() if phoneme in val ] + ["long_vowels"])[0]
 		for w_leng in phon_wordleng_dic[phoneme].keys():
 			pcat_dict[p_category][w_leng] += phon_wordleng_dic[phoneme][w_leng][2::3]
-	
-	phon_list = [key for key, val in pcat_dict.items()]
 
 	# Reduce the proportion lists to lists containing the coresponding median, mean, and mean+median/2
 	# Looks like: 
 	#   {"diphthongs" : { 1 : [median1, mean1, (m1+m1)/2 ], 2 : [median2, mean2, (m2+m2)/2] ...}, 
 	#    "long_vowels" : { 1 : [median, mean, (m+m)/2], ...}, ...}
-	pcat_dict_reduced = dict.fromkeys(pcat_dict.keys(), defaultdict(list))
-	for categ in phon_list:
-		wleng_list = [key for key, val in pcat_dict[categ].items()]
-		for w_leng in wleng_list:
-			if len(pcat_dict[categ]) != 0:
-				pcat_dict_reduced[categ][w_leng].append(round(np.median(pcat_dict[p_category][w_leng]), 3))   # append median (model = 0)
-				pcat_dict_reduced[categ][w_leng].append(round(np.mean(pcat_dict[p_category][w_leng]), 3))     # append mean (model = 1)
[…]

### 2017-04-03 py_files/get_syllables.py
Neues model

@@ -1,0 +1,211 @@
+import re
+from collections import defaultdict
+
+diph = ["aI", "aU", "OY"]
+vowels = ["a", "A", "a~", "e", "E", "I", "i", "O", "o", "U", "u", "Y", "y", "9", "2", "@" , "6", "a:", "a~:", "e:", "E:", "i:", "o:", "u:", "y:", "2:"]
+consonants = ["z", "S", "Z", "C", "x", "N", "b", "d", "f", "g", "h", "j", "k", "l", "m", "n", "p", "r", "s", "t", "v"]
+
+# Returns the position of the syllable containing phoneme inside word:
+# - "w_initial" = 1st syllable, "w_final" = last syllable, "w_middle" = between 1st and last syll
+# - "one_s" = if it's a one syllable word
+# The counting of syllable is done on actually realized syllables! (not the canonical transcription)
+def syl_position(datei, word_no):
+	#print("Cursor: " + str(cursor))
+	syls_word = get_sylsplit_word(datei, word_no)
+	w12 = mau_split(datei, word_no)#, cursor)
+	mau12, sylsplit_word = prepare_strings(w12, syls_word)
+	syl_dict = defaultdict(list)
+	#print("MAU1: " + mau1)
+	if sylsplit_word != "":
+		syl_dict = compare_strings(mau12, sylsplit_word)
+
+
+
+		#print(syl_dict.keys())
+#		first_syl = min(syl_dict.keys())
+#		last_syl = max(syl_dict.keys())
+#		#print("Last syl: " + str(last_syl))
+#
+#		if first_syl != 0:
+#			x = 0
+#			for i in range (first_syl, last_syl+1):
+#				syl_dict[x] = syl_dict.pop(i)
+#				x += 1
+#			last_syl = max(syl_dict.keys())
+#		#print("New last syl: " + str(last_syl))
+#		#print(syl_dict)
+#		try:
+#			syl_no = get_syl(mau1.index(mau1[-1]), syl_dict)-1
+#		except:
+#			syl_no = 0
+#		else:
+#			syl_no = 0
+#		#print("Syl no: " + str(syl_no))
+#	else:
+#		first_syl, last_syl = 0, 0
+#
+#	inword_pos = ""
+#	if first_syl != last_syl:
+#		if syl_no == 0:
+#			inword_pos = "w_initial"
+#		elif syl_no >= last_syl:
+#			inword_pos = "w_final"
+##		else:
[…]

### 2017-04-03 py_files/new_SR.py

@@ -1,0 +1,187 @@
+import re
+import model_utilities
+from collections import defaultdict
+
+######################################################## STATIC DATA #############################################################
+
+# List of phonemes approximated by our models. 
+# As there are also other phonetic items segmented in our database, we use this list to sort those out.
+# Example of excluded phonetic items: <p:>, <usb>
+valid_phonemes = ["a", "a~", "e", "E", "I", "i", "O", "o", "U", "u", "Y", "y", "9", "2", "a:", "a~:", "e:", "E:", "i:",
+					"o:", "u:", "y:", "2:", "OY", "aU", "aI", "@", "6", "z", "S", "Z", "C", "x", "N", "Q", "b", "d", "f", 
+					"g", "h", "j", "k", "l", "m", "n", "p", "r", "s", "t", "v"]
+
+vowel_list = ["a", "a~", "e", "E", "I", "i", "O", "o", "U", "u", "Y", "y", "9", "2", "a:", "a~:", "e:", "E:", "i:",
+				"o:", "u:", "y:", "2:", "OY", "aU", "aI", "@", "6"]
+
+######################################################## Help Functions #############################################################
+
+# Get a list of word-numbers to split the turn into phrases (before each number)
+def get_split_list(datei):
+	work_file = open(datei)
+	split_list = []
+	word_list = []
+	a = 0
+	# Variable to store the word number of the previous line
+	memo_nr = 0
+
+	# Searching MAU-tier for pauses inside turn
+	for line in work_file:
+		if re.match("MAU", line):
+			word_list.append(int(line.split()[3]))
+			if (memo_nr == -1) and (int(line.split()[3]) > 0):
+				split_list.append(int(line.split()[3]))
+			memo_nr = int(line.split()[3])
+	
+		# Searching PRO-tier for discourse split markings
+		if re.match("PRO", line):
+			if ";" in str(line.split()[1]):
+				i = str(line.split()[1]).find(";")+1
+				split_w = int(str(line.split()[1])[i:])
+				split_list.append(split_w)
+			else:
+				split_list.append(int(line.split()[1]))
+	
+	if -1 in word_list:
+		word_list = [x for x in word_list if x != -1]
+
+	work_file.close()
[…]

### 2017-04-03 py_files/new_arff_file_writer.py

@@ -1,0 +1,95 @@
+import new_phon_dict
+import model_utilities
+import time
+
+# Generates an almost ready for processing with weka text file. 
+
+######################################################## STATIC DATA #############################################################
+# Dataset used must contain only modified par-files
+dataset_path = model_utilities.get_path_list('C:/Users/alexutza_a/Abschlussarbeit/DB_Verbmobil/mod_dataset')
+
+# List of phonemes approximated by our models. 
+# As there are also other phonetic items segmented in our database, we use this list to sort those out.
+# Example of excluded phonetic items: <p:>, <usb>
+valid_phonemes = ["a", "a~", "e", "E", "I", "i", "O", "o", "U", "u", "Y", "y", "9", "2", "a:", "a~:", "e:", "E:", "i:",
+					"o:", "u:", "y:", "2:", "OY", "aU", "aI", "@", "6", "z", "S", "Z", "C", "x", "N", "Q", "b", "d", "f", 
+					"g", "h", "j", "k", "l", "m", "n", "p", "r", "s", "t", "v"]
+
+vowel_list = ["a", "a~", "e", "E", "I", "i", "O", "o", "U", "u", "Y", "y", "9", "2", "a:", "a~:", "e:", "E:", "i:",
+				"o:", "u:", "y:", "2:", "OY", "aU", "aI", "@", "6"]
+
+ctype_dict = {
+"plosive" : ["b", "d", "p", "t", "k", "g"], "fricative" : ["f", "v", "s", "S", "z", "Z", "x", "h", "C"],
+"nasal" : ["m", "n", "N"], "lateral" : ["l"], "japproximant" : ["j"], "other" : [ "Q", "r"]
+}
+
+wtype_dict = {
+"content_words" : ["ADJA", "ADJD", "ADV", "BS", "CARD", "FM", "NN", "PTKNEG", "TRUNC", "VVFIN", "VVIMP", "VVINF", "VVIZU", "VVPP", "VAIMP", "VMPP"],
+"function_words" : ["APPR", "APPRART", "APPO", "APZR", "ART", "KOUI", "KOUS", "KON", "KOKOM", "PDS", "PDAT", "PIS", "PIAT", "PIDAT", "PPER", "PPOSS", "PPOSAT",
[…]

### 2017-04-03 py_files/new_phon_dict.py

@@ -1,0 +1,283 @@
+import re
+import model_utilities
+from collections import defaultdict
+
+######################################################## STATIC DATA #############################################################
+
+# Dataset used must contain only modified par-files
+dataset_path = model_utilities.get_path_list('C:/Users/alexutza_a/Abschlussarbeit/DB_Verbmobil/Evaluation/my_test')
+
+# List of phonemes approximated by our models. 
+# As there are also other phonetic items segmented in our database, we use this list to sort those out.
+# Example of excluded phonetic items: <p:>, <usb>
+valid_phonemes = ["a", "a~", "e", "E", "I", "i", "O", "o", "U", "u", "Y", "y", "9", "2", "a:", "a~:", "e:", "E:", "i:",
+					"o:", "u:", "y:", "2:", "OY", "aU", "aI", "@", "6", "z", "S", "Z", "C", "x", "N", "Q", "b", "d", "f", 
+					"g", "h", "j", "k", "l", "m", "n", "p", "r", "s", "t", "v"]
+
+vowel_list = ["a", "a~", "e", "E", "I", "i", "O", "o", "U", "u", "Y", "y", "9", "2", "a:", "a~:", "e:", "E:", "i:",
+				"o:", "u:", "y:", "2:", "OY", "aU", "aI", "@", "6"]
+
+ctype_dict = {
+"plosive" : ["b", "d", "p", "t", "k", "g"], "fricative" : ["f", "v", "s", "S", "z", "Z", "x", "h", "C"],
+"nasal" : ["m", "n", "N"], "lateral" : ["l"], "japproximant" : ["j"], "other" : [ "Q", "r"]
+}
+
+wtype_dict = {
+"content_words" : ["ADJA", "ADJD", "ADV", "BS", "CARD", "FM", "NN", "PTKNEG", "TRUNC", "VVFIN", "VVIMP", "VVINF", "VVIZU", "VVPP", "VAIMP", "VMPP"],
+"function_words" : ["APPR", "APPRART", "APPO", "APZR", "ART", "KOUI", "KOUS", "KON", "KOKOM", "PDS", "PDAT", "PIS", "PIAT", "PIDAT", "PPER", "PPOSS", "PPOSAT",
[…]

### 2017-04-04 py_files/new_arff_file_writer.py
Slight change of the data write code.

@@ -63,6 +63,7 @@ arff_file.write("@attribute foll_p_type { v, c, none }\n")
 arff_file.write("@attribute foll_p_artic { v, p, f, n, l, j, o, none }\n")
 arff_file.write("@attribute foll_p_voiced { v, y, n, none }\n")
 arff_file.write("@attribute syl_part_foll_p { onset, nucleus, coda, single, none }\n")
+arff_file.write("@attribute speech_rate numeric\n")
 arff_file.write("@attribute duration numeric\n%\n")
 
 ######################################################## Writing file: data #############################################################
@@ -80,11 +81,18 @@ print("Time to create dictionary: " + str(t2 - t1))
 cart_list = []
 for phon in cart_dict.keys():
 	cart_list += cart_dict[phon]
+t3 = time.time()
+print("Time to build the list for writing: " + str(t3 - t2))
+
+list_len = len(cart_list)
 
 time1 = time.time()
+print("Time to get list length: " + str(time1 - t3))
+
 # Write elements of the cart_list to the arff_file
 for el in cart_list:
-		if (cart_list.index(el) + 1) % 18 == 0:
+	for ind in range(list_len):
+		if (ind + 1) % 19 == 0:
 			arff_file.write(str(el) + "\n")
 		else:
 			arff_file.write(str(el) + ", ")

### 2017-04-04 Paper/Paper_v1.tex
Added an example db.

@@
 
-Considering machine learning, it is straightforward that one needs to split the available dataset at least into two parts: a training dataset, used for discovering patterns in data through data mining, and a test dataset, used for testing the performance of the model resulting from the machine learning process.
+Considering machine learning, it is straightforward that one needs to split the available dataset at least into two parts: a (bigger) training dataset, used for training the model, meaning discover patterns in data, and a (smaller) test dataset, used for testing the performance of the model resulting from the machine learning process.
 
@@
 \chapter{Corpora}
-Segmented and annotated German speech corpora represent sets of text files, so-called ``textual data``, containing information about the speech data.
-Our minimum requirement for such corpora was that they contain speech segmented at phoneme level, allowing us to calculate the duration of each phoneme occurence.
+Language related databases are called corpora, and for our purpose we needed one containing segmented and annotated German speech as a set of text files, so-called ``textual data``.
+Our minimum requirement for such a corpus was that it contains speech segmented at phoneme level, allowing us to calculate the duration of each phoneme occurence.
 The unit in which this duration is expressed is not relevant.
+As phoneme duration may be influenced by several factors, and we did not intent to do any further processing of the audio files, we also needed that this textual database contains enough information to identify following attributes:
 
-For further processing of this information from our side, we considered the study of Rosen \cite{Rosen2005} on the application of the lognormal distribution for modeling the variation of speech segment duration.
-In accordance to this, during the data mining tasks we opted for a log-transformation of the observed phoneme durations using the natural logarithm, which proved to be a better description of the data in our database as well (fig.
-\ref{fig:aI_distrib}).
+\begin{itemize}
+	\item phrase boundaries
+	\item word boundaries
+	\item syllable boundaries
+	\item phoneme identification
+	\item vowel stress
+	\item pauses
+\end{itemize}
 
-\begin{figure}[htbp]
-\centering
-	\begin{subfigure}[h]{0.45\textwidth}
-	\centering
-	\includegraphics[width=\textwidth]{../Graphen/Hist_aims_no_title.eps}
-	\caption{Duration of /aI/ measured in seconds.}
-	\label{fig:ai_ms}
-	\end{subfigure}
-\hfill
-	\begin{subfigure}[h]{0.45\textwidth}
-	\centering
-	\includegraphics[width=\textwidth]{../Graphen/Hist_ailn_no_title.eps}
-	\caption{Duration of /aI/ measured in ln(sec).}
-	\label{fig:ai_ln}
-	\end{subfigure}
-\caption[Phoneme duration distribution ms vs ln(ms)]{Phoneme durations measured in seconds have a not normal distribution shape.
-However, when applying a logarithmic transformation in terms of ln(msec), the distribution is better shaped.}
-\label{fig:aI_distrib}
-\end{figure}
+There are three corpora which we had at our disposal: (1) The Kiel Corpus of Read (PhonDat) and of Spontaneous Speech, (2) Verbmobil, and (3) spoken Wikipedia.
+The first one contains two equal sized corpora of > 4 hours each: KCoRS, which is a collection of read speech, and KCoSS, which is a collection of spontaneous speech, both being manually segmented, which is a good indicator for good segmentation quality.
+However, as Pfitzinger \cite{Pfitzinger} also showed in his reassessment of the PhonDat-II corpus, for nowadays research the Kiel Corpus is rather small.
+Spoken Wikipedia is a new project comprising of a much larger collection of read speech.
+Unfortunately, the phoneme-level segmentation was not ready on time to evaluate in our project.
 
@@
 	\caption[Phoneme length variation between speakers]{Overview of \texttt{/a:/} length on the 5 most frequent speakers.
-Phoneme length is expressed as the natural logarithmus of the sample frequency.
+Phoneme length is expressed as the natural logarithm of the sample frequency.
 While the minimum duration is technically limited to an equal value corresponding to 0.04 sec, all other coordinates vary, e.g.

### 2017-04-04 py_files/new_phon_dict.py
Corrected mistake in line 275.

@@ -1,5 +1,6 @@
 import re
 import model_utilities
+import new_SR
 from collections import defaultdict
 
 ######################################################## STATIC DATA #############################################################
@@ -266,6 +267,9 @@ def cart_dict(dataset_path):
 						p_dict[p_key].append(get_syl_part(datei, following_phoneme[1], following_phoneme[0], following_phoneme[2]))
 					else:
 						p_dict[p_key].append("none")
+
+					# Speech rate calculated as word_duration / phoneme_count_in_word
+					p_dict[p_key].append(new_SR.get_word_SR(datei, int(line.split()[3]))[0])
 
 					# Duration of current phoneme in msec
 					p_dict[p_key].append(round(int(line.split()[2]) * 0.0625, 2))

### 2017-04-04 py_files/model_utilities.py
Set back some code lines

@@ -90,7 +90,7 @@ def word_statistics(datei, word_no):
 			if str(line.split()[4]) in vowel_list:
 				mau_syl_count += 1
 	
-	#word_duration *= 0.0000625
+	word_duration *= 0.0625
 	work_file.close()
 
 	return word_duration, phoneme_count, mau_syl_count

### 2017-04-04 py_files/new_SR.py

@@ -185,3 +185,5 @@ def get_word_SR(datei, word_nr):
 	n_phon_WSR = round(phon_count / word_dur, 2)
 
 	return p_phon_WSR, n_phon_WSR
+
+#print(get_word_SR('C:/Users/alexutza_a/Abschlussarbeit/DB_Verbmobil/mod_dataset/g001acn1_000_AAJ_a.par', 1)[0])

### 2017-04-04 py_files/new_arff_file_writer.py

@@ -6,7 +6,7 @@ import time
 
 ######################################################## STATIC DATA #############################################################
 # Dataset used must contain only modified par-files
-dataset_path = model_utilities.get_path_list('C:/Users/alexutza_a/Abschlussarbeit/DB_Verbmobil/mod_dataset')
+dataset_path = model_utilities.get_path_list('C:/Users/alexutza_a/Abschlussarbeit/DB_Verbmobil/Evaluation/Test_a')
 
 # List of phonemes approximated by our models. 
 # As there are also other phonetic items segmented in our database, we use this list to sort those out.
@@ -33,7 +33,7 @@ wtype_dict = {
 
 ######################################################## Writing file: attribute description #############################################################
 
-arff_file = open("C:/Users/alexutza_a/Abschlussarbeit/wekadateien/cart_58db.txt", "w")
+arff_file = open("C:/Users/alexutza_a/Abschlussarbeit/wekadateien/cart_test.txt", "w")
 
 arff_file.write("% ARFF file for Verbmobil\n% \n@relation cart\n%\n% List of attributes:\n")
 arff_file.write("% Duration is expressed in miliseconds\n%\n")
@@ -84,20 +84,23 @@ for phon in cart_dict.keys():
 t3 = time.time()
 print("Time to build the list for writing: " + str(t3 - t2))
 
-list_len = len(cart_list)
+i = 0
 
 time1 = time.time()
 print("Time to get list length: " + str(time1 - t3))
 
 # Write elements of the cart_list to the arff_file
 for el in cart_list:
-	for ind in range(list_len):
-		if (ind + 1) % 19 == 0:
-			arff_file.write(str(el) + "\n")
-		else:
-			arff_file.write(str(el) + ", ")
+	if  (cart_list.index(el) + 1):
+		arff_file.write(str(el) + "\n")
+		i += 1
+	else:
+		arff_file.write(str(el) + ", ")
+		i += 1
+
 
 arff_file.close()
 
 time2 = time.time()
 print("Time to write the data: " + str(time2-time1))
+#

### 2017-04-04 py_files/new_phon_dict.py

@@ -90,7 +90,7 @@ def get_syl_part(datei, phoneme, word_no, syl_no):
 		nucleus_ind = "NaN"
 
 	# Get the index of current phoneme in syl_list
-	p_index = syl_list.index([x for x in syl_list if x == phoneme][0])
+	p_index = syl_list.index(phoneme)
 
 	if type(nucleus_ind) == int:
 		if p_index < nucleus_ind:

### 2017-04-04 py_files/new_arff_file_writer.py
eliminated extra var

@@ -84,8 +84,6 @@ for phon in cart_dict.keys():
 t3 = time.time()
 print("Time to build the list for writing: " + str(t3 - t2))
 
-i = 0
-
 time1 = time.time()
 print("Time to get list length: " + str(time1 - t3))
 
@@ -93,10 +91,8 @@ print("Time to get list length: " + str(time1 - t3))
 for el in cart_list:
 	if  (cart_list.index(el) + 1):
 		arff_file.write(str(el) + "\n")
-		i += 1
 	else:
 		arff_file.write(str(el) + ", ")
-		i += 1
 
 
 arff_file.close()

### 2017-04-05 Paper/Paper_v1.tex
Some things on speech rate.

@@
 \bibliography{../PaperBib}
+\usepackage{amsmath}
 \usepackage{graphicx}
@@
 The first one contains two equal sized corpora of > 4 hours each: KCoRS, which is a collection of read speech, and KCoSS, which is a collection of spontaneous speech, both being manually segmented, which is a good indicator for good segmentation quality.
-However, as Pfitzinger \cite{Pfitzinger} also showed in his reassessment of the PhonDat-II corpus, for nowadays research the Kiel Corpus is rather small.
+However, as Pfitzinger \cite{Pfitzinger2002} also showed in his reassessment of the PhonDat-II corpus, for nowadays research the Kiel Corpus is rather small.
 Spoken Wikipedia is a new project comprising of a much larger collection of read speech.
@@
 phonetic transcription, segmentation, labeling, etc.) is captured in so-called ``par files``, representing files in the ``BAS Partitur Format`` \cite{Burger2000}.
+Our dataset included 604355 automatically labeled phoneme instances.
 
@@
 Using the classical phonetic conventions, we can group these phonemes on a first level according to their phonetic features using a tree structure into vowels, consonants and a glottal stop, although the latter one is sometimes treated as a consonant \cite{Kohler1995}.
+Vowels represent 24\% of the total phoneme inventory analyzed.
 Further relevant groupings are being examined under sections ``Vowels`` and ``Consonants``.
@@
 
-We selected ... because...
+As we recreated the tree model of Brinckmann \& Trouvain \cite{Brinckmann_2003} for the most part, we also selected most of the features used by them. Here just a rough listing of these factors:
+
+\begin{itemize}
+	\item phoneme id
+	\item vowel and syllable stress
+	\item manner of articulation for consonants
+	\item position of phoneme in syllable
+	\item position of syllable in word
+	\item position of word in turn
+	\item part-of-speech: function word, content word, proper name, interjection-or-answer-particle
+	\item speech rate
+\end{itemize}
+
+A detailed description of the model implementation and phoneme features used follows in Chapter 6.
 
@@
 \chapter{Defining speech rate}
-Under speech rate one always tends to understand a fraction of something divided by total time ocupied by that something.
+Under speech rate or speech tempo we understand a measure of how fast someone speaks.
+This may be seen as a discrete variable, with discrete values as \textit{slow}, \textit{normal}, and \textit{fast} or as a continuous one, with continuous numerical values.
+We opted for the continuous version, as it is better suited for our purpose.
+
+Expressed mathematically, this would be a fraction of something divided by total time ocupied by that something.
 The German word ``Sprechgeschwindigkeit`` involves the concept of ``speed`` in the equation, which would be space divided by time.
 In the case of speech this would be number of units divided by total time ocupied by those units.
-However, this approach doesn't consider the contribution of the single phoneme durations to the whole segment, but rather supposes an equal length for all phonemes in the numerator.
+This is fine, if one has a fixed time interval and is interested in the number of units that may fit in.
+However, this way of looking at it does not consider the fact, that the length of the unit may vary.
+We are not interested in variing the number of units, which should actually stay constant.
+We are rather interested in variing the time interval while keeping the number of units constant.
+This is what happends when someone is speaking faster or slower than a specific norm.
+Therefore we inverted the typical speed-like calculating formula, as you can see in Eq.
+\ref{eq1}, so we can investigate the length variation in the speech units, with focus on units = phonemes.
 
-Speech rate may be seen as a discrete or as a continuous variable. We opted for the continuous version, as it is better suited for our purpose.
-Speech rate may be calculated as global, local or relative speech rate.
+\begin{equation}
+\label{eq1}
+	SR = \frac{segment \quad duration}{unit\quad count}
+\end{equation}
 
-Pfitzinger \cite{Pfitzinger1998} showed that syllable count per second is a better approximation of the real speech rate than calculating the number of phones per second.
-He also proposed a formula combining the two rates, which should further improve this result for the calculation of the local speech rate, which was confirmed in later studies \cite{Pfitzinger1999}.
+The next thing to think about is the size and identity of the fraction elements.
+Unit size may be phoneme, syllable or word.
+Variing the scope from local to globa, the segment on which we calculate speech rate may be a syllable, a word, a phrase, a speech turn or the entire corpus.
+In the literature \cite{Pfitzinger1998} we also found mentions about a relative speech rate, which combines the speech rate calculated to two different unit sizes.
 
-(We decided to calculate global speech rate as a ratio of number of syllables per second.
-As syllables were not annotated in our database, we decided to use the number of (realized?) vowels as an approximation for the number of syllables \cite{Yishan_Jiao_2015} \cite{Kohler1995}.)
+From the beginning, we did not consider the variant of calculating a speech rate for the entire corpus.
+Given its structure of dialogues recorded by many speakers, it is straightfoward that speech rate may vary tremendousely not only between speakers, but also between different speech turns, and consequently be a bad measure for characterizing our corpus.
+The size of a speech turn in a dialog proved to be a bad measure as well.
+The reason for this is again the nature of spontaneous speech as ``live`` creation of the human mind.
+Hesitations, thinking, and emotions have a much greater influence as in read speech, and result in a variing tempo at which we produce speech.
+In figure \ref{fig:SR_example} we are illustrating an example of this phenomenon.
 
-In our study we avoided expressing speech rate neither as a fraction, nor as a number.
-We took the segment duration (i.e.
-denominator) alone, varied it, and tried to figure out what happends in the numerator so that the proportion stays.
-In other words, what we do is to say that when the speech rate changes, the duration of changes as well.
-Now given one has an algorithm to predict the duration change of single words in a phrase when speech rate changes, we only care about phoneme duration change inside word of given durations.
+\begin{figure}[htbp]
+	\centering
+	\noindent\makebox[\textwidth]{\includegraphics[width=\paperwidth, clip]{../Graphen/SR_example.eps}}
+	\vspace{-25mm}
+	\caption[Example of speech rate variation inside a turn]{This is the content of the speech turn number 22 of speaker AAJ in dialoge g001.
+The closer the points are to oneanother, the faster he speaks.
+Orange lines mark prosodic boundaries.}
+	\label{fig:SR_example}
+\end{figure}
+
+Seing this, there is no wonder that we got no good correlation values as long as we calculated speech rate on entire turns.
+This got better when we reduced the size of the segment to phrase, and then further down to words.
+An additional measure we used to evaluate how well a speech rate calculating formula may suit our purpose was the correlation value between speech rate and phoneme duration.
+As long vowels proved to be the most sensible to the variation of the speaking tempo, we present below the values we obtained for this group, together with the speech rate calculated at different segment sizes.
+
+%\begin{table}[htbp]
+%\caption {Duration of long vowels vs. speech rate calculated at turn, phrase, and word level} 
+%\centering
+\vspace{10mm}
+\begin{tabular}{|l|l|}
+\hline
+Segment	 & Correlation coefficient (Pearson)\\
+\hline
+\hline
+Turn 		&0.727178570436 \\
+	\hline
+Phrase  &\\
+	\hline
+Word  &  0.794901825792\\
+	\hline
+\end{tabular}
+%\end{table}
+
+%\end{table}
 

### 2017-04-05 Paper/Paper_v1.tex
some more changes to sr
some adds to chapter basics

@@
 
-\chapter{Fundamentals}
+\chapter{Basics}
 \section{Data Mining}
@@
 Computers help process within resonable time amounts of data that would need years of manual processing.
+
 \section{Machine Learning}
 Machine Learning means building a program or model which can make predictions about an unseen dataset, like filling in missing information, based on data mining results from a given dataset.
+It consists of a so-called training phase, in which the training data is being processed by the selected machine learning method and hypotheses or models are formed.
+These can then be used to make predictions about new data.
+The training data consists of a set of instances, each of them being described by a feature vector.
+The training phase is followed by an evaluation phase, in which predictions about unseen data are being made and compared to the actual values.
+
+We used machine learning to solve a task involving both classification and numeric prediction.
+We used for this two model tree methods: M5P and REPTree.
+The reasons we chose them is because trees are easy to interpret, they can perform both classification and numeric prediction tasks, we could easily generate them using Weka, and are the closest to classification and regression (CART) trees, also used in the literature \cite{Brinckmann2003} for phoneme duration prediction, so we have data to compare with.
+
+\subsection{REPTree}
+
+\subsection{M5P}
+The M5P model tree is a decision tree based on the M5 algorithm introduced by Quinlan (1992). It
+
 \section{Preparing the Data}
@@
 \end{enumerate}
+
+\section{Tools}
 
@@
 
-%\begin{table}[htbp]
-%\caption {Duration of long vowels vs. speech rate calculated at turn, phrase, and word level} 
+\begin{table}[htbp]
+\caption {Duration of long vowels vs. speech rate calculated at turn, phrase, and word level} 
 %\centering
 \vspace{10mm}
-\begin{tabular}{|l|l|}
+\begin{tabular}{|l|l|c|}
 \hline
-Segment	 & Correlation coefficient (Pearson)\\
+Segment	 & Formula & Correlation coefficient (Pearson)\\
 \hline
 \hline
-Turn 		&0.727178570436 \\
+Turn 		& $turn\_dur / phon\_count$ &0.727178570436 \\
 	\hline
-Phrase  &\\
+	&  $phon\_count / turn\_dur$ & -0.270392865503 \\
 	\hline
-Word  &  0.794901825792\\
+Phrase  & $phrase\_dur / phon\_count$ & 0.561266362391\\
+	\hline
+	&  $phon\_count / phrase\_dur$ & -0.377058296098\\
+	\hline
+Word  & $word\_dur / phon\_count$ &  0.794901825792\\
+	\hline
+	&  $phon\_count / phrase\_dur$ & -0.616539202577\\
 	\hline
 \end{tabular}
-%\end{table}
+\end{table}
 
-%\end{table}
+\begin{figure}[htbp]
+\centering
+	\begin{subfigure}[h]{0.48\textwidth}
+	\centering
+	\includegraphics[width=\textwidth]{../Graphen/SR_turn.eps}
+	\label{fig:tsr}
+	\end{subfigure}
+\hfill
+	\begin{subfigure}[h]{0.48\textwidth}
+	\centering
+	\includegraphics[width=\textwidth]{../Graphen/SR0.eps}
+	\label{fig:phraseSR}
+	\end{subfigure}
+\hfill
+	\begin{subfigure}[h]{0.48\textwidth}
+	\centering
+	\includegraphics[width=\textwidth]{../Graphen/SR4.eps}
+	\label{fig:wsr}
+	\end{subfigure}
+\caption[Speech rate vs long vowel duration]{Phoneme durations measured in msec plotted against different speech rates.}
+\label{fig:SR_corr}
+\end{figure}
 

### 2017-04-05 Paper/Paper_v1.tex
pictures

@@
 \subsection{M5P}
-The M5P model tree is a decision tree based on the M5 algorithm introduced by Quinlan (1992). It
+The M5P model tree \cite{Witten2011} is a decision tree based on the M5 algorithm introduced by Quinlan (1992).
+It builds a usual classification tree out of the attributes, and retrieves a linear regression model in each leaf out of the attributes contained in that specific subtree.
 

### 2017-04-05 Paper/Paper_v1.tex
Added some more things to Basics.

@@
 \chapter{Basics}
-\section{Data Mining}
-Data Mining means in foreground looking for patterns in data with the help of computers.
-Patterns are helpful for humans to understand data.
-They reduce the amount of data, and of available features to the relevant ones.
-Computers help process within resonable time amounts of data that would need years of manual processing.
 
@@
 
+Data Mining means in foreground looking for patterns in data with the help of computers.
+Patterns are helpful for humans to understand data.
+They reduce the amount of data, and of available features to the relevant ones.
+Computers help process within resonable time amounts of data that would need years of manual processing.
+
 We used machine learning to solve a task involving both classification and numeric prediction.
@@
 
-\subsection{REPTree}
+\subsection{Decision trees}
+Decision trees divide the attribute space into well defined clusters, identified by a class name or a numeric value.
+In the first case we are dealing with a classification tree, in the second one with a regression tree.
+If one needs to combine the two methods, he/she may use a classification and regression tree, called CART, or a model tree.
+The advantage of trees is that new instances are easily assigned to a class, or a numeric value, or a linear model which leads to such a numeric value, and they are easy to unterstand.
 
-\subsection{M5P}
-The M5P model tree \cite{Witten2011} is a decision tree based on the M5 algorithm introduced by Quinlan (1992).
+The \textbf{M5P model tree} \cite{Witten2011} is a decision tree based on the M5 algorithm introduced by Quinlan (1992).
 It builds a usual classification tree out of the attributes, and retrieves a linear regression model in each leaf out of the attributes contained in that specific subtree.
+Branching is realised based on the reduction of variation within the resulting subset.
+
+A \textbf{REPTree} is a decision and regression tree similar to M5P, but optimized for speed. This means it sorts values for numeric attributes only once.
+
+Both trees
 
@@
 \section{Performance Evaluation}
-Models produced by machine learning may have different performance levels, which may be evaluated in several ways. We decided for the following three:
+Models produced by machine learning may have different performance levels, which may be evaluated in several ways.
+As they assess different features of the models, and treat outliers differently, we didn't want to rely only on one of them.
+Furthermore, we wanted to be able to compare the results with results presented in the literature.
+After studying the common evaluation methods used for other phoneme prediction models, we chose the following three:
 \begin{enumerate}
@@
 It is computed using the average of the absolute errors: $MAE = \frac{\sum\left|predicted-actual\right|}{n}$
-	\item The \textbf{correlation coefficient} is a number between -1 and 1 that quantifies the correlation between two variables, each having a separate value set.
+	\item The \textbf{Pearson correlation coefficient}, a popular measure for continuous data, is a number between -1 and 1 that quantifies the correlation between two variables, each having a separate value set.
 We can interpret this as how much the variable movement through its value set corresponds to the movement of the other variable in its own value set.
@@
 \section{Tools}
+We used \textbf{Weka} 3.8 for building and training the model trees for the phoneme duration prediction.
+It is an open source freeware using a Java VM, and contains a collection of tools for all machine learning tasks we needed, from data pre-processing over classification and regression till model evaluation and data visualization.
+
+In order to extract the needed information from our database, and put it in a form ready to use with Weka, we used self-made code written in Python 3.6.
+This is an easy to learn programming language with many features dedicated to statistical analysis and even machine learning.
+Most of the figures in this paper were produced in python as well.
 

### 2017-04-05 Paper/Paper_v1.tex
Started adding / modifying the part on method

@@
 Phoneme length adjustment is the aspect we wish to adress in this work.
-In particular, we want to find an algorithm to adjust the phoneme duration according to the speech rate.
+In particular, we want to find an algorithm for adjusting phoneme durations according to the variations in speech rate.
 The resulting phoneme duration should be close to the actual duration that occurs when the speaking rate alters.
@@
 There are several models for phoneme duration approximation, that work pretty well for some languages, German inclusive.
-However, they consistently don't take the speech rate into consideration, their performance level being acquired within a specific speech rate.
+However, they consistently don't take the speech rate into consideration, their performance level being acquired rather within a specific speech rate.
 When this varies, model performance deteriorates, like van Santen states for his model \cite{Santen1994}.
@@
 
-The purpose is therefore not to find another duration approximation model for TTS-systems, but to improve the algorithm of phoneme duration adjustment, given known reference phoneme and word durations, and a target word duration, as presented in Table \ref{tab:Question}.
-Consequently, the modified speech unit would sound more natural at a faster or a slower speech rate than being modified with a linear function.
+The purpose is therefore not to find another duration approximation model for TTS-systems, but to improve the algorithm of phoneme duration prediction by considering the speech tempo.
+Table \ref{tab:Question} vizualises our challenge: we want to put all available/visible information from the table in relation so that we can replace the question marks with values that are close to the real ones.
+Consequently, the modified speech unit would sound more natural at a faster or a slower speech rate as when being modified with a linear function.
 
@@
 \end{tabular}
-\caption*{The first column shows our inputs: duration of each phoneme in German word ``genau``, given a speech rate of 0.07 sec, or a word duration of 0.28 sec.
-The second column presents our challenge: how long is the duration of each phoneme of the given word, when it is spoken slower, taking e.g.
-0.42 sec, speech rate of approx.
-0.11 sec?}
+\caption{The first column shows our inputs: duration of each phoneme in German word ``genau``, given a speech rate of 0.07 sec, or a word duration of 0.28 sec.
+The second column contains our challenge: How long is the duration of each phoneme of the given word, at a speech rate of 0.11 sec, equivalent of a word length of 0.42 s ?}
 \label{tab:Question} 
@@
 
-Both trees
+Both trees cope well with missing data. However, both need a very large training set in order to perform well.
 
@@
 0.86.
+The drawback of this model is its intransparency.
 \end{itemize}
@@
 \section{Method}
-Whichever phoneme duration prediction model one takes, and whichever size of the segment one takes in the denominator, in terms of syllable, word, foot or phrase duration...
-Larger the segment size, larger the number of influencing factors for phoneme duration.
-We opted for a progressive approach, starting at the syllable level ?.
+Our starting premise was that phoneme durations are being influenced by the speaking tempo.
+In addition to this, we assumed that this influence is not equal on all phonemes, and on all situations, and therefore also not linear.
+So we formulated our aim as to investigate the influence of speech rate on the duration of different phonemes and find a way of using this information to improve the performance of phoneme prediction models.
 
-Our approach is based on the observation that specific phonemes, respectively phoneme groups, occupy a specific percent of the total word duration (fig.
-\ref{fig:boxplot_genau}), and that certain phonemes seem to modify their duration according to that of the segment they're in to a much higher degree than others, as you can see in fig.
-\ref{fig:w_dur_vs_pho}.
+In order to achieve this aim, we first examined the phonemes with respect to their duration.
+We found out that the classical phonetics book grouping of phonemes is not respected in terms of duration.
+There are consonants like \texttt{/N/}, and \texttt{/m/} which may be longer than some long vowels, and long vowels which are shorter than some short vowels or even several consonants, like \texttt{/y:/}.
+If one would apply a clustering method on phonemes to group them on seven classes based on their duration interval coordinates using, say k-means method, the resulting groups would certainly differ from those listed in Chapter 4.
+Fig?!
 
-\begin{figure}[htbp]
-	\includegraphics[width=\textwidth]{../Graphen/Stats_for_genau.eps}
-	\centering
-	\caption[Word duration of ``genau`` vs.
-its phoneme durations]{We ordered ascendingly the 706 occurences of the word ``genau`` based on word duration and ploted recorded durations of the component phonemes against it.
-The picture suggests clearly that the diphthong \texttt{/aU/} occupies a rather fixed steak of the total word duration, while the others show no adjustment to the increasing word total duration.
-}
-	\label{fig:w_dur_vs_pho}
-\end{figure}
+Therefore we decided to capture phoneme identity as well in our model, although it inflates the number of attributes and attribute values to compute.
 
-\begin{figure}[htbp]
-	\includegraphics[width=.55\linewidth]{../Graphen/Sound_boxplot_genau.eps}
-	\centering
-	\caption[Phoneme proportions in ``genau``]{Proportions taken by the component phonemes of ``genau`` suggest the possibility of grouping them in disjunct classes.
-Here we explored the phoneme distribution for a relatively short occurence of our reference word, of only 0.3 sec found 25 times.
-This one, as well as others, correlates well with the overall phoneme distribution on all occurences.}
-	\label{fig:boxplot_genau}
-\end{figure}
+Then we looked for a definition of speech rate that can be used to model phoneme durations.
+This supposed the presence of a good correlation between speech rate and phoneme duration, at least for some phoneme groups.
+We explained our choice whith respect to this in Chapter 5.
+
+\section{Experiment}
+
+\section{Results}
 

### 2017-04-06 Paper/Paper_v1.tex
Added method and experiment

@@
 
-\chapter{Defining speech rate}
-Under speech rate or speech tempo we understand a measure of how fast someone speaks.
-This may be seen as a discrete variable, with discrete values as \textit{slow}, \textit{normal}, and \textit{fast} or as a continuous one, with continuous numerical values.
-We opted for the continuous version, as it is better suited for our purpose.
+\chapter{Speech rate}
+Speech rate is a measurement of the ``tempo`` of speech, that is how quickly someone speaks.
+This may be seen as a discrete/qualitative measure, with discrete values as \textit{slow}, \textit{normal}, and \textit{fast} or as a continuous one, with continuous numerical values.
+We opted for the continuous version, because it is quantifyable and so can be a useful input into a regression, see Section 2.1.1.
 
-Expressed mathematically, this would be a fraction of something divided by total time ocupied by that something.
-The German word ``Sprechgeschwindigkeit`` involves the concept of ``speed`` in the equation, which would be space divided by time.
-In the case of speech this would be number of units divided by total time ocupied by those units.
+Expressed mathematically, speech rate can be defined similarly to speed in spatial displacement, which would be distance covered by time.
+In the case of speech, distance is not counted in the spatial domain but in relevant linguistic units.
+Other than in the case of distance, the length of the speech units is not standardized (like kilometers, miles, etc.).
+Moreover it's size is also expressed in units of time: phoneme \texttt{/a/} has the length of x seconds, miliseconds, etc.
+
 This is fine, if one has a fixed time interval and is interested in the number of units that may fit in.
@@
 We are rather interested in variing the time interval while keeping the number of units constant.
-This is what happends when someone is speaking faster or slower than a specific norm.
+This is what happens when someone is speaking faster or slower than a specific norm.
 Therefore we inverted the typical speed-like calculating formula, as you can see in Eq.
@@
 \label{eq1}
-	SR = \frac{segment \quad duration}{unit\quad count}
+	SR = \frac{word \quad duration}{phoneme\quad count}
 \end{equation}
@@
 	\centering
-	\begin{subfigure}[h]{0.48\textwidth}
-	\centering
-	\includegraphics[width=\textwidth]{../Graphen/SR_turn.eps}
-	\label{fig:tsr}
-	\end{subfigure}
-\hfill
-	\begin{subfigure}[h]{0.48\textwidth}
-	\centering
-	\includegraphics[width=\textwidth]{../Graphen/SR0.eps}
-	\label{fig:phraseSR}
-	\end{subfigure}
-\hfill
-	\begin{subfigure}[h]{0.48\textwidth}
-	\centering
 	\includegraphics[width=\textwidth]{../Graphen/SR4.eps}
+	\label{fig:}
+\caption[Speech rate vs long vowel duration]{Phoneme durations measured in msec plotted against speech rates.}
 \label{fig:wsr}
-	\end{subfigure}
-\caption[Speech rate vs long vowel duration]{Phoneme durations measured in msec plotted against different speech rates.}
-\label{fig:SR_corr}
 \end{figure}
@@
 
-Therefore we decided to capture phoneme identity as well in our model, although it inflates the number of attributes and attribute values to compute.
+Therefore we decided to capture phoneme identity in our model as well, although it inflates the number of attributes and attribute values to compute.
+The relation between phoneme duration and factors other than speech rate was treated only marginally, as these have already been extensively examined in the literature on various corpora.
+At this stage, we considered such factors only to the extent they might have helped identify groups of phonemes that may be best suited to evaluate the influence of speech rate.
+Such a factor is primary stress, which creates a group of very elastic vowels, which are therefore also sensitive to changes in speech tempo.
 
@@
 We explained our choice whith respect to this in Chapter 5.
+The fact, that speech rate doesn't show any correlation with some phonemes, like \texttt{/r/}, while it has a pretty good correlation with others, like \texttt{/a:/} suggests the need for classification in whichever model we choose.
 
 \section{Experiment}
+In order to test whether our definition of speech rate has an influence on the prediction of the exact phoneme duration by comparing a duration prediction model with and without a feature for speech rate.
+We chose a tree model, because it is simple to implement, and to interpret.
+One can easily examine this kind of model to see which features are being used, and which weights they bear.
+
+We first reconstructed to a great extent a previous experiment \cite{Brinckmann2003} that used a CART tree to predict phoneme duration, and reached good approximation values, because we wanted to have a reference for comparison.
+There is only one feature out of the 19 used in that experiment, which we could not recreate: degree of word accentuation.
+We lacked prosodic information for doing this.
+However, we tried to capture this kind of information by adding two extra word type classes: emphatic function words, like interjections and answer particles, and emphatic content words, like proper names and spelled items.
+
+Next, we pruned the used feature space by using two feature selection algorithms: forward and backward feature selection. Consequently, following features were selected:
+
+\begin{itemize}
+	\item PHONEME INSTRINSIC FEATURES
+
+	\begin{itemize}
+		\item phoneme id: all the 52 phoneme definitions used in Verbmobil
+		\item phoneme type: vowel, or consonant
+		\item articulation manner: vowel, plosive, fricative, nasal, lateral, approximant, other
+	\end{itemize}
+
+	\item SYLLABLE RELATED FEATURES
+
+	\begin{itemize}
+		\item position of phoneme in syllable: initial, final
+		\item syllable part containing current phoneme: onset, nucleus, code, single-phoneme-syllable
+		\item lexical stress of the syllable (using the stress type of the nucleus vowel)
+	\end{itemize}
+
+	\item PHONEMIC CONTEXT
+
+	\begin{itemize}
+		\item previous phoneme type: vowel, consonant, or none
+		\item following phoneme type: vowel, or consonant, or none
+		\item articulation manner of following phoneme: vowel, plosive, fricative, nasal, lateral, approximant, other, none
+		\item if following phoneme is voiced: yes, no, or none
+		\item syllable part to which following phoneme belongs: onset, nucleus, code, single-phoneme-syllable
+	\end{itemize}
+\end{itemize}
+
+Note that all features consider only actually realised units in speech. 
+After pruning, the \textit{word} related features got completely eliminated, together with \textit{syllable-place-in-word} and \textit{articulation-of-realised-previous phoneme}.
 

### 2017-04-06 Paper/Paper_v1.tex
Added my approach (section 6)

@@
 \section{Experiment}
-In order to test whether our definition of speech rate has an influence on the prediction of the exact phoneme duration by comparing a duration prediction model with and without a feature for speech rate.
+In order to test whether adding speech rate information really improves the predictions of a phoneme duration model, we tested the same model with and without the speech rate feature.
 We chose a tree model, because it is simple to implement, and to interpret.
@@
 
-We first reconstructed to a great extent a previous experiment \cite{Brinckmann2003} that used a CART tree to predict phoneme duration, and reached good approximation values, because we wanted to have a reference for comparison.
+For the sake of comparison, and to see whether we lie on the right track, we created at the beginning two basic models:
+
+\begin{enumerate}
+	\item The first one predicted the phoneme duration as the mean of its durations across the whole corpus. 
+	\item The second one build a two level tree, which split the root on the 49 realized phonemes, and then each phoneme on two specific speech rate values.
+The leaves contained three predicted values for each phoneme.
+These values were calculated using the mean duration of the specific phoneme across the corpus modified or not by a proportion of its standard deviation, based on the speech rate value split.
+Figure ...
+shows the structure of this tree.
+\end{enumerate}
+
+As the results of the initial two models were encouraging, we proceeded with a ``real`` model.
+For this, we reconstructed to a great extent a previous experiment \cite{Brinckman2003} that used a CART tree to predict phoneme duration, and reached good approximation values.
+By using this model we also had the advantage of having a reference for comparison.
 There is only one feature out of the 19 used in that experiment, which we could not recreate: degree of word accentuation.
@@
 However, we tried to capture this kind of information by adding two extra word type classes: emphatic function words, like interjections and answer particles, and emphatic content words, like proper names and spelled items.
+For such words it is generally known that they usually carry phrasal stress.
 
-Next, we pruned the used feature space by using two feature selection algorithms: forward and backward feature selection. Consequently, following features were selected:
+Next, we pruned the used feature space by using two feature selection algorithms: forward and backward feature selection.
+Their principle is easy to understand and they are available in Weka.
+Consequently, following features were selected:
 
@@
 Note that all features consider only actually realised units in speech.
+We also had to reduce the amount of data used for the model to 58\% of the total corpus, due to capacity limitations imposed by Weka on our system.
 After pruning, the \textit{word} related features got completely eliminated, together with \textit{syllable-place-in-word} and \textit{articulation-of-realised-previous phoneme}.
 
+To this collection of features we added \textit{speech rate} in the second phase and compare the results.
+
 \section{Results}
+We trained and evaluated both a M5P and a REPTree model.
+The results were very similar, so present here only the results obtained using the model tree M5P, as they were slightly better.
+As can be seen in Table ...
+adding the speech rate feature caused a significant improvement of all our models.
+
+Table
+
+However, the performance of our model tree is smaller than the performance of the CART built by Brinckmann \& Trouvain \cite{Brinckman2003}.
+This is explainable if we consider the nature of the used datasets.
+The CART tree used PhonDat, a rather homogenous dataset in terms of speech rate and number of speakers.
+Furthermore, the model was trained and tested separately on data produced by one speaker.
+The reported results show even for PhonDat a difference in performance between speakers.
+Another aspect that can have a great impact on the model result is the quality of the input data.
+While PhonDat was manually segmented and annotated, Verbmobil was processed automatically using the MAUS segmentation system.
+When evaluating outliers we found many examples of wrong segmentation, as we also discussed in Section 4.2.
+
+On the other hand, our model tree shows robustness with respect to factors like inter-speaker variability and speech rate.
+The results obtained for a test set containing data from a speaker not included in the training data were similar to those obtained on known speakers on terms of correlation coefficient, and showed less error rates, as can be seen in table ....
+Moreover, if the model is to be evaluated on data produced by only one speaker, it produces good results even when using less than 10\% of the amount of data used to train on the multiple speaker corpus.
+The most frequent speaker in our database still only has 185 recorded turns, consisting of 9315 identifyiable uttered phonemes, which represent 0.02\% of the total phoneme instances in our corpus, so we could not test our model on a larger dataset of a single speaker.
+In comparison to this, the dataset used for CART contained 624 sentences, and a total of 4932 orthographically different words, which results in more than a double sized corpus.
 

### 2017-04-06 Paper/Paper_v1.tex
a table
Another table

@@
 As the results of the initial two models were encouraging, we proceeded with a ``real`` model.
-For this, we reconstructed to a great extent a previous experiment \cite{Brinckman2003} that used a CART tree to predict phoneme duration, and reached good approximation values.
+For this, we reconstructed to a great extent a previous experiment \cite{Brinckmann_2003} that used a CART tree to predict phoneme duration, and reached good approximation values.
 By using this model we also had the advantage of having a reference for comparison.
@@
 The results were very similar, so present here only the results obtained using the model tree M5P, as they were slightly better.
-As can be seen in Table ...
-adding the speech rate feature caused a significant improvement of all our models.
+As can be seen in Table \ref{tab:perfM} adding the speech rate feature caused a significant improvement on all our models.
 
-Table
+\begin{table}[htbp]
+\centering
+%\vspace{10mm}
+\begin{tabular}{|l|c|c|c|c|c|}
 
-However, the performance of our model tree is smaller than the performance of the CART built by Brinckmann \& Trouvain \cite{Brinckman2003}.
+\hline
+	 & Basic - SR & Basic + SR & M5P - SR & M5P + SR & CART\\
+\hline
+\hline
+Correlation coef 		& .42  &  & .50 & .79 & .86 / .83\\
+	\hline
+MAE  &  31.02 & & 34.21 & 25.19 & 22.46 / 21.40\\
+	\hline
+RMSE  &  47.09 & & 62.91 & 44.70 & NA \\
+	\hline
+
+\end{tabular}
+\caption{Overview of the performance results of 3 models with (+) and without (-) speech rate (SR).
+CART is the model created by Brinckmann \& Trouvain \cite{Brinckmann_2003}.
+The two values in the CART cells represent the performance obtained for 2 different speakers.}
+\label{tab:perfM}
+\end{table}
+
+However, the performance of our model tree is smaller than the performance of the CART built by Brinckmann \& Trouvain \cite{Brinckmann_2003}.
 This is explainable if we consider the nature of the used datasets.
@@
 In comparison to this, the dataset used for CART contained 624 sentences, and a total of 4932 orthographically different words, which results in more than a double sized corpus.
+
+\begin{table}[htbp]
+\centering
+%\vspace{10mm}
+\begin{tabular}{|l|c|c|c|c|}
+
+\hline
+	 & \shortstack{Train: whole \\ Test: unseen speaker} & \shortstack{Train\&Test (+SR): \\ speaker AKX} & \shortstack{Train\&Test (-SR): \\ speaker AKX} \\ 
+\hline
+\hline
+Correlation coef 		& .76   & .75 & .63 \\
+	\hline
+MAE  &  22.91 & 22.47 & 26.95 \\
+	\hline
+RMSE  &  40.36 & 33.87 & 40.30 \\
+	\hline
+
+\end{tabular}
+\caption{} 
+\label{tab:perfM}
+\end{table}
 

### 2017-04-06 Paper/Paper_v1.tex
.

@@
 On the other hand, our model tree shows robustness with respect to factors like inter-speaker variability and speech rate.
-The results obtained for a test set containing data from a speaker not included in the training data were similar to those obtained on known speakers on terms of correlation coefficient, and showed less error rates, as can be seen in table ....
+The results obtained for a test set containing data from a speaker not included in the training data were similar to those obtained on known speakers on terms of correlation coefficient, and showed less error rates, as can be seen in table \ref{tab:perfM5P}.
 Moreover, if the model is to be evaluated on data produced by only one speaker, it produces good results even when using less than 10\% of the amount of data used to train on the multiple speaker corpus.
@@
 \end{tabular}
-\caption{} 
-\label{tab:perfM}
+\caption{Overview of the performance of our M5P model tree on different training and test sets.} 
+\label{tab:perfM5P}
 \end{table}

### 2017-04-06 Paper/Paper_v1.tex
some things on vowels

@@
 The phoneme inventory used in Verbmobil includes 52 classes defined according to the German SAMPA.
-Using the classical phonetic conventions, we can group these phonemes on a first level according to their phonetic features using a tree structure into vowels, consonants and a glottal stop, although the latter one is sometimes treated as a consonant \cite{Kohler1995}.
+However, only 49 of these occur in the examined corpus.
+The missing ones are \texttt{/2/} as in \textit{Ökonom}, and the nasals \texttt{/E\~/} and \texttt{/O\~/} as in the loanwords \textit{Teint, Saison}.
+Using the classical phonetic conventions, we can group these phonemes using a tree structure.
+The first branching splits them into vowels, consonants and a glottal stop, although the latter one is sometimes treated as a consonant \cite{Kohler1995}.
 Vowels represent 24\% of the total phoneme inventory analyzed.
@@
 
-The literature records on many potential influencing factors for the duration of phonemes, which vary not only from one author to another but also from one language to another.
-For the German language, intuition aside, we have information on such factors provided by: Kohler \cite{Kohler1992}, Riedi \cite{Riedi1995}, Moebius \cite{Moebius1996}, and Brinckmann \& Trouvain \cite{Brinckmann_2003}, who applied some of the existent segment duration models to the German language.
+\section{Vowels}
+Vowels are an important class, because they show a high degree of elasticity in terms of duration, and correlate well with the speech rate.
+The minimum value recorded for vowels lies at 0.03 sec, while their maximum is 2.3 sec.
+The variation inside the value group may be expressed in terms of standard deviation with a $\sigma$ value of 0.07 sec.
+However, most values lie within a much smaller range, and 75\% of the vowels are shorter than 0.1 sec.
+If we remove 1\% of the data at the higher end, $\sigma$ reduces to 0.5 sec and the maximum duration at 0.3 sec.
+After carefully analyzing some of the outliers, we decided to keep them.
+We observed that some vowels may last as long a one can hold breath, like in the case of \texttt{/a:/}.
+Due to this high elasticity, vowels are the best candidates to start working on any phoneme duration model, as they are extremely sensitive to duration influencing factors.
 
-As we recreated the tree model of Brinckmann \& Trouvain \cite{Brinckmann_2003} for the most part, we also selected most of the features used by them. Here just a rough listing of these factors:
-
-\begin{itemize}
-	\item phoneme id
-	\item vowel and syllable stress
-	\item manner of articulation for consonants
-	\item position of phoneme in syllable
-	\item position of syllable in word
-	\item position of word in turn
-	\item part-of-speech: function word, content word, proper name, interjection-or-answer-particle
-	\item speech rate
-\end{itemize}
-
-A detailed description of the model implementation and phoneme features used follows in Chapter 6.
-
-\section{Vowels}
-Vowels are an important class, because their variation in terms of duration correlates well with the speech rate.
 One can say that vowels generally tend to get proportionally longer when the words themselves are spoken slower.
+We investigated this assumption by comparing the vowel duration inside words containing a specific number of phonemes, as well as the variation of the word time proportion occupied by them in such words.
+The results confirmed this hypothesis, as you may see in fig.
+\ref{fig:w_dur_vs_pho} and \ref{fig:boxplot_prop}.
+This also means that it is a good idea to look for the correlation between vowels and speech rate.
 We obtained an overall correlation coefficient of 0.72 for vowels against speech rate.
@@
 0.47.
+
+\begin{figure}[htbp]
+	\noindent\makebox[\textwidth]{\includegraphics[width=\textwidth, clip]{../Graphen/Stats_for_genau.eps}}
+	\centering
+	\vspace{-10mm}
+	\caption[Word duration of ``genau`` vs.
+its phoneme durations]{We ordered ascendingly the 706 occurences of the word ``genau`` based on word duration and ploted recorded durations of the component phonemes against it.
+The picture suggests clearly that the diphthong \texttt{/aU/} occupies a rather fixed steak of the total word duration, while the others show no adjustment to the increasing word total duration.
+}
+	\label{fig:w_dur_vs_pho}
+\end{figure}
+
+\begin{figure}[htbp]
+	\includegraphics[width=.55\linewidth]{../Graphen/box3_oyauaiaab.eps}
+	\centering
+	\caption[Proportions of diphthong in 3-phoneme-words]{The time slice occupied by vowels in words of specific length  vary relatively little, indicating a possible linear correlation between phoneme duration and the variation of word duration.
+The y-axis measures time in sec.}
+	\label{fig:boxplot_prop}
+\end{figure}
 
@@
 The relation between phoneme duration and factors other than speech rate was treated only marginally, as these have already been extensively examined in the literature on various corpora.
-At this stage, we considered such factors only to the extent they might have helped identify groups of phonemes that may be best suited to evaluate the influence of speech rate.
+For the German language, intuition aside, we have information on such factors provided by: Kohler \cite{Kohler1992}, Riedi \cite{Riedi1995}, Moebius \cite{Moebius1996}, and Brinckmann \& Trouvain \cite{Brinckmann_2003}, who applied some of the existent segment duration models to the German language.At this stage, we considered such factors only to the extent they might have helped identify groups of phonemes that may be best suited to evaluate the influence of speech rate.
 Such a factor is primary stress, which creates a group of very elastic vowels, which are therefore also sensitive to changes in speech tempo.

### 2017-04-06 Paper/Paper_v1.tex
Phonemes added.

@@
 The correlation value for the long vowel group is even greather than that of the primary stressed one, amounting to 0.79.
-	\item Short vowels in opposition to the above mentioned vowel categories.
+	\item Short vowels, in opposition to the above mentioned vowel categories, group phonemes that are to be considered as vowels from the point of view of their production, but which show a relatively low elasticity with respect to duration, if we compare them to the other three vowel groups.
 \end{itemize}
+
+Table \ref{tab:vowel_stats} below presents the main duration information regarding the above mentioned classes, sustaining our comments on them.
+\begin{table}[htbp]
+\centering
+%\vspace{10mm}
+\begin{tabular}{|l|c|c|c|c|c|c|c|}
+
+\hline
+	 & Min value & Max value & Mean & St. Dev. & Q25 & Q75 & Count\\
+\hline
+\hline
+Diphthongs  	& .04  & .71 & .12 & .06 & .08 & .15 & 8495\\
+	\hline
+Long vowels   & .04 & 2.3 &  .11 & .08 & .06 & .13 & 30425\\
+	\hline
+Schwa  &  .03 & 1.02 & .08 & .07 & .04 & .09 & 21637 \\
+	\hline
+Short vowels & .03 & .81 & .07 & .04 & .04 & .08 & 42570 \\
+	\hline
+\end{tabular}
+\caption*{} 
+\label{tab:vowel_stats}
+\end{table}
+
+The vowel tree can be continued further down, introducing several further quality related features from the domain of the articulation manner, which leads to single phonemes at the leaves.
+However, our purpose was to group vowels, in order to get an overview about the phoneme durations for the beginning, so we don't go further down the tree, but just show for the record fig.
+\ref{fig:vow_square} on locating vowels according to their place of creation.
+
+\begin{figure}[htbp]
+	\includegraphics[width=.55\linewidth]{../Graphen/vow_art.eps}
+	\centering
+	\caption[Vowel square]{Vowels are produced by the air flow in an open vowel tract, and they differ based on the modification we add to the shape of this cavity.
+The only way for humans to produce such modifications are by moving the thongue and/or the lips.
+These changes are usually represented in the so-called vowel square, which may be further simplified to a vowel triangle.
+Here you can see how to place a vowel square containing the main vowels in the mouth cavity.
+}
+	\label{fig:vow_square}
+\end{figure}
 
@@
 \section{Consonants}
+In general, we can say that consonants present different degrees of elasticity, which are, however, generally lower than the ones of vowels.
+
 Based on the correlation between phoneme duration and speech rate we created a subgroup of consonants containing the nasals and \texttt{/x/, /C/, /h/} with a relatively good correlation value $\geq 0.7$, and considered plosives to be a subgroup because of their consistently low correlation $\leq 0.42$.
@@
 
-Below an overview of the classical consonant categories.
+Same as vowels, the consonant branch of the tree may be split further down into smaller categories.
+We chose to include them in our phoneme overview, as some of these categories proved to be relevant for the modelling of phoneme durations as well.
+Consequently, level 2 split shows for for the consonant node following categories:
 \begin{itemize}
-	\item Plosives, which are generally short, is due to their articulation characteristics.
-	\item Fricatives, which show a legth variation almost comparable to that of vowels.
-	\item Nasals: /m/, /n/, /N/.
-	\item Specials: /h/, /l/, /C/, /j/, /r/
+	\item Plosives, which are generally short, due to their articulation characteristics.
+Their production involves two separate phases: closure, when the vocal tract is completely closed, and release, when the accumulated air flow is let out as in a burst.
+Considering this, it is straightforward, that the first phase may be longer, limited only by the human physiology, while the second one is technically extremely short.
+Some speech corpora, like PhonDat, label the two phases differently.
+Verbmobil doesn't, so we had to take them as a whole.
+We keep these considerations in mind, when referring to the special case of \texttt{/t/} that we found in our corpus.
+	\item Fricatives, which show a duration variation almost comparable to that of vowels.
+This fact can also be explained by their production process: the vocal tract is not completely closed, but the air flow is forced out through a narrowing at some point in the vocal tract, so that a friction sound is produced.
+	\item Nasals also show an elasticity comparable to that of vowels.
+Again this is explainable through their production: the oral passage is blocked, while lowering the velum, so that the air can escape through the nose.
+Nasals may also constitute the nucleus of a syllable, when the vowel gets reduced.
+/m/, /n/, /N/.
+	\item Liquids, also called approximants, are produced by partial closure of the mouth produced by the thongue.
+German knows two liquids \texttt{/l/, /r/}.
+One relevant feature is that they have the greatest freedom to occur in consonant clusters, that is a group of consonants belonging to the same syllable part.
+This also means that they must be particularly short, as we identified vowels as the greater steak holders in word duration.
+Another fact we considered, is the \texttt{/r/} is relatively seldom in German, being often reduced to a schwa, and when it is articulated, then it is generally extremely short, which durations comparable to the ones of the glottal stop.
 \end{itemize}
+
+Plosives and fricatives may be split further down into voiced, if the vocal cords are participating in the production of the sound, and voiceless, when the vocal cords do not participate.
+We didn't analyse these groups in particular, as we focused on vowels.
+
+Table \ref{tab:cons_stats} below presents the main duration information regarding the above mentioned classes, sustaining our comments on them.
+\begin{table}[htbp]
+\centering
+%\vspace{10mm}
+\begin{tabular}{|l|c|c|c|c|c|c|c|}
+
+\hline
+	 & Min value & Max value & Mean & St. Dev. & Q25 & Q75 & Count\\
+\hline
+\hline
+Plosives  	& .03  & 1.50 & .06 & .05 & .03 & .07 & 45477\\
+	\hline
+Fricatives   & .03 & 1.89 &  .06 & .05 & .03 & .08 & 51667\\
+	\hline
+Nasals  &  .03 & 1.46 & .08 & .08 & .04 & .09 & 39936 \\
+	\hline
+Liquids /l/ & .03 & 1.51 & .06 & .06 & .03 & .07 & 8019 \\
+	\hline
+Liquids /r/ & .03 & .46 & .05 & .03 &  .03 & .06 & 5453 \\
+	\hline
+\end{tabular}
+\caption{} 
+\label{tab:cons_stats}
+\end{table}
 
@@
 	\hline
-
 \end{tabular}

### 2017-04-06 Paper/Paper_v1.tex
changed a title
Cleaned and organized wekafiles folder. Created a Readme for the wekafiles. Only cleaned py_file folder.

@@
-\documentclass[a4paper]{scrreprt}
+\documentclass[11pt,a4paper]{scrbook}
+\usepackage{geometry}	
 \usepackage[backend=bibtex,style=numeric]{biblatex}
@@
 \bibliography{../PaperBib}
+\usepackage{colortbl}	
+\usepackage{xcolor}
+\usepackage{soul}
 \usepackage{amsmath}
@@
 \usepackage{scrhack}
+\renewcommand{\familydefault}{\sfdefault}
+\definecolor{uhhred}{cmyk}{0,100,100,0}
 \usepackage{tabularx}
@@
 
-\title{Improving the modelling of tempo and phoneme duration}
-\author{Alexandra Krah}
-\date{06. April ~2017}
-\maketitle
-\chapter*{}
+\frontmatter
+\newgeometry{centering,left=2cm,right=2cm,top=2cm,bottom=2cm}
+\begin{titlepage}
+\includegraphics[scale=0.3]{../Graphen/UHH-Logo_2010_Farbe_CMYK.pdf}
+\vspace*{2cm}
+\Large
+\begin{center} 
+      {\color{uhhred}\textbf{\so{BACHELORTHESIS}}}
+%oder {\color{uhhred}\textbf{\so{MASTERTHESIS}}}
+\vspace*{2.0cm}\\
+{\LARGE \textbf{Improving the modelling of tempo and phoneme duration}}
+\vspace*{2.0cm}\\
+vorgelegt von
+\vspace*{0.4cm}\\
+Alexandra Krah
+\end{center}
+\vspace*{3.9cm}
+
+\noindent 
+MIN-Fakultät \vspace*{0.4cm} \\ 
+Fachbereich Informatik \vspace*{0.4cm} \\ 
+%Ggf. Professur/Institut \vspace*{0.4cm} \\
+Studiengang: Informatik \vspace*{0.4cm} \\ 
+Matrikelnummer: 6415559 \vspace*{0.8cm} \\ 
+Erstgutachter: Dr. Timo Baumann \vspace*{0.4cm} \\ 
+Zweitgutachter: Prof. Dr. Wolfgang Menzel
+
+\end{titlepage}
+
+\restoregeometry
+
 \tableofcontents
-\listoffigures
 
@@
 The fact that we understand people at all natural speech rates is certainly largely due to the quality features of sounds (phonemes) in the given language.
-However, if we have a perfect audio recording of speech, and play it faster or slower, we immediatly notice how naturalness deteriorates.
-Synthetic speech shows a similar behaviour when changing the speaking rate.
+However, if we have a perfect audio recording of speech, and play it faster or slower, we immediately notice how naturalness deteriorates.
+Synthetic speech shows a similar behavior when changing the speaking rate.
 In both cases phoneme quality doesn't need to change.
@@
 
-The attempts to improve the quality of speech synthesis when tempo changes include linguistic processing, and phoneme length manipulation before the actual synthesis takes place.
+Attempts to improve the quality of speech synthesis when tempo changes also include linguistic processing, and phoneme length manipulation before the actual synthesis takes place.
 As mentioned in a patent documentation of Fujitsu Limited, phoneme length manipulation plays an important role in this context and one cannot rely on a linear adjustment function \cite{nishiike2008}.
@@
 Phoneme length adjustment is the aspect we wish to adress in this work.
-In particular, we want to find an algorithm for adjusting phoneme durations according to the variations in speech rate.
+In particular, we want to find an algorithm for adjusting phoneme durations according to the variation of the speech rate.
 The resulting phoneme duration should be close to the actual duration that occurs when the speaking rate alters.
-This would be the same as playing an audio file faster or slower.
+This would be the same as playing an audio recording faster or slower.
 
-There are several models for phoneme duration approximation, that work pretty well for some languages, German inclusive.
+There are several models for phoneme duration approximation that work pretty well for some languages, German inclusive.
 However, they consistently don't take the speech rate into consideration, their performance level being acquired rather within a specific speech rate.
@@
 Moreover, all these models have an upper asymptote for their accuracy situated at less than 100\%.
-It has been suggested, that the influence of so-called ``macroscopic`` \cite{Cummins1999} or para-linguistic \cite{Santen1994} factors should be examined in order to move this asymptote further up.
-So we focused on speech rate.
+It has been suggested, that the influence of so-called ``macroscopic``or para-linguistic \cite{Santen1994} factors should be examined in order to move this asymptote further up.
+Such a factor is speech rate, and we took the challenge of integrating it into a speech model.
 
-The purpose is therefore not to find another duration approximation model for TTS-systems, but to improve the algorithm of phoneme duration prediction by considering the speech tempo.
-Table \ref{tab:Question} vizualises our challenge: we want to put all available/visible information from the table in relation so that we can replace the question marks with values that are close to the real ones.
+The purpose is therefore not to create another duration approximation model for TTS-systems, but to improve the algorithm of phoneme duration prediction by considering the tempo of speech.
+Table \ref{tab:Question} visualizes our challenge: we want to put all available/visible information from the table in relation so that we can replace the question marks with values that are close to the real ones.
 Consequently, the modified speech unit would sound more natural at a faster or a slower speech rate as when being modified with a linear function.
@@
 \end{tabular}
-\caption{The first column shows our inputs: duration of each phoneme in German word ``genau``, given a speech rate of 0.07 sec, or a word duration of 0.28 sec.
-The second column contains our challenge: How long is the duration of each phoneme of the given word, at a speech rate of 0.11 sec, equivalent of a word length of 0.42 s ?}
+\caption{The first column shows our inputs: Duration of each phoneme in German word ``genau``, given a speech rate of 0.07 sec, or a word duration of 0.28 sec.
+The second column visualizes our challenge: How long is the duration of each phoneme of the given word, at a speech rate of 0.11 sec, equivalent of a word length of 0.42 sec?}
 \label{tab:Question} 
@@
 phonemes) \cite{Hoequist1986}.
-Indeed, if one simply ``stretches`` all phonemes in the same manner to accomodate a new speech rate, the result is rather dissapointing, as one can see in Fig.
+Indeed, if one simply ``stretches`` all phonemes in the same manner to accommodate a new speech rate, the result is rather disappointing, as one can see in Fig.
 \ref{fig:linear_approx}.
-
 \begin{figure}[htbp]
@@
 	\includegraphics[scale=0.7]{../Graphen/Linear_approx.eps}
-	\caption[Linear phoneme duration approximation]{This is a linear solution to the issue presented in table \ref{tab:Question} for the German word ``genau``}
+	\caption[Linear phoneme duration approximation]{This is a linear solution to the challenge presented in table \ref{tab:Question}.}
 	\label{fig:linear_approx}
@@
 
-\subsubsection*{Notizen}
-Syllable and foot structure of the utternace belong to the important factors influencing the segment length modification  \cite{Hoequist1986}.
-
-More important as the segmentation of the utterance in words is its segmentation in prosodical segments, so-called feet.
-In spontaneous speech, some word parts may fall, others may be uttered together with other words so that the resulting phoneme duration tends to relate more to the structure of the resulting prosodic segment as to the original word \cite{Kohler1986}.
-
-Accross the paper we will use the German SAMPA notations when referring to phonemes.
+Across the paper we use the German SAMPA notations when referring to phonemes.
 
 \subsection*{Outline}
-We start our approach by defining our methods in chapter 2.
-In the following chapter we examine our database so that we can proceed with the actual task in chapter 4, where we analyze the German phonemes.
-We dedicate chapter 5 to the challenge of finding an adequate speech rate definition for our database and purpose.
+We start our challenge by defining our methods in chapter 2.
+In the following chapter we examine the available resources so that we can proceed with the actual task in chapter 4, where we analyze the German phonemes.
+We dedicate chapter 5 to the challenge of finding an adequate speech rate definition for our database, and which can serve our purpose.
 Chapter 6 finally deals with phoneme duration prediction models and the challenge of finding a solution to our above presented issue.
 
-\chapter{Basics}
+\chapter{Means and Methods}
 
@@
 
-We used machine learning to solve a task involving both classification and numeric prediction.
-We used for this two model tree methods: M5P and REPTree.
-The reasons we chose them is because trees are easy to interpret, they can perform both classification and numeric prediction tasks, we could easily generate them using Weka, and are the closest to classification and regression (CART) trees, also used in the literature \cite{Brinckmann2003} for phoneme duration prediction, so we have data to compare with.
+We used machine learning to solve a task involving both classification and numeric prediction, and selected from its repertoire two model tree methods: M5P and REPTree.
+The reasons we chose are: trees are easy to interpret, they can perform both classification and numeric prediction tasks, we could easily generate them using Weka, and they are the closest to classification and regression (CART) trees, also used in the literature \cite{Brinckmann_2003} for phoneme duration prediction, so we have data to compare with.
 
@@
 If one needs to combine the two methods, he/she may use a classification and regression tree, called CART, or a model tree.
-The advantage of trees is that new instances are easily assigned to a class, or a numeric value, or a linear model which leads to such a numeric value, and they are easy to unterstand.
+The advantage of trees is that new instances are easily assigned to a class, or a numeric value, or a linear model which leads to a numeric value, and they are easy to unterstand.
 
-The \textbf{M5P model tree} \cite{Witten2011} is a decision tree based on the M5 algorithm introduced by Quinlan (1992).
-It builds a usual classification tree out of the attributes, and retrieves a linear regression model in each leaf out of the attributes contained in that specific subtree.
+The \textbf{M5P model tree} is a decision tree based on the M5 algorithm introduced by Quinlan (1992).
+It builds an ordinary classification tree out of the selected attributes, and retrieves a linear regression model in each leaf out of the attributes contained in that specific subtree.
 Branching is realised based on the reduction of variation within the resulting subset.
@@
 
-Both trees cope well with missing data. However, both need a very large training set in order to perform well.
+Both trees cope well with missing data. However, both need large training sets in order to perform well.
 
@@
 A very important and time consuming step in data mining is data preparation, which is the next step after data collection.
-At this stage one needs to asses the quality of the data, expressed as noise ratio, and strategies for dealing with it.
-If the data used has been originally collected for another purpose, then it needs to be assessed in terms of features it contains, actually needed features, and organization resp.
+At this stage, one needs to assess the quality of the data, expressed as noise ratio, and the strategies for dealing with it.
+If the data used has been originally collected for another purpose, then it needs to be assessed in terms of features it contains, actually needed features, and organization, resp.
 reorganization of data to serve new purpose.
 
-Considering machine learning, it is straightforward that one needs to split the available dataset at least into two parts: a (bigger) training dataset, used for training the model, meaning discover patterns in data, and a (smaller) test dataset, used for testing the performance of the model resulting from the machine learning process.
+Considering machine learning, it is straightforward that one needs to split the available dataset at least into two parts: a (bigger) training dataset, used for training the model, meaning to discover patterns in data, and a (smaller) test dataset, used for testing the performance of the model resulting from the machine learning process.
+We used the well-established split of 90\% training data and 10\% test data.
 
@@
 We can interpret this as how much the variable movement through its value set corresponds to the movement of the other variable in its own value set.
-Values close to 1 mean that greater values of the one variable imply greater values for the second variable, negative values mean the two variables move in opposite directions, and values close to 0 mean there is no correlation between the two variables.
+Values close to 1 mean that greater values of the one variable imply greater values for the second variable, negative values mean the two variables move in opposite directions, while values close to 0 mean there is no correlation between the two variables.
 However, one must keep in mind that correlation does not imply causality, so the presence of correlation between two variables does not mean that one of them has any influence on the other one.
 If we compare red Ferrari cars to blue Renault cars, one may find a correlation between car color red and car speed.
-However, this does not imply that red cars are faster.
+However, this does not imply that red cars are generally faster.
 \end{enumerate}
@@
 \section{Tools}
-We used \textbf{Weka} 3.8 for building and training the model trees for the phoneme duration prediction.
+We used \textbf{Weka} 3.8 for building and training the model trees for phoneme duration prediction.
 It is an open source freeware using a Java VM, and contains a collection of tools for all machine learning tasks we needed, from data pre-processing over classification and regression till model evaluation and data visualization.
@@
 There are three corpora which we had at our disposal: (1) The Kiel Corpus of Read (PhonDat) and of Spontaneous Speech, (2) Verbmobil, and (3) spoken Wikipedia.
-The first one contains two equal sized corpora of > 4 hours each: KCoRS, which is a collection of read speech, and KCoSS, which is a collection of spontaneous speech, both being manually segmented, which is a good indicator for good segmentation quality.
+The first one contains two equal sized corpora of > 4 hours each: KCoRS, which is a collection of read speech, and KCoSS, which is a collection of spontaneous speech, both being manually segmented, a good indicator for good segmentation quality.
 However, as Pfitzinger \cite{Pfitzinger2002} also showed in his reassessment of the PhonDat-II corpus, for nowadays research the Kiel Corpus is rather small.
@@
 This was an important point in selecting this database for our purpose, as phoneme duration is influenced by many factors, and the Verbmobil textual database provides most of them.
-We used 286 recordings from the last verbmobil phase (Verbmobil II) created by 131 speakers.
-The total VMII corpus used amounts to 11783 conversational turns representing recorded speech time of 15,5 hours.
+We used 286 recordings from the last verbmobil phase (VM-II) created by 131 speakers.
+The total VM-II corpus used amounts to 11783 conversational turns representing recorded speech time of 15.5 hours.
 The ``symbolic data`` corresponding to each turn (e.g.
@@
 This means one starts his turn at a normal speech rate, then gets to a point where he/she hesitates and gets slower, like when weighting different transport means to get to the appointment they are to agree upon.
-This results in the challenge of finding an appropriate description for the speech rate to be used for modeling phoneme durations across the corpus.
+This results in the challenge of finding an appropriate description for the speech rate to be used for the modeling of phoneme durations across the corpus.
 We handle this issue of speech rate separately in chapter 5.
@@
 The exact coordinates of phoneme durations vary across speakers, as one can see in Fig.
-\ref{fig:speaker_cmp} for the case of \texttt{/a:/}.
+\ref{fig:speaker_cmp}, for the case of the phoneme \texttt{/a:/}.
 
@@
 
-\chapter{Phoneme durations and the dynamics of speech}
+\chapter{German phoneme system}
 The phoneme inventory used in Verbmobil includes 52 classes defined according to the German SAMPA.
@@
 Vowels represent 24\% of the total phoneme inventory analyzed.
-Further relevant groupings are being examined under sections ``Vowels`` and ``Consonants``.
+Further relevant groupings are discussed in subsections ``Vowels`` and ``Consonants``.
 
-Pauses may be caused and/or be quantitatively modified both by linguistic (e.g.
+Pauses may be caused by and/or be quantitatively modified both by linguistic (e.g.
 phonological) and non-linguistic factors, like technical, psychological, dialog dynamics, etc.
-Consequently pause length variation needs to be treated separately from phoneme duration variation, as a not negligeable large set of other factors are to be considered for this, which extends beyond the purpose of this work.
+Consequently pause length variation needs to be treated separately from phoneme duration variation, as a not negligible large set of other factors are to be considered for this, such as dialogue dynamics, hesitations, technical issues of the recording, which extends beyond the purpose of this work.
 
 \section{Vowels}
-Vowels are an important class, because they show a high degree of elasticity in terms of duration, and correlate well with the speech rate.
+Vowels are an important class with respect to tempo modelling, because they show a high degree of elasticity in terms of duration, and correlate well with the speech rate.
 The minimum value recorded for vowels lies at 0.03 sec, while their maximum is 2.3 sec.
@@
 However, most values lie within a much smaller range, and 75\% of the vowels are shorter than 0.1 sec.
-If we remove 1\% of the data at the higher end, $\sigma$ reduces to 0.5 sec and the maximum duration at 0.3 sec.
+If we remove 1\% of the data at the higher end, $\sigma$ reduces to 0.05 sec and the maximum duration at 0.3 sec.
 After carefully analyzing some of the outliers, we decided to keep them.
-We observed that some vowels may last as long a one can hold breath, like in the case of \texttt{/a:/}.
+We observed that some vowels may last as long as one can hold breath, like in the case of \texttt{/a:/}.
 Due to this high elasticity, vowels are the best candidates to start working on any phoneme duration model, as they are extremely sensitive to duration influencing factors.
@@
 We investigated this assumption by comparing the vowel duration inside words containing a specific number of phonemes, as well as the variation of the word time proportion occupied by them in such words.
-The results confirmed this hypothesis, as you may see in fig.
+The results confirmed this hypothesis, as you may see in Fig.
 \ref{fig:w_dur_vs_pho} and \ref{fig:boxplot_prop}.
 This also means that it is a good idea to look for the correlation between vowels and speech rate.
-We obtained an overall correlation coefficient of 0.72 for vowels against speech rate.
-Inside the vowel group we could identify the primary stressed vowels as the actual correlation owners, with a group correlation value of 0.75, while the unstressed and secondary stressed ones showed rather modest correlation values of 0.56, resp.
-0.47.
+We obtained an overall correlation coefficient of $r = 0.72$ for vowels against speech rate.
+Inside the vowel group, we could identify the primary stressed vowels as the actual correlation owners, with a group correlation value of $r = 0.75$, while the unstressed and secondary stressed ones showed rather modest correlation values of $r = 0.56$, respectively $r = 0.47$.
 
@@
 its phoneme durations]{We ordered ascendingly the 706 occurences of the word ``genau`` based on word duration and ploted recorded durations of the component phonemes against it.
-The picture suggests clearly that the diphthong \texttt{/aU/} occupies a rather fixed steak of the total word duration, while the others show no adjustment to the increasing word total duration.
+The picture suggests clearly that the diphthong \texttt{/aU/} occupies a rather fixed steak of the total word duration, while the other phonemes show no adjustment to the increasing word total duration.
 }
@@
 	\caption[Proportions of diphthong in 3-phoneme-words]{The time slice occupied by vowels in words of specific length  vary relatively little, indicating a possible linear correlation between phoneme duration and the variation of word duration.
-The y-axis measures time in sec.}
+The y-axis measures time in seconds.}
 	\label{fig:boxplot_prop}
@@
 \begin{itemize}
-	\item Diphthongs represent two vowels that are articulated together in a syllable.
+	\item Diphthongs represent a slide between two vowels that are articulated together in a syllable.
 They are generally characterized by longer durations.
@@
 It is also called a reduction vowel \cite{Kohler1995}, as most of the other monophthongs tend to be articulated as schwa in non-stressed vowels.
-Verbmobil differentiates two types of schwa for the German language: \texttt{/@/} like in the last syllable of \textit{lesen} and \texttt{/6/} like in \textit{Leser}, where the reduced vowel is followed by an \texttt{/r/}.
+Verbmobil differentiates two types of schwa for the German language: \texttt{/@/} like in the last syllable of \textit{lesen} and \texttt{/6/} like in \textit{Leser}, where the reduced vowel is followed by an \texttt{/r/} in writing.
 	\item Long vowels.
 As long vowels usually carry the primary stress, as some german phonetics manuals such as the one of Kohler \cite{Kohler1995} suggest, we considered the long vowels as a standalone subclass of vowels.
-The correlation value for the long vowel group is even greather than that of the primary stressed one, amounting to 0.79.
+The correlation value for the long vowel group is even greather than that of the primary stressed one, amounting to $r = 0.79$.
 	\item Short vowels, in opposition to the above mentioned vowel categories, group phonemes that are to be considered as vowels from the point of view of their production, but which show a relatively low elasticity with respect to duration, if we compare them to the other three vowel groups.
@@
 
-Table \ref{tab:vowel_stats} below presents the main duration information regarding the above mentioned classes, sustaining our comments on them.
+Table \ref{tab:vowel_stats} below presents the main duration information regarding the above mentioned classes that consolidates the discussion presented above.
 \begin{table}[htbp]
@@
 \end{tabular}
-\caption*{} 
+\caption{Numeric information on the duration of the main vowel classes. The unit of the values presented in all columns except ``Count`` is second.} 
 \label{tab:vowel_stats}
@@
 The vowel tree can be continued further down, introducing several further quality related features from the domain of the articulation manner, which leads to single phonemes at the leaves.
-However, our purpose was to group vowels, in order to get an overview about the phoneme durations for the beginning, so we don't go further down the tree, but just show for the record fig.
-\ref{fig:vow_square} on locating vowels according to their place of creation.
+It is good to have an overview on the articulation of vowels, i.e.
+on locating vowels according to their place of creation, and Fig.
+\ref{fig:vow_square} illustrates this very well.
+However, our purpose was to group vowels in order to get an overview about the phoneme durations for the beginning, so we don't go further down the tree.
 
@@
 	\caption[Vowel square]{Vowels are produced by the air flow in an open vowel tract, and they differ based on the modification we add to the shape of this cavity.
-The only way for humans to produce such modifications are by moving the thongue and/or the lips.
+The only way for humans to produce such modifications are by moving the tongue and/or the lips.
 These changes are usually represented in the so-called vowel square, which may be further simplified to a vowel triangle.
@@
 Some phonetic manuals don't event consider it a phoneme \cite{Ternes2012}.
-Furthermore, it's acoustical properties make it look like a (filled) pause in speech, as you may see in Fig.
-..., which represents a challenge for the transcription.
+Furthermore, its acoustical properties make it look like a (filled) pause in speech, which represents a challenge for the transcription.
 As our files were segmented automatically, there is place for doubt about the accuracy of the collected data for \texttt{/Q/}.
@@
 
-The information collected from our database show a very large variation of it's duration from 0.04 sec.
-to 2.05 sec., with $\sigma = 0.06$, consistent with the official data on the complete VMII.
+The information collected from our database show a very large variation of its duration from 0.04 sec.
+to 2.05 sec., with $\sigma = 0.06$, consistent with the official data on the complete VM-II.
 However, if we remove the upper 0.8 \% of the duration data, we obtain a dataset having only half of the mentioned variation with $\sigma = 0.03$ and a maximum duration of 0.3 seconds.
@@
 \hline
-Turn 		& $turn\_dur / phon\_count$ &0.727178570436 \\
-	\hline
-	&  $phon\_count / turn\_dur$ & -0.270392865503 \\
-	\hline
+
 Phrase  & $phrase\_dur / phon\_count$ & 0.561266362391\\
 	\hline
-	&  $phon\_count / phrase\_dur$ & -0.377058296098\\
-	\hline
 Word  & $word\_dur / phon\_count$ &  0.794901825792\\
-	\hline
-	&  $phon\_count / phrase\_dur$ & -0.616539202577\\
 	\hline
@@
 If one would apply a clustering method on phonemes to group them on seven classes based on their duration interval coordinates using, say k-means method, the resulting groups would certainly differ from those listed in Chapter 4.
-Fig?!
 
@@
 These values were calculated using the mean duration of the specific phoneme across the corpus modified or not by a proportion of its standard deviation, based on the speech rate value split.
-Figure ...
-shows the structure of this tree.
 \end{enumerate}
@@
 One possible result of such an approach would be the prediction of phonemes reaching $duration = 0$ at specific speech rates, which means predict phoneme drop phenomena.
-
-\subsubsection*{Notizen}
-Analyze duration changes of vowels in interrupted words.
-Consider actual realized syllables in speech instead of syllables based on the canonical transcription?.
 In spontaneous speech, the word and syllable boundaries change, and this change influences other phonetic aspects such as stress, syllable duration, phoneme duration.
@@
 \printbibliography
+\backmatter
+
+\thispagestyle{empty}
+
+\vspace*{\fill}
+\pagestyle{empty}
+{\normalsize
+\begin{center}\textbf{Eidesstattliche Erklärung}\end{center}
+Hiermit versichere ich an Eides statt, dass ich die vorliegende Arbeit im Bachelorstudiengang Wirtschaftsinformatik selbstständig verfasst und keine anderen als die angegebenen Hilfsmittel – insbesondere keine im Quellenverzeichnis nicht benannten Internet-Quellen – benutzt habe.
+Alle Stellen, die wörtlich oder sinngemäß aus Veröffentlichungen entnommen wurden, sind als solche kenntlich gemacht.
+Ich versichere weiterhin, dass ich die Arbeit vorher nicht in einem anderen Prüfungsverfahren eingereicht habe und die eingereichte schriftliche Fassung der auf dem elektronischen Speichermedium entspricht.
+\vspace*{1cm}\\
+Hamburg, den 06.04.2017
+\hspace*{\fill}\begin{tabular}{@{}l@{}}\hline
+\makebox[5cm]{Vorname Nachname}
+\end{tabular}
+\vspace*{3cm}
+
+}
+\vspace*{\fill} 
 \end{document}

### 2017-04-06 py_files/arff_file_builder.py

@@ -1,88 +1,0 @@
-import phon_dict
-import fileinput
-import re
-import os
-import sys
-import time
-
-# Generates an almost ready for processing with weka text file. 
-time1 = time.time()
-# List of phonemes to search for
-#phoneme_list = ["a", "a:", "E", "E:", "e", "e:", "O", "o", "o:", "@", "6", "9"] # needs 4,5 hrs
-#phoneme_list = ["p", "b", "t", "d", "k", "g"] # needs cca 3hrs (2hrs 45 min)
-#phoneme_list = ["f", "v", "s", "S", "z", "x", "h", "C", "Z"] # ca. 3hrs (3hrs 14 min)
-#phoneme_list = ["m", "n", "N", "l", "r", "j"] # ca. 3 hrs 40 min
-phoneme_list = ["@", "6"] # ca. 1hr 15 min
-# All diphthongs:
-#phoneme_list = ["OY", "aU", "aI"]
-# All long vowels:
-#phoneme_list = ["a:", "e:", "E:", "i:", "o:", "u:", "y:", "2:", "a~:"]
-# All vowels:
-#phoneme_list = ["a", "a~", "e", "E", "I", "i", "O", "o", "U", "u", "Y", "y", "9", "2", "a:", "a~:", "e:", "E:", "i:", "o:", "u:", "y:", "2:", "OY", "aU", "aI"]
-
-# Get the dictionary of phonemes (keys) and their attributes (values)
-phon_dict = phon_dict.phon_dict(phoneme_list)
-#dict_of_lengths = duration_dict.phon_dict(phoneme_list)
-
-# Put all dict items in one list
-dict_list = []
-for el in phoneme_list:
-	dict_list += phon_dict[el]
-
-# Get the labels for the attribute: word
-word_list = set(dict_list[3::13])
-
-
-#test_file = open("C:/Users/alexutza_a/Abschlussarbeit/sp_kurzes_a.txt", "w") -> gibt falsche Anzahl von speech rates
-arff_file = open("C:/Users/alexutza_a/Abschlussarbeit/wekadateien/schwa_usw", "w")
-
-arff_file.write("% ARFF file for Verbmobil\n% Phonemes: @, 6\n%\n@relation schwa\n\n% List of attributes:\n%\n")
-arff_file.write("% Duration is expressed in seconds\n")
-
-arff_file.write("@attribute phoneme { ") 
-for el in phoneme_list:
-	arff_file.write(el + ", ")
[…]

### 2017-04-06 py_files/calc_errors.py

@@ -1,83 +1,0 @@
-import performance_measures
-import numpy as np
-from collections import defaultdict
-import pd_LSR
-
-#phon_dict, test_dict, tm_dict = prepare_data.calculate_pho_mean(prepare_data.read_trainig_files())
-# A dictionary giving the values of Mean and SD in a list for each encountered phoneme.
-stat_dict = pd_LSR.phone_stats(pd_LSR.read_trainig_files())
-#print(test_dict)
-
-#testlist = prepare_data.read_testfiles()
-# Looks like: ["a", 583, 0.5, "b", 12, 0.78, "a", 489, 0.12, ...]
-testlist = pd_LSR.read_testfiles()
-#print(testlist)
-
-predictions = pd_LSR.create_prediction_list(testlist, stat_dict)
-actual = pd_LSR.read_testfiles()[1::3]
-
-
-#print(len(predictions))
-#print(prepare_data.read_testfiles())
-#print(len(actual))
-
-global_rmse, mae = performance_measures.calc_error_for_data(predictions, actual)
-corrCoef = performance_measures.corrCoef(predictions, actual)[0]
-print(global_rmse, mae, corrCoef)
-
-
-### Following is rmse calculation individually (per phoneme) ###
-
-
-
-# Set of occuring phonemes to iterate on when creating dictionary
-#set_phon = set(prepare_data.read_testfiles()[::2])
-set_phon = set(pd_LSR.read_testfiles()[::3])
-#print(set_phon)
-
-# A list of phonemes in occuring order in the actual and pred lists
-#a_phone_list = prepare_data.read_testfiles()[::2]
-a_phone_list = pd_LSR.read_testfiles()[::3]
-#print(a_phone_list)
-
-# Use list comprehension to interweave the phoneme list with the list of predicted durations
-#   so we can create the dictionary of predictions per phoneme
-pred_list = [x for y in zip (a_phone_list, predictions) for x in y]
-
-
-# Dictionary of actual values per phoneme
-a_dict = defaultdict(list)
-# Dictionary of predicted values per phoneme
-p_dict = defaultdict(list)
-
-
-# Populate dictionary of actual values
-for el in set_phon:
-	i = 0
-	for y in testlist:
-		if el == y:
-			a_dict[el].append(testlist[i+1])
-		i += 1
-#print(a_dict)
-# Populate dictionary of predicted values
-for el in set_phon:
-	j = 0
-	for z in pred_list:
[…]

### 2017-04-06 py_files/create_eval_lists.py

@@ -13,6 +13,8 @@ valid_phonemes = ["a", "a~", "e", "E", "I", "i", "O", "o", "U", "u", "Y", "y", "
                  "o:", "u:", "y:", "2:", "OY", "aU", "aI", "@", "6", "z", "S", "Z", "C", "x", "N", "Q", "b", "d", "f", 
                  "g", "h", "j", "k", "l", "m", "n", "p", "r", "s", "t", "v"]
 
+path_list_test = model_utilities.get_path_list('C:/Users/alexutza_a/Abschlussarbeit/DB_Verbmobil/Evaluation/Test_a')
+
 # Model type uses individual phonemes and their median, mean, or both in words having specific no. of phonemes
 # Actually creates the lists of predictions and actual values to evaluate model
 # @param model: 0, 1, or 2 meaning (median, mean, or (median + mean)/2)

### 2017-04-06 py_files/eval_model.py

@@ -9,45 +9,14 @@ import time
 # Create the two lists of values to be compared, for each model a pair
  # - 1st numeric parameter allows chosing the unit on which one evaluates the model (0 = median, 1 = mean, 2 = (median + mean)/2 )
  # - 2nd numeric parameter should not be changed (it's just the indices of the elem in the result vector to be used)
-#evaluation_lists_initial = [(create_eval_lists.create_modelWL_lists(1)[0], create_eval_lists.create_modelWL_lists(1)[1]), 
-#					(create_eval_lists.create_modelWLIP_lists(1)[0], create_eval_lists.create_modelWLIP_lists(1)[1]), 
-#					(create_eval_lists.create_modelIPdur_lists(1)[0], create_eval_lists.create_modelIPdur_lists(1)[1])]
 
 # Splitting dataset 10 times into disjunct training- and test-datasets (90% to 10%)
 rmse_list, mae_list, corr_list = [], [], []
 
-
 dataset = model_utilities.get_path_list('C:/Users/alexutza_a/Abschlussarbeit/DB_Verbmobil/Evaluation/Training')
-num_folds = 10
-subset_size = int(len(dataset)/num_folds)
-for i in range(num_folds):
-	t1 = time.time()
-	path_list_test = dataset[i*subset_size:][:subset_size]
-	path_list_training = dataset[:i*subset_size] + dataset[(i+1)*subset_size:]
 
-	# Taking the best model for further testing
-	evaluation_lists = [(create_eval_lists.create_modelWLIP_lists(1, path_list_test, path_list_training)[0], 
-		create_eval_lists.create_modelWLIP_lists(1, path_list_test, path_list_training)[1])]
+global_rmse, mae = performance_measures.calc_error_for_data(create_eval_lists.create_modelIPdur_lists(1)[0], create_eval_lists.create_modelIPdur_lists(1)[1])
+corrCoef = performance_measures.corrCoef(create_eval_lists.create_modelIPdur_lists(1)[0], create_eval_lists.create_modelIPdur_lists(1)[1])[0]
 
-	for i in range(len(evaluation_lists)):
-		global_rmse, mae = performance_measures.calc_error_for_data(evaluation_lists[i][0], evaluation_lists[i][1])
-		corrCoef = performance_measures.corrCoef(evaluation_lists[i][0], evaluation_lists[i][1])[0]
-	rmse_list.append(global_rmse)
-	mae_list.append(mae)
[…]

### 2017-04-06 py_files/fetch_data.py

@@ -1,47 +1,0 @@
-import os
-import glob
-import re
-import numpy as np
-from collections import defaultdict
-
-# Returns a dictionary of the official phoneme means for VM1+2
-def standard_value_means():
-	o_dict = defaultdict(float)
-	official_file = open("C:/Users/alexutza_a/Abschlussarbeit/DB_Verbmobil/Evaluation/Training/Basic_german_phone_list.txt")
-	for line in official_file:
-		if len(line.split()[0]) < 3:
-			o_dict[line.split()[0]] = int(round((float(line.split()[6])/0.0000625), 0))
-
-	#print(o_dict)
-	return o_dict
-
-# Returns a dictionary of phoneme occurences (keys) in the training data and their durations (values)
-def read_trainig_files():
-	t_dict = defaultdict(list)
-	os.chdir("C:/Users/alexutza_a/Abschlussarbeit/DB_Verbmobil/Evaluation/Training")
-	
-	#Iterate over the training files
-	for file in glob.glob("*.par"):
-		work_file = open(file)
-
-		for line in work_file:
-			if re.match("MAU", line):
-				t_dict[line.split()[4]].append(int(line.split()[2]))
-		work_file.close()
-	
-	# Remove breaks from the data
-	x = t_dict.pop("<p:>")
-
-	return t_dict
-
-# Calculates the mean of phoneme durations / phoneme and puts it at the beginning of the value lists
-# Returns the phoneme dict of the training data updated with the mean duration of each phoneme 
-# 	at the beginning of each value list
-def calculate_pho_mean(phon_dict):
-	for key in phon_dict.keys():
-		key_mean = np.mean(phon_dict[key])
-		phon_dict[key].insert(0, int(round((key_mean), 0)))
-	#print(phon_dict)
-	return phon_dict
-
-calculate_pho_mean(read_trainig_files())

### 2017-04-06 py_files/find_syl_position.py

@@ -189,11 +189,11 @@ def get_syl(p_index, syl_dict):
 #print(prepare_strings("Qan6", "thal", "Q''an.d6+t.h'alp"))
 #print(prepare_strings("fE6aI", "nma:6n", "f6.+Q'aIn.+ba:.r#+@n"))
 #print(prepare_strings("ve", "a:t6b@zu:x", "te.Qa:t6.b@.zu:x"))
-#print(compare_strings("an6", "thal", "an.d6t.halp"))	# anderthalb
-#print(compare_strings("fE6aI", "nma:6n", "f6.aIn.ba:.r@n"))	# vereinbaren
-#print(compare_strings("b@zICt", "IgN", "b@.zIC.tI.g@n"))	# besichtigen
-#print(compare_strings("naIn", "halpte:gIg@s", "aI.naIn.halp.tE:.gi.g@s")) # eineinhalbtägiges
-#print(compare_strings("ar", "baItstrEfn", "a6.baIts.trE.f@n")) # Arbeitstreffen
+print(compare_strings("an6thal", "an.d6t.halp"))	# anderthalb
+print(compare_strings("fE6aInma:6n", "f6.aIn.ba:.r@n"))	# vereinbaren
+print(compare_strings("b@zICtIgN", "b@.zIC.tI.g@n"))	# besichtigen
+print(compare_strings("naInhalpte:gIg@s", "aI.naIn.halp.tE:.gi.g@s")) # eineinhalbtägiges
+print(compare_strings("arbaItstrEfn", "a6.baIts.trE.f@n")) # Arbeitstreffen
 #print(compare_strings("tE6m", "i:nmE:sIC", "tE6.mi:n.mE:.sIC"))
 #print(compare_strings("neC", "stn", "n'E:Cs.t@n"))
 #print(compare_strings("vea:t6b@zu:x", "te.a:.t6.b@.zu:x"))

### 2017-04-06 py_files/get_cursor_pos.py

@@ -1,17 +1,0 @@
-import re
-
-def cursor_pos():
-	work_file = open("C:/Users/alexutza_a/Abschlussarbeit/DB_Verbmobil/Evaluation/Test/g002acn1_000_AAJ.par")
-	#work_file = open("C:/LP/ad_lesbiam.txt")
-	zeile = 0
-	p_zeile = []
-
-	for line in work_file:
-		zeile += 1
-		if re.match("MAU", line) and (int(line.split()[3]) == 9) and (str(line.split()[4]) == "n"):
-			p_zeile.append(zeile)
-			print(line)
-	work_file.close()
-	return p_zeile
-#cursor_pos()
-print(cursor_pos())

### 2017-04-06 py_files/get_sp_perDialog.py

@@ -1,24 +1,0 @@
-import model_utilities
-from collections import defaultdict
-
-speaker_list = open('C:/Users/alexutza_a/Abschlussarbeit/DB_Verbmobil/speaker_perDialog.txt', "w")
-SL = []
-speaker_list.close()
-for pfad in model_utilities.get_path_list('C:/Users/alexutza_a/Abschlussarbeit/DB_Verbmobil/verbmobil_par'):
-	#speaker_list.write(str(pfad)[-20:-16] + "\t" + str(pfad)[-7:-4] + "\n")
-	if (str(pfad)[-7:-4] not in SL) and (str(pfad)[-16] != "b"):
-		SL.append(str(pfad)[-7:-4])
-
-speaker_set = set(SL)
-print(len(speaker_set))
-
-speaker_dict = defaultdict(set)
-for pfad in model_utilities.get_path_list('C:/Users/alexutza_a/Abschlussarbeit/DB_Verbmobil/verbmobil_par'):
-	speaker_dict[str(pfad)[-7:-4]].add(str(pfad)[-20:-15])
-
-for speaker in speaker_dict.keys():
-	speaker_list = open('C:/Users/alexutza_a/Abschlussarbeit/DB_Verbmobil/speaker_perDialog.txt', "a")
-	speaker_list.write("\n" + speaker + "\t")
-	for el in speaker_dict[speaker]:
-		speaker_list.write(el + "\t")
-	speaker_list.close()

### 2017-04-06 py_files/model_ipdur.py

@@ -13,7 +13,7 @@ import model_utilities
 
 # Create list of filepaths to explore. This one uses entire data set, for exploration purposes.
 # path_list_training = verbmo_par_files('C:/Users/alexutza_a/Abschlussarbeit/DB_Verbmobil/verbmobil_par')
-path_list_training = model_utilities.get_path_list('C:/Users/alexutza_a/Abschlussarbeit/DB_Verbmobil/Evaluation/Training')
+path_list_training = model_utilities.get_path_list('C:/Users/alexutza_a/Abschlussarbeit/DB_Verbmobil/Evaluation/Training_a')
 
 valid_phonemes = ["a", "a~", "e", "E", "I", "i", "O", "o", "U", "u", "Y", "y", "9", "2", "a:", "a~:", "e:", "E:", "i:",
                  "o:", "u:", "y:", "2:", "OY", "aU", "aI", "@", "6", "z", "S", "Z", "C", "x", "N", "Q", "b", "d", "f", 

### 2017-04-06 py_files/new_SR.py

@@ -186,4 +186,11 @@ def get_word_SR(datei, word_nr):
 
 	return p_phon_WSR, n_phon_WSR
 
-#print(get_word_SR('C:/Users/alexutza_a/Abschlussarbeit/DB_Verbmobil/mod_dataset/g001acn1_000_AAJ_a.par', 1)[0])
+#SR_list = []
+#for el in range(13):
+#	SR_list.append(get_word_SR('C:/Users/alexutza_a/Abschlussarbeit/DB_Verbmobil/mod_dataset/g001acn1_035_AAJ_a.par', el)[0])
+#SR_list = [ (x/10)*2 for x in SR_list]
+
+#print(SR_list)
+
+#print(get_word_SR('C:/Users/alexutza_a/Abschlussarbeit/DB_Verbmobil/mod_dataset/g001acn1_017_AAJ_a.par', 1)[0])

### 2017-04-06 py_files/new_arff_file_writer.py

@@ -1,12 +1,13 @@
 import new_phon_dict
 import model_utilities
+import phon_list
 import time
 
 # Generates an almost ready for processing with weka text file. 
 
 ######################################################## STATIC DATA #############################################################
 # Dataset used must contain only modified par-files
-dataset_path = model_utilities.get_path_list('C:/Users/alexutza_a/Abschlussarbeit/DB_Verbmobil/Evaluation/Test_a')
+dataset_path = model_utilities.get_path_list('C:/Users/alexutza_a/Abschlussarbeit/DB_Verbmobil/Evaluation/Training_a')
 
 # List of phonemes approximated by our models. 
 # As there are also other phonetic items segmented in our database, we use this list to sort those out.
@@ -33,7 +34,7 @@ wtype_dict = {
 
 ######################################################## Writing file: attribute description #############################################################
 
-arff_file = open("C:/Users/alexutza_a/Abschlussarbeit/wekadateien/cart_test.txt", "w")
+arff_file = open("C:/Users/alexutza_a/Abschlussarbeit/wekadateien/cart_min_trainAKX.txt", "w")
 
 arff_file.write("% ARFF file for Verbmobil\n% \n@relation cart\n%\n% List of attributes:\n")
 arff_file.write("% Duration is expressed in miliseconds\n%\n")
@@ -72,28 +73,43 @@ arff_file.write("@data\n%\n")
 
 t1 = time.time()
 # Create dictionary to be written in the arff-file
-cart_dict = new_phon_dict.cart_dict(dataset_path)
-
+c_dict = new_phon_dict.cart_dict(dataset_path)
+#cart_l = phon_list.cart_list(dataset_path)
 t2 = time.time()
 print("Time to create dictionary: " + str(t2 - t1))
 
 # Combine all value lists of the dictionary into one huge list
-cart_list = []
-for phon in cart_dict.keys():
-	cart_list += cart_dict[phon]
-t3 = time.time()
-print("Time to build the list for writing: " + str(t3 - t2))
+#cart_list = []
+#for phon in cart_dict.keys():
+#	cart_list += cart_dict[phon]
+#t3 = time.time()
+#print("Time to build the list for writing: " + str(t3 - t2))
 
 time1 = time.time()
-print("Time to get list length: " + str(time1 - t3))
-
[…]

### 2017-04-06 py_files/old_dict.py

@@ -1,86 +1,0 @@
-import duration_dict
-from collections import defaultdict
-from statistics import *
-
-# Creates a dictionary of phonemes, which incorporates the duration-dictionary for each phoneme and adds stats
-# A much too complicated way of building the dictionary of phonemes out of a dictionary of durations.
-
-# List of phonemes to search for
-phoneme_list = ["OY", "aU", "aI"]
-# Get the dictionary of lengths
-dict_of_lengths = duration_dict.phon_dict(phoneme_list)
-
-
-# Returns a list of phoneme duration occurencies (duration_list_per_phoneme) 
-# 	and a list containing the values of all keys (duration_flat_list).
-# Source: already existent dictionary.
-def create_list_of_durations():
-	# Initialize variables
-	duration_list_per_phoneme = []
-	value_list = []
-	duration_flat_list = []
-
-	# Populate list with keys from duration dictionary
-	key_list = dict_of_lengths.keys()
-
-	# Iterate over the keys of the dictionary to get the value lists
-	for item in key_list:
-		value_list.extend(dict_of_lengths[item])	# Get value list of a specific key
-		number_of_key_occurences = int(len(value_list) / 7)		# Divide value list by number of different element-
-																# types (attributes) = no. of occurences of a specific duration
-		duration_list_per_phoneme.extend([item] * number_of_key_occurences) # list of all duration value occurencies
-		#duration_flat_list.append(item)
-		duration_flat_list.extend(value_list)
-		value_list.clear()  # Reset list, for new key-value pair
-	return duration_list_per_phoneme, duration_flat_list
-duration_list_per_phoneme, duration_flat_list = create_list_of_durations()
-
-
-# Writes speech rates to a text file, one column
-def write_speech_rate():
-	sp_file = open("C:/Users/alexutza_a/Abschlussarbeit/Datenanalyse_Verbmobil/SpeechRate_Histogram.txt", "w")
-	#sr_l=duration_flat_list[6::7]
-	for elem in duration_flat_list[6::7]:
-		sp_file.write(str(round(float(elem), 2)) + "\n")
-	sp_file.close()
-#write_speech_rate()
-
-
-
-# Replaces the speech rate values in the duration_flat_list with normalized values
-# Param flat_list = duration_flat_list; sp_list = list of not normalized speech rates; norm_sp_list = list of normalized speech rates
[…]

### 2017-04-06 py_files/pd_LSR.py

@@ -1,186 +1,0 @@
-import os
-import glob
-import re
-import numpy as np
-from collections import defaultdict
-import speech_rate
-
-# Returns a dictionary of phoneme occurences (keys) in the training data and their durations (values), and LSR (values)
-# Looks like {a: [1452, 0.8, 799, 0.5], b : [655, 0.5, 799, 0.45]...}
-def read_trainig_files():
-	training_dict = defaultdict(list)
-	os.chdir("C:/Users/alexutza_a/Abschlussarbeit/DB_Verbmobil/Evaluation/Training")
-	
-	#Iterate over the training files
-	for file in glob.glob("*.par"):
-		work_file = open(file)
-
-		for line in work_file:
-			if re.match("MAU", line):
-				training_dict[line.split()[4]].append(int(line.split()[2]))
-				word_duration, phon_count, syl_count = speech_rate.word_duration(file, int(line.split()[3]))
-				training_dict[line.split()[4]].append((word_duration/0.0000625)/phon_count)
-				#training_dict[line.split()[4]].append(speech_rate.local_speech_rate(file, int(line.split()[3])))
-
-		work_file.close()
-	
-	# Remove breaks from the data
-	x = training_dict.pop("<p:>")
-
-	return training_dict
-
-# A dictionary giving the values of Mean and SD in a list for each encountered phoneme.
-def phone_stats(training_dict):
-	stat_dict = defaultdict(list)
-
-	for phoneme in training_dict.keys():
-		stat_dict[phoneme].append(int(round(np.mean(training_dict[phoneme][::2]), 0)))
-		stat_dict[phoneme].append(int(round(np.std(training_dict[phoneme][::2]), 0)))
-
-	return stat_dict
-
-# Not used
-def mean_of_means(stat_dict):
-	m1 = 0
-
-	for p in stat_dict:
-		m1 += stat_dict[p][0]
-	mom = m1 / len(stat_dict.keys())
-	return mom 
-
-# Returns a list of phoneme occuring in the test files, followed by their respective durations
-# Looks like: ["a", 583, 0.5, "b", 12, 0.78, "a", 489, 0.12, ...]
-def read_testfiles():
-	compare_list = []
-	os.chdir("C:/Users/alexutza_a/Abschlussarbeit/DB_Verbmobil/Evaluation/Test")
-
-	#Iterate over the test files
-	for file in glob.glob("*.par"):
-		work_file = open(file)
-
[…]

### 2017-04-06 py_files/phon_counter.py

@@ -1,18 +1,0 @@
-import phon_dict
-import re
-
-file_list, path_list = phon_dict.verbmo_par_files()
-
-def phon_counter():
-	count = 0
-	for file in path_list:
-		datei = open(file)
-		for line in datei:
-				if re.match("MAU", line) and (line.split()[4] == "a"):
-					count += 1
-				elif re.match("PRO", line):
-					break
-	datei.close()
-	return count
-
-print(str(phon_counter()))

### 2017-04-06 py_files/phon_dict.py

@@ -1,190 +1,0 @@
-import os
-import glob
-import fnmatch
-import re
-from collections import defaultdict
-import stress
-import speech_rate
-import math
-import find_syl_position
-import model_utilities
-import time
-
-# Creates a dictionary of durations (of specific phoneme)
-dur_file = open("C:/Users/alexutza_a/Abschlussarbeit/word_durations.txt", "w")
-speech_rate_list = []
-# List of all german .par files in verbmobil, where the recording was done acn (main scenario dialogue 
-	#recorded means neckband microphone)
-path_list = model_utilities.get_path_list('C:/Users/alexutza_a/Abschlussarbeit/DB_Verbmobil/verbmobil_par')
-
-
-# Dictionary of the given phoneme instance length in ms (keys) in Verbmobil and its attributes (values in list)
-# Attributes are: Filename, Position in Speech (Start, Middle, End), Containing word, Stress, Overlapping, Speech rate
-def phon_dict(phoneme_list):
-	laut_schluessel = defaultdict(list)
-
-	# Iterate over the given list of phonemes, and
-	for phoneme in phoneme_list:
-		#iterate over all .par files in g016a and open/close one file at a time
-		#for file in glob.glob(r"C:/Users/alexutza_a/Abschlussarbeit/DB_Verbmobil/verbmobil_par/g016a/g*acn*.par"):
-		#Iterate over all .par files in the Verbmobil database and open/close one file at a time
-		# search in the open file for lines starting with "MAU", and containing given phoneme.
-		# Get the attributes for current phoneme, and place them as keys in the list of values.
-		for file in path_list:
-			datei = open(file)
-			filename = str(file)
-			#print(filename)
-			words = total_words(datei)
-			overlapping_list = overlapped_word(datei)
-			sp_pho, sp_w, speech_rate_msyl, speech_rate_ksyl = speech_rate.speech_rate(datei)
-			zeile = 0
-			
-		
-			for line in datei:
-				zeile += 1
-				if re.match("MAU", line) and (line.split()[4] == phoneme):
-					wort = int(line.split()[3])
-					# Attribute: phoneme, just for easier handling
-					laut_schluessel[phoneme].append(phoneme)
-					# Attribute: filename 
-					laut_schluessel[phoneme].append(filename[-20:])
-					# Attribute: Position in Speech: Start, Middle, End
-					laut_schluessel[phoneme].append(position_in_speechround(words, line))
[…]

### 2017-04-06 py_files/prepare_data.py

@@ -1,105 +1,0 @@
-import os
-import glob
-import re
-import numpy as np
-from collections import defaultdict
-
-# Returns a dictionary of the official phoneme means for VM1+2
-def standard_value_means():
-	o_dict = defaultdict(float)
-	omedian_dict = defaultdict(float)
-	official_file = open("C:/Users/alexutza_a/Abschlussarbeit/DB_Verbmobil/Evaluation/Training/Basic_german_phone_list.txt")
-	for line in official_file:
-		if len(line.split()[0]) < 3:
-			o_dict[line.split()[0]] = int(round((float(line.split()[6])/0.0000625), 0))
-			omedian_dict[line.split()[0]] = int(round((float(line.split()[9])/0.0000625), 0))
-	#print(o_dict)
-	return o_dict, omedian_dict
-o_dict, omedian_dict = standard_value_means()
-
-# Returns a dictionary of phoneme occurences (keys) in the training data and their durations (values)
-def read_trainig_files():
-	training_dict = defaultdict(list)
-	os.chdir("C:/Users/alexutza_a/Abschlussarbeit/DB_Verbmobil/Evaluation/Training")
-	
-	#Iterate over the training files
-	for file in glob.glob("*.par"):
-		work_file = open(file)
-
-		for line in work_file:
-			if re.match("MAU", line):
-				training_dict[line.split()[4]].append(int(line.split()[2]))
-		work_file.close()
-	
-	# Remove breaks from the data
-	x = training_dict.pop("<p:>")
-
-	return training_dict
-
-# Returns a list of phoneme occuring in the test files, followed by their respective durations
-# Looks like: ["a", 583, "b", 12, "a", 489, ...]
-def read_testfiles():
-	compare_list = []
-	os.chdir("C:/Users/alexutza_a/Abschlussarbeit/DB_Verbmobil/Evaluation/Test")
-
-	#Iterate over the test files
-	for file in glob.glob("*.par"):
-		work_file = open(file)
-
-		for line in work_file:
-			if re.match("MAU", line):
-				compare_list.append(str(line.split()[4]))
-				compare_list.append(int(line.split()[2]))
-		work_file.close()
-	
-	# Remove breaks from the data
-
-	# Get list of indexes for occurences of <p:>
-	pause_index = [i for i, val in enumerate(compare_list) if val == "<p:>"]
-	#print(pause_index)
-	pause_dur = [i + 1 for i in pause_index]
-	p_list = [x for y in zip (pause_index, pause_dur) for x in y]
-	#print(p_list)
[…]

### 2017-04-06 py_files/read_lex_file.py

@@ -1,10 +1,0 @@
-# Erstellt eine Datei mit 2 Spalten:
-# 1. orthografische Form des Wortes
-# 2. phonetische Form des Wortes, mit Silbentrennzeichen.
-
-lex_file = open("C:/Users/alexutza_a/Abschlussarbeit/DB_Verbmobil/verbmobil.lex")
-syl_file = open("C:/Users/alexutza_a/Abschlussarbeit/DB_Verbmobil/syl.lex", "w")
-
-for line in lex_file:
-	syl_file.write(line.split()[0] + "\t")
-	syl_file.write(line.split()[3] + "\n")

### 2017-04-06 py_files/speech_rate.py

@@ -1,168 +1,0 @@
-from __future__ import division
-import re
-import glob
-
-
-#speech_rate2 = 0
-vowel_list = ["Q", "a", "a~", "e", "E", "I", "i", "O", "o", "U", "u", "Y", "y", "9", "2", "@", "6", "a:", "a~:", "e:", "E:", "i:", "o:", "u:", "y:", "2:", "OY", "aU", "aI"]
-
-# Calculates:
-# - total speech duration of a turn in seconds, without breaks,
-# - total number of realized phonemes in a turn,
-# - total number of words in a turn
-# - total number of realized syllables in a turn,
-# 
-def speech_duration_pho(datei):
-	#datei = open(d)
-	sp_dur = 0
-	phoneme_count = 0
-	word_count = 0
-	mau_syl_count = 0
-	word = ""
-	kan_syl_count = 0
-
-	# Iterate over the lines starting with "KAN", which do not correspond to a pause.
-	for line in datei:
-		if re.match("KAN", line):
-			word_count += 1
-			word = str(line.split()[2])
-			# Eliminate diphthongs, so their syllables don't get double counted
-			for diph in ["aI", "aU", "OY"]:
-				if diph in word:
-					kan_syl_count += word.count(diph)
-					word = word.replace(diph, "")
-			# Count rest of vowels
-			for v in vowel_list[:-12]:
-				if v in word:
-					kan_syl_count += word.count(v)
-		# Stop search when KAN tier passed, so cursor doesn't reach EOF
-		elif re.match("SYN", line):
-			break
-
-	# Iterate over the lines starting with "MAU", which do not correspond to a pause.
-	for line in datei:
-		if re.match("MAU", line) and ("<p" not in line):
-			phoneme_count += 1 							# Get total no. of phonemes
-			sp_dur += int(line.split()[2])			# Get total duration without pauses
-			if str(line.split()[4]) in vowel_list:	# As syllable nuclei are always vowels, we count them to get the syllable count
-				mau_syl_count += 1
-
-	datei.seek(91)
-
-	duration = sp_dur * 0.0000625				# Transform total duration in seconds
-
-	return duration, phoneme_count, word_count, mau_syl_count, kan_syl_count
-
-
[…]

### 2017-04-06 py_files/stress.py

@@ -17,8 +17,9 @@ def stress_type(datei, word_number, phoneme, cursor):
 			#print(phoneme)
 		
 	mau_rank = p_rank_in_mau(work_file, word_no, phoneme, cursor)
-	stress_type = ""
-	word = ""
+	#print("MAU-rank: " + str(mau_rank))
+	#stress_type = ""
+	#word = ""
 	
 	for line in work_file:	
 		if re.match("KAN", line) and (int(line.split()[1]) == word_no):
@@ -136,6 +137,7 @@ def short_mono_stress(mau, kan, mau_rank, phoneme):
 		# If given phoneme has same count in both strings, 
 		#  then we can simply search for the needed occurence in the kan string, no other checks necessary
 		if count_m == count_k: 
+			#print("ok")
 			k_index = kan.find(phoneme)
 			stress_type = set_label(k_index, kan)
 

### 2017-04-06 py_files_old/Check_file.py

@@ -1,0 +1,14 @@
+import model_utilities
+import re
+
+dataset = model_utilities.get_path_list('C:/Users/alexutza_a/Abschlussarbeit/DB_Verbmobil/mod_dataset')
+file_error_list = []
+
+for datei in dataset:
+	w_f = open(datei)
+	for line in w_f:
+		if re.match("POS", line):
+			if "," in str(line.split()[1]):
+				file_error_list.append(str(datei)[-22:])
+	w_f.close()
+print(file_error_list)

### 2017-04-06 py_files_old/arff_file_builder.py

@@ -1,0 +1,88 @@
+import phon_dict
+import fileinput
+import re
+import os
+import sys
+import time
+
+# Generates an almost ready for processing with weka text file. 
+time1 = time.time()
+# List of phonemes to search for
+#phoneme_list = ["a", "a:", "E", "E:", "e", "e:", "O", "o", "o:", "@", "6", "9"] # needs 4,5 hrs
+#phoneme_list = ["p", "b", "t", "d", "k", "g"] # needs cca 3hrs (2hrs 45 min)
+#phoneme_list = ["f", "v", "s", "S", "z", "x", "h", "C", "Z"] # ca. 3hrs (3hrs 14 min)
+#phoneme_list = ["m", "n", "N", "l", "r", "j"] # ca. 3 hrs 40 min
+phoneme_list = ["@", "6"] # ca. 1hr 15 min
+# All diphthongs:
+#phoneme_list = ["OY", "aU", "aI"]
+# All long vowels:
+#phoneme_list = ["a:", "e:", "E:", "i:", "o:", "u:", "y:", "2:", "a~:"]
+# All vowels:
+#phoneme_list = ["a", "a~", "e", "E", "I", "i", "O", "o", "U", "u", "Y", "y", "9", "2", "a:", "a~:", "e:", "E:", "i:", "o:", "u:", "y:", "2:", "OY", "aU", "aI"]
+
+# Get the dictionary of phonemes (keys) and their attributes (values)
+phon_dict = phon_dict.phon_dict(phoneme_list)
+#dict_of_lengths = duration_dict.phon_dict(phoneme_list)
+
+# Put all dict items in one list
+dict_list = []
+for el in phoneme_list:
+	dict_list += phon_dict[el]
+
+# Get the labels for the attribute: word
+word_list = set(dict_list[3::13])
+
+
+#test_file = open("C:/Users/alexutza_a/Abschlussarbeit/sp_kurzes_a.txt", "w") -> gibt falsche Anzahl von speech rates
+arff_file = open("C:/Users/alexutza_a/Abschlussarbeit/wekadateien/schwa_usw", "w")
+
+arff_file.write("% ARFF file for Verbmobil\n% Phonemes: @, 6\n%\n@relation schwa\n\n% List of attributes:\n%\n")
+arff_file.write("% Duration is expressed in seconds\n")
+
+arff_file.write("@attribute phoneme { ") 
+for el in phoneme_list:
+	arff_file.write(el + ", ")
[…]

### 2017-04-06 py_files_old/calc_errors.py

@@ -1,0 +1,83 @@
+import performance_measures
+import numpy as np
+from collections import defaultdict
+import pd_LSR
+
+#phon_dict, test_dict, tm_dict = prepare_data.calculate_pho_mean(prepare_data.read_trainig_files())
+# A dictionary giving the values of Mean and SD in a list for each encountered phoneme.
+stat_dict = pd_LSR.phone_stats(pd_LSR.read_trainig_files())
+#print(test_dict)
+
+#testlist = prepare_data.read_testfiles()
+# Looks like: ["a", 583, 0.5, "b", 12, 0.78, "a", 489, 0.12, ...]
+testlist = pd_LSR.read_testfiles()
+#print(testlist)
+
+predictions = pd_LSR.create_prediction_list(testlist, stat_dict)
+actual = pd_LSR.read_testfiles()[1::3]
+
+
+#print(len(predictions))
+#print(prepare_data.read_testfiles())
+#print(len(actual))
+
+global_rmse, mae = performance_measures.calc_error_for_data(predictions, actual)
+corrCoef = performance_measures.corrCoef(predictions, actual)[0]
+print(global_rmse, mae, corrCoef)
+
+
+### Following is rmse calculation individually (per phoneme) ###
+
+
+
+# Set of occuring phonemes to iterate on when creating dictionary
+#set_phon = set(prepare_data.read_testfiles()[::2])
+set_phon = set(pd_LSR.read_testfiles()[::3])
+#print(set_phon)
+
+# A list of phonemes in occuring order in the actual and pred lists
+#a_phone_list = prepare_data.read_testfiles()[::2]
+a_phone_list = pd_LSR.read_testfiles()[::3]
+#print(a_phone_list)
+
+# Use list comprehension to interweave the phoneme list with the list of predicted durations
+#   so we can create the dictionary of predictions per phoneme
+pred_list = [x for y in zip (a_phone_list, predictions) for x in y]
+
+
+# Dictionary of actual values per phoneme
+a_dict = defaultdict(list)
+# Dictionary of predicted values per phoneme
+p_dict = defaultdict(list)
+
+
+# Populate dictionary of actual values
+for el in set_phon:
+	i = 0
+	for y in testlist:
+		if el == y:
+			a_dict[el].append(testlist[i+1])
+		i += 1
+#print(a_dict)
+# Populate dictionary of predicted values
+for el in set_phon:
+	j = 0
+	for z in pred_list:
[…]

### 2017-04-06 py_files_old/check_file_composition.py

@@ -1,0 +1,53 @@
+import model_utilities
+import re
+
+dataset_path = model_utilities.get_path_list('C:/Users/alexutza_a/Abschlussarbeit/DB_Verbmobil/verbmobil_par')
+
+dataset_list, pos_list, pro_list = [], [], []
+
+for datei in dataset_path:
+	d_name = str(datei)[-20:]
+	dataset_list.append(d_name)
+
+	work_file = open(datei)
+	for line in work_file:
+		if re.match("PRO", line):
+			pro_list.append(d_name)
+		if re.match("POS", line):
+			pos_list.append(d_name)
+			break
+	work_file.close()
+
+pro_set = set(pro_list)
+
+neg_pro_list = []
+for el in dataset_path:
+	if el[-20:] not in pro_set:
+		neg_pro_list.append(el)
+
+w_count_list = []
+for datei in neg_pro_list:
+	work_file = open(datei)
+	w_counter = 0
+	for line in work_file:
+		if re.match("ORT", line):
+			w_counter += 1
+	w_count_list.append(w_counter)
+	work_file.close()
+
+i = 0
+nok_npro_list, ok_pro_list = [], []
+for el in w_count_list:
+	if el > 7:
+		nok_npro_list.append(neg_pro_list[i])
+	else:
+		ok_pro_list.append(neg_pro_list[i])
+	i += 1
+
+nok_npro_set = set(nok_npro_list)
+
+#print(len(dataset_list))
+#print(len(pos_list))
+print(len(nok_npro_list))
+print(nok_npro_list[:5])
+#print(nok_npro_list[:15])

### 2017-04-06 py_files_old/create_eval_lists.py

@@ -1,0 +1,93 @@
+import re
+import model_utilities
+import model_wl
+import model_wl_ip
+import model_wl_ipdur
+import model_ipdur
+
+
+# List of paths to the files in our test data, to later iterate on
+#path_list_test = model_utilities.get_path_list('C:/Users/alexutza_a/Abschlussarbeit/DB_Verbmobil/Evaluation/Test')
+
+valid_phonemes = ["a", "a~", "e", "E", "I", "i", "O", "o", "U", "u", "Y", "y", "9", "2", "a:", "a~:", "e:", "E:", "i:",
+                 "o:", "u:", "y:", "2:", "OY", "aU", "aI", "@", "6", "z", "S", "Z", "C", "x", "N", "Q", "b", "d", "f", 
+                 "g", "h", "j", "k", "l", "m", "n", "p", "r", "s", "t", "v"]
+
+path_list_test = model_utilities.get_path_list('C:/Users/alexutza_a/Abschlussarbeit/DB_Verbmobil/Evaluation/Test_a')
+
+# Model type uses individual phonemes and their median, mean, or both in words having specific no. of phonemes
+# Actually creates the lists of predictions and actual values to evaluate model
+# @param model: 0, 1, or 2 meaning (median, mean, or (median + mean)/2)
+# Return: 2 lists of integers
+def create_modelIPdur_lists(model):
+	predictions_list, actuals_list = [], []
+	for datei in path_list_test:
+		work_file = open(datei)
+		for line in work_file:
+			# Restrain model to relevant phonemes
+			if re.match("MAU", line) and (str(line.split()[4]) in valid_phonemes):
+				predicted_pdur = model_ipdur.pdur_prediction_value(datei, int(line.split()[3]), str(line.split()[4]), model)
+				predictions_list.append(predicted_pdur)
+
+				actuals_list.append(int(line.split()[2]))
+		work_file.close()
+	#print(len(predictions_list))
+	#print(len(actuals_list))
+	return predictions_list, actuals_list
+
+# Model type uses individual phonemes and their median, mean, or both
+# Actually creates the lists of predictions and actual values to evaluate model
+# @param model: 0, 1, or 2 meaning (median, mean, or (median + mean)/2)
+# Return: 2 lists of integers
+def create_modelWLIP_lists(model, path_list_test, path_list_training):
+	predictions_list, actuals_list = [], []
+	for datei in path_list_test:
+		work_file = open(datei)
+		for line in work_file:
+			# Restrain model to relevant phonemes
[…]

### 2017-04-06 py_files_old/eval_model.py

@@ -1,0 +1,22 @@
+import performance_measures
+import numpy as np
+import create_eval_lists
+import model_utilities
+import time
+
+#### Running this file prints out the results of the performance evaluation of the differen models ####
+
+# Create the two lists of values to be compared, for each model a pair
+ # - 1st numeric parameter allows chosing the unit on which one evaluates the model (0 = median, 1 = mean, 2 = (median + mean)/2 )
+ # - 2nd numeric parameter should not be changed (it's just the indices of the elem in the result vector to be used)
+
+# Splitting dataset 10 times into disjunct training- and test-datasets (90% to 10%)
+rmse_list, mae_list, corr_list = [], [], []
+
+dataset = model_utilities.get_path_list('C:/Users/alexutza_a/Abschlussarbeit/DB_Verbmobil/Evaluation/Training')
+
+global_rmse, mae = performance_measures.calc_error_for_data(create_eval_lists.create_modelIPdur_lists(1)[0], create_eval_lists.create_modelIPdur_lists(1)[1])
+corrCoef = performance_measures.corrCoef(create_eval_lists.create_modelIPdur_lists(1)[0], create_eval_lists.create_modelIPdur_lists(1)[1])[0]
+
+print(global_rmse, mae)
+print(corrCoef)

### 2017-04-06 py_files_old/find_syl_position.py

@@ -1,0 +1,213 @@
+import re
+from collections import defaultdict
+
+diph = ["aI", "aU", "OY"]
+vowels = ["a", "A", "a~", "e", "E", "I", "i", "O", "o", "U", "u", "Y", "y", "9", "2", "@" , "6", "a:", "a~:", "e:", "E:", "i:", "o:", "u:", "y:", "2:"]
+consonants = ["z", "S", "Z", "C", "x", "N", "b", "d", "f", "g", "h", "j", "k", "l", "m", "n", "p", "r", "s", "t", "v"]
+
+# Returns the position of the syllable containing phoneme inside word:
+# - "w_initial" = 1st syllable, "w_final" = last syllable, "w_middle" = between 1st and last syll
+# - "one_s" = if it's a one syllable word
+# The counting of syllable is done on actually realized syllables! (not the canonical transcription)
+def syl_position(datei, word_no, cursor):
+	#print("Cursor: " + str(cursor))
+	syls_word = get_sylsplit_word(datei, word_no)
+	w1, w2 = mau_split(datei, word_no, cursor)
+	mau1, mau2, sylsplit_word = prepare_strings(w1, w2, syls_word)
+	#print("MAU1: " + mau1)
+	if sylsplit_word != "":
+		syl_dict = compare_strings(mau1+mau2, sylsplit_word)
+
+		#print(syl_dict.keys())
+		first_syl = min(syl_dict.keys())
+		last_syl = max(syl_dict.keys())
+		#print("Last syl: " + str(last_syl))
+
+		if first_syl != 0:
+			x = 0
+			for i in range (first_syl, last_syl+1):
+				syl_dict[x] = syl_dict.pop(i)
+				x += 1
+			last_syl = max(syl_dict.keys())
+		#print("New last syl: " + str(last_syl))
+		#print(syl_dict)
+		try:
+			syl_no = get_syl(mau1.index(mau1[-1]), syl_dict)-1
+		except:
+			syl_no = 0
+		else:
+			syl_no = 0
+		#print("Syl no: " + str(syl_no))
+	else:
+		first_syl, last_syl = 0, 0
+
+	inword_pos = ""
+	if first_syl != last_syl:
+		if syl_no == 0:
+			inword_pos = "w_initial"
+		elif syl_no >= last_syl:
+			inword_pos = "w_final"
+		else:
+			inword_pos = "w_middle"
+	else:
[…]

### 2017-04-06 py_files_old/get_sp_perDialog.py

@@ -1,0 +1,24 @@
+import model_utilities
+from collections import defaultdict
+
+speaker_list = open('C:/Users/alexutza_a/Abschlussarbeit/DB_Verbmobil/speaker_perDialog.txt', "w")
+SL = []
+speaker_list.close()
+for pfad in model_utilities.get_path_list('C:/Users/alexutza_a/Abschlussarbeit/DB_Verbmobil/verbmobil_par'):
+	#speaker_list.write(str(pfad)[-20:-16] + "\t" + str(pfad)[-7:-4] + "\n")
+	if (str(pfad)[-7:-4] not in SL) and (str(pfad)[-16] != "b"):
+		SL.append(str(pfad)[-7:-4])
+
+speaker_set = set(SL)
+print(len(speaker_set))
+
+speaker_dict = defaultdict(set)
+for pfad in model_utilities.get_path_list('C:/Users/alexutza_a/Abschlussarbeit/DB_Verbmobil/verbmobil_par'):
+	speaker_dict[str(pfad)[-7:-4]].add(str(pfad)[-20:-15])
+
+for speaker in speaker_dict.keys():
+	speaker_list = open('C:/Users/alexutza_a/Abschlussarbeit/DB_Verbmobil/speaker_perDialog.txt', "a")
+	speaker_list.write("\n" + speaker + "\t")
+	for el in speaker_dict[speaker]:
+		speaker_list.write(el + "\t")
+	speaker_list.close()

### 2017-04-06 py_files_old/model_ipdur.py

@@ -1,0 +1,57 @@
+import numpy as np
+from collections import defaultdict
+import model_utilities
+
+# This is a relatively static phoneme duration prediction model which only considers the median (or mean) value of a specific phoneme
+#
+# Predicted value = median or mean of phon_dur in provided (training) dataset
+
+###################################################### START ##################################################################
+
+############## STATIC DATA STRUCTURES (not dataset specific) ##################################
+
+
+# Create list of filepaths to explore. This one uses entire data set, for exploration purposes.
+# path_list_training = verbmo_par_files('C:/Users/alexutza_a/Abschlussarbeit/DB_Verbmobil/verbmobil_par')
+path_list_training = model_utilities.get_path_list('C:/Users/alexutza_a/Abschlussarbeit/DB_Verbmobil/Evaluation/Training_a')
+
+valid_phonemes = ["a", "a~", "e", "E", "I", "i", "O", "o", "U", "u", "Y", "y", "9", "2", "a:", "a~:", "e:", "E:", "i:",
+                 "o:", "u:", "y:", "2:", "OY", "aU", "aI", "@", "6", "z", "S", "Z", "C", "x", "N", "Q", "b", "d", "f", 
+                 "g", "h", "j", "k", "l", "m", "n", "p", "r", "s", "t", "v"]
+
+#################################### DYNAMIC DATA STRUCTURES (training data) ################################
+
+# Processed version of the phon_dur_dict (defined in module model_utilities): 
+# Returns: a flat dictionary. Looks like: 
+# {"a" : [median_1, mean_1, (m_1 + m_1) /2], "b" : [median_2, mean_2, (m_2 + m_2) /2], ...}
+def phon_dur_compressed_dict(phon_dur_dict):
+	phon_dur_compressed_dict = dict( (i, []) for i in valid_phonemes)
+
+	for phon in valid_phonemes:
+		phon_dur_compressed_dict[phon] = [round(np.median(phon_dur_dict[phon]), 3)]
+		phon_dur_compressed_dict[phon].append(round(np.mean(phon_dur_dict[phon]), 3))
+		phon_dur_compressed_dict[phon].append(round((np.median(phon_dur_dict[phon]) + np.mean(phon_dur_dict[phon]))/2, 3))
+	#print(round(np.median(phon_dur_dict["v"]), 3))
+	#print(round(np.median(phon_dur_dict["aU"]), 3))
[…]

### 2017-04-06 py_files_old/model_utilities.py

@@ -1,0 +1,99 @@
+import os
+import fnmatch
+import re
+from collections import defaultdict
+
+# This file contains help functions that are used across models
+
+############################### STATIC DATA #################################
+
+# List of phonemes approximated by our models. 
+# As there are also other phonetic items segmented in our database, we use this list to sort those out.
+# Example of excluded phonetic items: <p:>, <usb>
+valid_phonemes = ["a", "a~", "e", "E", "I", "i", "O", "o", "U", "u", "Y", "y", "9", "2", "a:", "a~:", "e:", "E:", "i:",
+                 "o:", "u:", "y:", "2:", "OY", "aU", "aI", "@", "6", "z", "S", "Z", "C", "x", "N", "Q", "b", "d", "f", 
+                 "g", "h", "j", "k", "l", "m", "n", "p", "r", "s", "t", "v"]
+
+vowel_list = ["Q", "a", "a~", "e", "E", "I", "i", "O", "o", "U", "u", "Y", "y", "9", "2", "@", "6", "a:", "a~:", "e:", "E:", "i:", "o:", "u:", "y:", "2:", "OY", "aU", "aI"]
+
+############################### FUNCTIONS #################################
+
+
+# Return: List of all german .par filepaths in provided data set, 
+#    where the recording was done acn (main scenario dialogue recorded means neckband microphone)
+# @param dataset_path: (string) path to the folder containing needed files (data)
+def get_path_list(dataset_path):
+	pattern = 'g*acn*.par'  # Pattern to be used for filtering filenames
+	path_list = []
+	for path, subfolder, filenames in os.walk(dataset_path):
+		for filename in fnmatch.filter(filenames, pattern):
+			path_list.append(os.path.join(path, filename))
+	return path_list
+
+
+# Creates a flat dictionary of phonemes and their durations in the given dataset
+# Return looks like: {"a" : [1599, 2348, 959, ...], "b" : [...], ...}
+def phon_dur_dict(path_list):
+    phon_dur_dict = dict( (i, []) for i in valid_phonemes)
+
+    for datei in path_list:
+        work_file = open(datei)
+        
+        for line in work_file:
+            if re.match("MAU", line) and (str(line.split()[4]) in valid_phonemes):
+                phon_dur_dict[str(line.split()[4])].append(int(line.split()[2]))
+        work_file.close()
+    #print(l)
[…]

### 2017-04-06 py_files_old/model_wl.py

@@ -1,0 +1,143 @@
+import re
+import numpy as np
+from collections import defaultdict
+import model_utilities
+
+# This is a phoneme duration prediction model for varying word duration.
+# It supposes we know the target word duration, and is based only on that feature.
+#
+# Before predicting phoneme duration, the phoneme is classified in one of the following categories: 
+#		diphthong, long_vowel, short_vowel, etc -> wehere the class members don't correspond completely to the ones listed in the grammar books.
+# Information from the training data is used to compute the steak each class member takes in a n-phoneme-long word. (? or maybe better if this is also resumed they way pred-values are ?)
+#
+# Predicted value = steak*w_dur/class-member-count-in-word
+
+###################################################### START ##################################################################
+
+
+############## STATIC DATA STRUCTURES (not dataset specific) ##################################
+
+
+# Create list of filepaths to explore. This one uses entire data set, for exploration purposes.
+# path_list_training = verbmo_par_files('C:/Users/alexutza_a/Abschlussarbeit/DB_Verbmobil/verbmobil_par')
+path_list_training = model_utilities.get_path_list('C:/Users/alexutza_a/Abschlussarbeit/DB_Verbmobil/Evaluation/Training')
+
+# Classification of phonemes used by this model
+phon_class_dict = {"diphthong" : ["aI", "aU", "OY"], "long_vowels" : ["a:", "E:", "e:", "i:", "2:", "@", "m", "k", "N"], 
+					"short_vowels" : ["a", "u:", "o:", "e", "O", "E", "C", "6", "U", "f", "y", "o", "S", "j", "y:"], 
+					"cons_allg" : ["x", "h", "l", "n", "I", "9", "z", "s", "Y", "v", "t"], "short_cons" : ["p", "b", "d", "g"], 
+					"others" : ["Q", "r"]}
+
+valid_phonemes = ["a", "a~", "e", "E", "I", "i", "O", "o", "U", "u", "Y", "y", "9", "2", "a:", "a~:", "e:", "E:", "i:",
+                 "o:", "u:", "y:", "2:", "OY", "aU", "aI", "@", "6", "z", "S", "Z", "C", "x", "N", "Q", "b", "d", "f", 
+                 "g", "h", "j", "k", "l", "m", "n", "p", "r", "s", "t", "v"]
+
[…]

### 2017-04-06 py_files_old/model_wl_ip.py

@@ -1,0 +1,100 @@
+import re
+import numpy as np
+from collections import defaultdict
+import model_utilities
+
+# This is a phoneme duration prediction model for varying word duration.
+# It supposes we know the target word duration, and is based only on that feature.
+#
+# Information from the training data is used to compute the steak each phoneme takes in a n-phoneme-long word.
+#
+# Predicted value = steak*w_dur/class-member-count-in-word
+
+###################################################### START ##################################################################
+
+############## STATIC DATA STRUCTURES (not dataset specific) ##################################
+
+
+# Create list of filepaths to explore. This one uses entire data set, for exploration purposes.
+# path_list_training = verbmo_par_files('C:/Users/alexutza_a/Abschlussarbeit/DB_Verbmobil/verbmobil_par')
+#path_list_training = model_utilities.get_path_list('C:/Users/alexutza_a/Abschlussarbeit/DB_Verbmobil/Evaluation/Training')
+
+valid_phonemes = ["a", "a~", "e", "E", "I", "i", "O", "o", "U", "u", "Y", "y", "9", "2", "a:", "a~:", "e:", "E:", "i:",
+                 "o:", "u:", "y:", "2:", "OY", "aU", "aI", "@", "6", "z", "S", "Z", "C", "x", "N", "Q", "b", "d", "f", 
+                 "g", "h", "j", "k", "l", "m", "n", "p", "r", "s", "t", "v"]
+
+#################################### DYNAMIC DATA STRUCTURES (training data) ################################
+
+# Processed version of the phon_wordleng_dict (defined in module model_utilities): 
+#    - the value-lists of the inner dict include only the phoneme proportions for the given word length (inner key)
+# Returns: a dictionary. Looks like: 
+# {"a" : { 1 : [median_a1, mean_a1, (m_a1 + m_a1) /2], 2 : [median_a2, mean_a2, (m_a2 + m_a2) /2], ...}, 
+#  "b" : { 1 : [...], ...}, ...}
+def phon_wl_compressed_dict(phon_wordleng_dict):
+	phon_wl_compressed_dict = dict( (i, defaultdict(list)) for i in phon_wordleng_dict.keys() )
+	pho_key_list = [key for key, val in phon_wordleng_dict.items()]
+
+	for phon in pho_key_list:
[…]

### 2017-04-06 py_files_old/new_SR.py

@@ -1,0 +1,196 @@
+import re
+import model_utilities
+from collections import defaultdict
+
+######################################################## STATIC DATA #############################################################
+
+# List of phonemes approximated by our models. 
+# As there are also other phonetic items segmented in our database, we use this list to sort those out.
+# Example of excluded phonetic items: <p:>, <usb>
+valid_phonemes = ["a", "a~", "e", "E", "I", "i", "O", "o", "U", "u", "Y", "y", "9", "2", "a:", "a~:", "e:", "E:", "i:",
+					"o:", "u:", "y:", "2:", "OY", "aU", "aI", "@", "6", "z", "S", "Z", "C", "x", "N", "Q", "b", "d", "f", 
+					"g", "h", "j", "k", "l", "m", "n", "p", "r", "s", "t", "v"]
+
+vowel_list = ["a", "a~", "e", "E", "I", "i", "O", "o", "U", "u", "Y", "y", "9", "2", "a:", "a~:", "e:", "E:", "i:",
+				"o:", "u:", "y:", "2:", "OY", "aU", "aI", "@", "6"]
+
+######################################################## Help Functions #############################################################
+
+# Get a list of word-numbers to split the turn into phrases (before each number)
+def get_split_list(datei):
+	work_file = open(datei)
+	split_list = []
+	word_list = []
+	a = 0
+	# Variable to store the word number of the previous line
+	memo_nr = 0
+
+	# Searching MAU-tier for pauses inside turn
+	for line in work_file:
+		if re.match("MAU", line):
+			word_list.append(int(line.split()[3]))
+			if (memo_nr == -1) and (int(line.split()[3]) > 0):
+				split_list.append(int(line.split()[3]))
+			memo_nr = int(line.split()[3])
+	
+		# Searching PRO-tier for discourse split markings
+		if re.match("PRO", line):
+			if ";" in str(line.split()[1]):
+				i = str(line.split()[1]).find(";")+1
+				split_w = int(str(line.split()[1])[i:])
+				split_list.append(split_w)
+			else:
+				split_list.append(int(line.split()[1]))
+	
+	if -1 in word_list:
+		word_list = [x for x in word_list if x != -1]
+
[…]

### 2017-04-06 py_files_old/new_arff_file_writer.py

@@ -1,0 +1,118 @@
+import new_phon_dict
+import model_utilities
+import phon_list
+import time
+
+# Generates an almost ready for processing with weka text file. 
+
+######################################################## STATIC DATA #############################################################
+# Dataset used must contain only modified par-files
+dataset_path = model_utilities.get_path_list('C:/Users/alexutza_a/Abschlussarbeit/DB_Verbmobil/Evaluation/Training_a')
+
+# List of phonemes approximated by our models. 
+# As there are also other phonetic items segmented in our database, we use this list to sort those out.
+# Example of excluded phonetic items: <p:>, <usb>
+valid_phonemes = ["a", "a~", "e", "E", "I", "i", "O", "o", "U", "u", "Y", "y", "9", "2", "a:", "a~:", "e:", "E:", "i:",
+					"o:", "u:", "y:", "2:", "OY", "aU", "aI", "@", "6", "z", "S", "Z", "C", "x", "N", "Q", "b", "d", "f", 
+					"g", "h", "j", "k", "l", "m", "n", "p", "r", "s", "t", "v"]
+
+vowel_list = ["a", "a~", "e", "E", "I", "i", "O", "o", "U", "u", "Y", "y", "9", "2", "a:", "a~:", "e:", "E:", "i:",
+				"o:", "u:", "y:", "2:", "OY", "aU", "aI", "@", "6"]
+
+ctype_dict = {
+"plosive" : ["b", "d", "p", "t", "k", "g"], "fricative" : ["f", "v", "s", "S", "z", "Z", "x", "h", "C"],
+"nasal" : ["m", "n", "N"], "lateral" : ["l"], "japproximant" : ["j"], "other" : [ "Q", "r"]
+}
+
+wtype_dict = {
+"content_words" : ["ADJA", "ADJD", "ADV", "BS", "CARD", "FM", "NN", "PTKNEG", "TRUNC", "VVFIN", "VVIMP", "VVINF", "VVIZU", "VVPP", "VAIMP", "VMPP"],
+"function_words" : ["APPR", "APPRART", "APPO", "APZR", "ART", "KOUI", "KOUS", "KON", "KOKOM", "PDS", "PDAT", "PIS", "PIAT", "PIDAT", "PPER", "PPOSS", "PPOSAT",
[…]

### 2017-04-06 py_files_old/new_phon_dict.py

@@ -1,0 +1,287 @@
+import re
+import model_utilities
+import new_SR
+from collections import defaultdict
+
+######################################################## STATIC DATA #############################################################
+
+# Dataset used must contain only modified par-files
+dataset_path = model_utilities.get_path_list('C:/Users/alexutza_a/Abschlussarbeit/DB_Verbmobil/Evaluation/my_test')
+
+# List of phonemes approximated by our models. 
+# As there are also other phonetic items segmented in our database, we use this list to sort those out.
+# Example of excluded phonetic items: <p:>, <usb>
+valid_phonemes = ["a", "a~", "e", "E", "I", "i", "O", "o", "U", "u", "Y", "y", "9", "2", "a:", "a~:", "e:", "E:", "i:",
+					"o:", "u:", "y:", "2:", "OY", "aU", "aI", "@", "6", "z", "S", "Z", "C", "x", "N", "Q", "b", "d", "f", 
+					"g", "h", "j", "k", "l", "m", "n", "p", "r", "s", "t", "v"]
+
+vowel_list = ["a", "a~", "e", "E", "I", "i", "O", "o", "U", "u", "Y", "y", "9", "2", "a:", "a~:", "e:", "E:", "i:",
+				"o:", "u:", "y:", "2:", "OY", "aU", "aI", "@", "6"]
+
+ctype_dict = {
+"plosive" : ["b", "d", "p", "t", "k", "g"], "fricative" : ["f", "v", "s", "S", "z", "Z", "x", "h", "C"],
+"nasal" : ["m", "n", "N"], "lateral" : ["l"], "japproximant" : ["j"], "other" : [ "Q", "r"]
+}
+
+wtype_dict = {
+"content_words" : ["ADJA", "ADJD", "ADV", "BS", "CARD", "FM", "NN", "PTKNEG", "TRUNC", "VVFIN", "VVIMP", "VVINF", "VVIZU", "VVPP", "VAIMP", "VMPP"],
+"function_words" : ["APPR", "APPRART", "APPO", "APZR", "ART", "KOUI", "KOUS", "KON", "KOKOM", "PDS", "PDAT", "PIS", "PIAT", "PIDAT", "PPER", "PPOSS", "PPOSAT",
[…]

### 2017-04-06 py_files_old/old_dict.py

@@ -1,0 +1,86 @@
+import duration_dict
+from collections import defaultdict
+from statistics import *
+
+# Creates a dictionary of phonemes, which incorporates the duration-dictionary for each phoneme and adds stats
+# A much too complicated way of building the dictionary of phonemes out of a dictionary of durations.
+
+# List of phonemes to search for
+phoneme_list = ["OY", "aU", "aI"]
+# Get the dictionary of lengths
+dict_of_lengths = duration_dict.phon_dict(phoneme_list)
+
+
+# Returns a list of phoneme duration occurencies (duration_list_per_phoneme) 
+# 	and a list containing the values of all keys (duration_flat_list).
+# Source: already existent dictionary.
+def create_list_of_durations():
+	# Initialize variables
+	duration_list_per_phoneme = []
+	value_list = []
+	duration_flat_list = []
+
+	# Populate list with keys from duration dictionary
+	key_list = dict_of_lengths.keys()
+
+	# Iterate over the keys of the dictionary to get the value lists
+	for item in key_list:
+		value_list.extend(dict_of_lengths[item])	# Get value list of a specific key
+		number_of_key_occurences = int(len(value_list) / 7)		# Divide value list by number of different element-
+																# types (attributes) = no. of occurences of a specific duration
+		duration_list_per_phoneme.extend([item] * number_of_key_occurences) # list of all duration value occurencies
+		#duration_flat_list.append(item)
+		duration_flat_list.extend(value_list)
+		value_list.clear()  # Reset list, for new key-value pair
+	return duration_list_per_phoneme, duration_flat_list
+duration_list_per_phoneme, duration_flat_list = create_list_of_durations()
+
+
+# Writes speech rates to a text file, one column
+def write_speech_rate():
+	sp_file = open("C:/Users/alexutza_a/Abschlussarbeit/Datenanalyse_Verbmobil/SpeechRate_Histogram.txt", "w")
+	#sr_l=duration_flat_list[6::7]
+	for elem in duration_flat_list[6::7]:
+		sp_file.write(str(round(float(elem), 2)) + "\n")
+	sp_file.close()
+#write_speech_rate()
+
+
+
+# Replaces the speech rate values in the duration_flat_list with normalized values
+# Param flat_list = duration_flat_list; sp_list = list of not normalized speech rates; norm_sp_list = list of normalized speech rates
[…]

### 2017-04-06 py_files_old/pd_LSR.py

@@ -1,0 +1,186 @@
+import os
+import glob
+import re
+import numpy as np
+from collections import defaultdict
+import speech_rate
+
+# Returns a dictionary of phoneme occurences (keys) in the training data and their durations (values), and LSR (values)
+# Looks like {a: [1452, 0.8, 799, 0.5], b : [655, 0.5, 799, 0.45]...}
+def read_trainig_files():
+	training_dict = defaultdict(list)
+	os.chdir("C:/Users/alexutza_a/Abschlussarbeit/DB_Verbmobil/Evaluation/Training")
+	
+	#Iterate over the training files
+	for file in glob.glob("*.par"):
+		work_file = open(file)
+
+		for line in work_file:
+			if re.match("MAU", line):
+				training_dict[line.split()[4]].append(int(line.split()[2]))
+				word_duration, phon_count, syl_count = speech_rate.word_duration(file, int(line.split()[3]))
+				training_dict[line.split()[4]].append((word_duration/0.0000625)/phon_count)
+				#training_dict[line.split()[4]].append(speech_rate.local_speech_rate(file, int(line.split()[3])))
+
+		work_file.close()
+	
+	# Remove breaks from the data
+	x = training_dict.pop("<p:>")
+
+	return training_dict
+
+# A dictionary giving the values of Mean and SD in a list for each encountered phoneme.
+def phone_stats(training_dict):
+	stat_dict = defaultdict(list)
+
+	for phoneme in training_dict.keys():
+		stat_dict[phoneme].append(int(round(np.mean(training_dict[phoneme][::2]), 0)))
+		stat_dict[phoneme].append(int(round(np.std(training_dict[phoneme][::2]), 0)))
+
+	return stat_dict
+
+# Not used
+def mean_of_means(stat_dict):
+	m1 = 0
+
+	for p in stat_dict:
+		m1 += stat_dict[p][0]
+	mom = m1 / len(stat_dict.keys())
+	return mom 
+
+# Returns a list of phoneme occuring in the test files, followed by their respective durations
+# Looks like: ["a", 583, 0.5, "b", 12, 0.78, "a", 489, 0.12, ...]
+def read_testfiles():
+	compare_list = []
+	os.chdir("C:/Users/alexutza_a/Abschlussarbeit/DB_Verbmobil/Evaluation/Test")
+
+	#Iterate over the test files
+	for file in glob.glob("*.par"):
+		work_file = open(file)
+
[…]

### 2017-04-06 py_files_old/performance_measures.py

@@ -1,0 +1,55 @@
+import numpy as np
+from scipy.stats.stats import pearsonr
+
+#predictions = [0.06, 0.06, 0.06, 0.06]
+#actual_values = [0.04, 0.06, 0.02, 0.06]
+
+# Returns the list of predicted values, for the training set
+def get_predictions(predictions_dict):
+	predictions = []
+	for key in predictions_dict.keys():
+		for i in range(1, len(predictions_dict[key])):
+			predictions.append(predictions_dict[key][0])
+	#	if key in vowel_list:
+	#		for i in range(1, len(durations_dict[key])):
+	#			predictions.append(1149)
+	#	else:
+	#		for i in range(1, len(durations_dict[key])):
+	#			predictions.append(durations_dict[key][0])
+	#print(durations_dict)
+	#print(predictions)
+
+	return predictions
+
+# Returns the list of actual values, for the training set
+def get_observed(observed_dict):
+	observed = []
+
+	for key in observed_dict.keys():
+		observed += observed_dict[key]
+	#print(observed)
+	return observed
+# Need
+def RMSE(prediction, actual):
+	return np.sqrt(((prediction - actual) ** 2).mean())
+
+
+# Need
+def MAE(prediction, actual):
+	return np.absolute(prediction - actual).mean()
+
+def corrCoef(prediction, actual):
+	return pearsonr(np.array(prediction), np.array(actual))
+
+#rmse_val = rmse(np.array(predictions), np.array(actual_values))
+#mae_val = mae(np.array(predictions), np.array(actual_values))
+
+# Need
+# Compares the two lists via rmse and mae.
+def calc_error_for_data(predictions, actuals):
+	rmse = RMSE(np.array(predictions), np.array(actuals))
+	rmse *= 0.0000625
+	mae = MAE(np.array(predictions), np.array(actuals))
+	mae *= 0.0000625
+
+	return rmse, mae

### 2017-04-06 py_files_old/phon_counter.py

@@ -1,0 +1,40 @@
+#import new_phon_dict
+import model_utilities
+import re
+import time
+import numpy as np
+
+dataset_path = model_utilities.get_path_list('C:/Users/alexutza_a/Abschlussarbeit/DB_Verbmobil/mod_dataset')
+#pcounter_file = open("C:/Users/alexutza_a/Abschlussarbeit/DB_Verbmobil/pcounter_file", "w")
+
+valid_phonemes = ["a", "a~", "e", "E", "I", "i", "O", "o", "U", "u", "Y", "y", "9", "2", "a:", "a~:", "e:", "E:", "i:",
+					"o:", "u:", "y:", "2:", "OY", "aU", "aI", "@", "6", "z", "S", "Z", "C", "x", "N", "Q", "b", "d", "f", 
+					"g", "h", "j", "k", "l", "m", "n", "p", "r", "s", "t", "v"]
+
+vowel_list = ["a", "a~", "e", "E", "I", "i", "O", "o", "U", "u", "Y", "y", "9", "2", "a:", "a~:", "e:", "E:", "i:",
+				"o:", "u:", "y:", "2:", "OY", "aU", "aI", "@", "6"]
+
+def phon_counter(dataset_path):
+	p_counter, v_counter = 0, 0
+	p_dur_list = []
+	for file in dataset_path:
+		datei = open(file)
+		for line in datei:
+				if re.match("MAU", line) and (line.split()[4] in valid_phonemes):
+					p_counter += 1
+					p_dur_list += int(line.split()[2])
+					if (line.split()[4] in vowel_list):
+						v_counter += 1
+				elif re.match("PRO", line):
+					break
+		datei.close()
+	return p_counter, v_counter, p_dur_list
+
+t1 = time.time()
+p_counter, v_counter = phon_counter(dataset_path)
+t2 = time.time()
+print(t2 - t1)
+#pcounter_file.close()
+
+print("Total phonemes: " + str(p_counter))
+print("Total vowels: " + str(v_counter))

### 2017-04-06 py_files_old/phon_dict.py

@@ -1,0 +1,190 @@
+import os
+import glob
+import fnmatch
+import re
+from collections import defaultdict
+import stress
+import speech_rate
+import math
+import find_syl_position
+import model_utilities
+import time
+
+# Creates a dictionary of durations (of specific phoneme)
+dur_file = open("C:/Users/alexutza_a/Abschlussarbeit/word_durations.txt", "w")
+speech_rate_list = []
+# List of all german .par files in verbmobil, where the recording was done acn (main scenario dialogue 
+	#recorded means neckband microphone)
+path_list = model_utilities.get_path_list('C:/Users/alexutza_a/Abschlussarbeit/DB_Verbmobil/verbmobil_par')
+
+
+# Dictionary of the given phoneme instance length in ms (keys) in Verbmobil and its attributes (values in list)
+# Attributes are: Filename, Position in Speech (Start, Middle, End), Containing word, Stress, Overlapping, Speech rate
+def phon_dict(phoneme_list):
+	laut_schluessel = defaultdict(list)
+
+	# Iterate over the given list of phonemes, and
+	for phoneme in phoneme_list:
+		#iterate over all .par files in g016a and open/close one file at a time
+		#for file in glob.glob(r"C:/Users/alexutza_a/Abschlussarbeit/DB_Verbmobil/verbmobil_par/g016a/g*acn*.par"):
+		#Iterate over all .par files in the Verbmobil database and open/close one file at a time
+		# search in the open file for lines starting with "MAU", and containing given phoneme.
+		# Get the attributes for current phoneme, and place them as keys in the list of values.
+		for file in path_list:
+			datei = open(file)
+			filename = str(file)
+			#print(filename)
+			words = total_words(datei)
+			overlapping_list = overlapped_word(datei)
+			sp_pho, sp_w, speech_rate_msyl, speech_rate_ksyl = speech_rate.speech_rate(datei)
+			zeile = 0
+			
+		
+			for line in datei:
+				zeile += 1
+				if re.match("MAU", line) and (line.split()[4] == phoneme):
+					wort = int(line.split()[3])
+					# Attribute: phoneme, just for easier handling
+					laut_schluessel[phoneme].append(phoneme)
+					# Attribute: filename 
+					laut_schluessel[phoneme].append(filename[-20:])
+					# Attribute: Position in Speech: Start, Middle, End
+					laut_schluessel[phoneme].append(position_in_speechround(words, line))
[…]

### 2017-04-06 py_files_old/phon_list.py

@@ -1,0 +1,286 @@
+import re
+import model_utilities
+import new_SR
+
+######################################################## STATIC DATA #############################################################
+
+# Dataset used must contain only modified par-files
+dataset_path = model_utilities.get_path_list('C:/Users/alexutza_a/Abschlussarbeit/DB_Verbmobil/Evaluation/my_test')
+
+# List of phonemes approximated by our models. 
+# As there are also other phonetic items segmented in our database, we use this list to sort those out.
+# Example of excluded phonetic items: <p:>, <usb>
+valid_phonemes = ["a", "a~", "e", "E", "I", "i", "O", "o", "U", "u", "Y", "y", "9", "2", "a:", "a~:", "e:", "E:", "i:",
+					"o:", "u:", "y:", "2:", "OY", "aU", "aI", "@", "6", "z", "S", "Z", "C", "x", "N", "Q", "b", "d", "f", 
+					"g", "h", "j", "k", "l", "m", "n", "p", "r", "s", "t", "v"]
+
+vowel_list = ["a", "a~", "e", "E", "I", "i", "O", "o", "U", "u", "Y", "y", "9", "2", "a:", "a~:", "e:", "E:", "i:",
+				"o:", "u:", "y:", "2:", "OY", "aU", "aI", "@", "6"]
+
+ctype_dict = {
+"plosive" : ["b", "d", "p", "t", "k", "g"], "fricative" : ["f", "v", "s", "S", "z", "Z", "x", "h", "C"],
+"nasal" : ["m", "n", "N"], "lateral" : ["l"], "japproximant" : ["j"], "other" : [ "Q", "r"]
+}
+
+wtype_dict = {
+"content_words" : ["ADJA", "ADJD", "ADV", "BS", "CARD", "FM", "NN", "PTKNEG", "TRUNC", "VVFIN", "VVIMP", "VVINF", "VVIZU", "VVPP", "VAIMP", "VMPP"],
+"function_words" : ["APPR", "APPRART", "APPO", "APZR", "ART", "KOUI", "KOUS", "KON", "KOKOM", "PDS", "PDAT", "PIS", "PIAT", "PIDAT", "PPER", "PPOSS", "PPOSAT",
[…]

### 2017-04-06 py_files_old/prepare_data.py

@@ -1,0 +1,105 @@
+import os
+import glob
+import re
+import numpy as np
+from collections import defaultdict
+
+# Returns a dictionary of the official phoneme means for VM1+2
+def standard_value_means():
+	o_dict = defaultdict(float)
+	omedian_dict = defaultdict(float)
+	official_file = open("C:/Users/alexutza_a/Abschlussarbeit/DB_Verbmobil/Evaluation/Training/Basic_german_phone_list.txt")
+	for line in official_file:
+		if len(line.split()[0]) < 3:
+			o_dict[line.split()[0]] = int(round((float(line.split()[6])/0.0000625), 0))
+			omedian_dict[line.split()[0]] = int(round((float(line.split()[9])/0.0000625), 0))
+	#print(o_dict)
+	return o_dict, omedian_dict
+o_dict, omedian_dict = standard_value_means()
+
+# Returns a dictionary of phoneme occurences (keys) in the training data and their durations (values)
+def read_trainig_files():
+	training_dict = defaultdict(list)
+	os.chdir("C:/Users/alexutza_a/Abschlussarbeit/DB_Verbmobil/Evaluation/Training")
+	
+	#Iterate over the training files
+	for file in glob.glob("*.par"):
+		work_file = open(file)
+
+		for line in work_file:
+			if re.match("MAU", line):
+				training_dict[line.split()[4]].append(int(line.split()[2]))
+		work_file.close()
+	
+	# Remove breaks from the data
+	x = training_dict.pop("<p:>")
+
+	return training_dict
+
+# Returns a list of phoneme occuring in the test files, followed by their respective durations
+# Looks like: ["a", 583, "b", 12, "a", 489, ...]
+def read_testfiles():
+	compare_list = []
+	os.chdir("C:/Users/alexutza_a/Abschlussarbeit/DB_Verbmobil/Evaluation/Test")
+
+	#Iterate over the test files
+	for file in glob.glob("*.par"):
+		work_file = open(file)
+
+		for line in work_file:
+			if re.match("MAU", line):
+				compare_list.append(str(line.split()[4]))
+				compare_list.append(int(line.split()[2]))
+		work_file.close()
+	
+	# Remove breaks from the data
+
+	# Get list of indexes for occurences of <p:>
+	pause_index = [i for i, val in enumerate(compare_list) if val == "<p:>"]
+	#print(pause_index)
+	pause_dur = [i + 1 for i in pause_index]
+	p_list = [x for y in zip (pause_index, pause_dur) for x in y]
+	#print(p_list)
[…]

### 2017-04-06 py_files_old/read_lex_file.py

@@ -1,0 +1,10 @@
+# Erstellt eine Datei mit 2 Spalten:
+# 1. orthografische Form des Wortes
+# 2. phonetische Form des Wortes, mit Silbentrennzeichen.
+
+lex_file = open("C:/Users/alexutza_a/Abschlussarbeit/DB_Verbmobil/verbmobil.lex")
+syl_file = open("C:/Users/alexutza_a/Abschlussarbeit/DB_Verbmobil/syl.lex", "w")
+
+for line in lex_file:
+	syl_file.write(line.split()[0] + "\t")
+	syl_file.write(line.split()[3] + "\n")

### 2017-04-06 py_files_old/speech_rate.py

@@ -1,0 +1,168 @@
+from __future__ import division
+import re
+import glob
+
+
+#speech_rate2 = 0
+vowel_list = ["Q", "a", "a~", "e", "E", "I", "i", "O", "o", "U", "u", "Y", "y", "9", "2", "@", "6", "a:", "a~:", "e:", "E:", "i:", "o:", "u:", "y:", "2:", "OY", "aU", "aI"]
+
+# Calculates:
+# - total speech duration of a turn in seconds, without breaks,
+# - total number of realized phonemes in a turn,
+# - total number of words in a turn
+# - total number of realized syllables in a turn,
+# 
+def speech_duration_pho(datei):
+	#datei = open(d)
+	sp_dur = 0
+	phoneme_count = 0
+	word_count = 0
+	mau_syl_count = 0
+	word = ""
+	kan_syl_count = 0
+
+	# Iterate over the lines starting with "KAN", which do not correspond to a pause.
+	for line in datei:
+		if re.match("KAN", line):
+			word_count += 1
+			word = str(line.split()[2])
+			# Eliminate diphthongs, so their syllables don't get double counted
+			for diph in ["aI", "aU", "OY"]:
+				if diph in word:
+					kan_syl_count += word.count(diph)
+					word = word.replace(diph, "")
+			# Count rest of vowels
+			for v in vowel_list[:-12]:
+				if v in word:
+					kan_syl_count += word.count(v)
+		# Stop search when KAN tier passed, so cursor doesn't reach EOF
+		elif re.match("SYN", line):
+			break
+
+	# Iterate over the lines starting with "MAU", which do not correspond to a pause.
+	for line in datei:
+		if re.match("MAU", line) and ("<p" not in line):
+			phoneme_count += 1 							# Get total no. of phonemes
+			sp_dur += int(line.split()[2])			# Get total duration without pauses
+			if str(line.split()[4]) in vowel_list:	# As syllable nuclei are always vowels, we count them to get the syllable count
+				mau_syl_count += 1
+
+	datei.seek(91)
+
+	duration = sp_dur * 0.0000625				# Transform total duration in seconds
+
+	return duration, phoneme_count, word_count, mau_syl_count, kan_syl_count
+
+
[…]

### 2017-04-06 py_files_old/stress.py

@@ -1,0 +1,248 @@
+import re
+
+short_vowels = ["a", "e", "E", "I", "i", "O", "o", "U", "u", "Y", "y", "9", "2"]
+
+# Sets the context of search at file level and finds the right word number
+# Returns the type of stress that falls on the given vowel of the given file of the given word 
+# none (unstressed), s_stress (secondary), p_stress (primary)
+def stress_type(datei, word_number, phoneme, cursor):
+	work_file = open(datei)
+	word_no = int(word_number)
+	z = 0
+#	Instructions for Q to get the stress of the following phoeneme
+#	for line in work_file:
+#		z += 1
+#		if z == (cursor):
+#			phoneme = line.split()[4]
+			#print(phoneme)
+		
+	mau_rank = p_rank_in_mau(work_file, word_no, phoneme, cursor)
+	#print("MAU-rank: " + str(mau_rank))
+	#stress_type = ""
+	#word = ""
+	
+	for line in work_file:	
+		if re.match("KAN", line) and (int(line.split()[1]) == word_no):
+			word = str(line.split()[2])
+			stress_type = set_stress(datei, word, phoneme, word_number, mau_rank, cursor)
+			break
+		else:
+			continue
+
+	work_file.close()
+	return stress_type
+
+# Sets the context of search at word level
+# Actually determines the stress of the given phoneme in the given word.
+def set_stress(datei, word, phoneme, word_number, mau_rank, cursor):
+	stress_type = ""
+	p_index = -1
+	mau, kan = kan_mau_string_builder(datei, word, word_number)
+
+	# Schwa is never stressed
+	if phoneme in ["@", "6"]:
+		stress_type = "none"
+	# Q occurs always before a vowel, either at word initial position or inside word-composita bevor a stressed vowel morpheme initial
+	# As cases with more than two Q in a word are extremely rare in our database, we decided to make only 2 distinctions for its position
+	elif phoneme == "Q":
+		qPosition = find_qPosition(datei, cursor)
+		stress_type = set_QLabel(qPosition, kan)
+	# Set stress for short vowel monophthongs (except "a~", which is included in the diphthong search)
+	else:#if phoneme in short_vowels:
+		stress_type = short_mono_stress(mau, kan, mau_rank, phoneme)
+
+	return stress_type
+
[…]

### 2017-04-06 py_files_old/t_modify_files.py

@@ -1,0 +1,99 @@
+import os
+import re
+import model_utilities
+import stress
+import get_syllables
+import time
+
+dataset = model_utilities.get_path_list('C:/Users/alexutza_a/Abschlussarbeit/DB_Verbmobil/Evaluation/Rest')
+vowel_list = ["a", "a~", "e", "E", "I", "i", "O", "o", "U", "u", "Y", "y", "9", "2", "a:", "a~:", "e:", "E:", "i:", "o:", "u:", "y:", "2:", "OY", "aU", "aI", "@", "6"]
+
+t1 = time.time()
+# Returns a list containing the syllable number for each phoneme in given word number
+# Looks like: [0011] for the word genau realized as /g@naU/
+def get_syl_no(datei, word_no):
+	syl_dict = get_syllables.syl_position(datei, word_no)
+	syl_numbers_list = []
+	key_list = sorted(syl_dict.keys())
+	#print(syl_dict)
+	i = 0
+	for syl_no in key_list:
+		for value in syl_dict[syl_no]:
+			# Consider cases where elements of diphthongs, nasals, or long vowels are listed separately in the value list
+			if value not in ["~", ":", "I", "U", "Y"]:
+				syl_numbers_list.append(i)
+			elif value in ["I", "U"] and "a" not in syl_dict[syl_no]:
+				syl_numbers_list.append(i)
+			elif value == "Y" and "O" not in syl_dict[syl_no]:
+				syl_numbers_list.append(i)
+
+		i += 1
+	#print(syl_numbers_list)
+	return syl_numbers_list
+
+i = 0
+# Here starts the file annotation
+for datei in dataset:
+	if i%180 == 0:
+		print(i)
+
+	z = 0
+	work_file = open(datei)
+	m_file = open('C:/Users/alexutza_a/Abschlussarbeit/DB_Verbmobil/mod_dataset/' + str(datei)[-20:-4] + "_a" + str(datei)[-4:], "w")
+	syl_list_counter = 0
+	for line in work_file:
+		z += 1
+		if re.match("MAU", line):
+			syl_numbers_list = get_syl_no(datei, int(line.split()[3]))
+
+			# Annotate only when no break, and syllable dictionary not empty
+			if int(line.split()[3]) >= 0 and len(syl_numbers_list) > 0:
+				
+				# Case1: current phoneme is a vowel
+				if (str(line.split()[4]) in vowel_list):
+					phon_stress = stress.stress_type(datei, int(line.split()[3]), str(line.split()[4]), z)
[…]

### 2017-04-06 py_files_old/word_stats.py

@@ -1,0 +1,88 @@
+import re
+import phon_dict
+import speech_rate
+from collections import defaultdict
+
+path_list = phon_dict.verbmo_par_files()
+
+word_list = ["genau", "eineinhalb", "Arbeitstreffen", "Flughafen", "auf", "einen", "fliegen", "Restaurant", "Regul"]
+
+# Returns a dictionary of the reference words containing:
+#    filename, word_duration, list of sounds and their durations in this word
+# 	 All durations are in samples. Searches the complete verbmobil database.
+def create_word_dict():
+	word_dict = defaultdict(list)
+	for datei in path_list:
+		filename = str(datei)
+		work_file = open(datei)
+		for line in work_file:
+			if re.match("ORT", line) and (line.split()[2] in word_list):
+				# Filename
+				word_dict[line.split()[2]].append(datei[-20:])
+				# Word duration
+				word_duration, phoneme_count, mau_syl_count = speech_rate.word_duration(datei, int(line.split()[1]))
+				word_dict[line.split()[2]].append(word_duration)
+				#word_dict[line.split()[2]].append(int(round(word_duration/0.0000625, 0)))
+				# List of word sounds and their respective durations
+				word_dict[line.split()[2]].append(word_sounds(datei, line.split()[1]))
+		work_file.close()
+	return word_dict
+
+
+# Returns a list of the sounds in the given word and their resp. durations (in samples)
+#    Looks like: ["a", 648, "b", 756, ...]
+def word_sounds(datei, word_no):
+	work_file = open(datei)
+	sound_list = []
+	for line in work_file:
+		if re.match("MAU", line) and int(line.split()[3]) == int(word_no):
+			sound_list.append(line.split()[4])
+			sound_list.append(int(line.split()[2]))
+
+	work_file.close()
+	return sound_list
+
+word_dict = create_word_dict()
+# Create a list of durations of word "genau"
+genau_dauer = word_dict["genau"][1::3]
+#genau_dauer = [ float(x)*0.0625 for x in genau_dauer]
+print(genau_dauer)
+
+sound_lists = word_dict["genau"][2::3]
+#print(sound_lists)
+#print(sound_lists[0])
+#print(len(sound_lists))
+
+def create_word_phon_dict():
+	wordSound_dict = defaultdict(list)
+	k = 0
+	for el in sound_lists:
[…]

### 2017-05-02 Paper/Paper_v2.tex
Fixed formatting errors and corrected some chapters in the paper. See Paper_v2 for changes.

--- 2017-04-06 Paper/Paper_v1.tex
@@
 \usepackage{geometry}	
-\usepackage[backend=bibtex,style=numeric]{biblatex}
-\usepackage{ngerman}
-\usepackage[T1]{fontenc}
+%\usepackage[backend=bibtex,style=numeric]{biblatex}
+%\usepackage{ngerman}
+%\usepackage[T1]{fontenc}
 \usepackage[utf8]{inputenc}
 \usepackage[style=numeric]{biblatex}
-\bibliography{../PaperBib}
+\bibliography{../PaperBib2}
 \usepackage{colortbl}	
 \usepackage{xcolor}
-\usepackage{soul}
+%\usepackage{soul}
 \usepackage{amsmath}
@@
 \usepackage{textcomp}
-\usepackage{floatrow}
+%\usepackage{floatrow}
 \renewcommand{\figurename}{Fig.}
-\usepackage{caption}
+\usepackage[skip=0pt]{caption}
+\captionsetup{format=hang}
 \setcapindent{0pt}
@@
 \usepackage{scrhack}
-\renewcommand{\familydefault}{\sfdefault}
+%\renewcommand{\familydefault}{\sfdefault}
 \definecolor{uhhred}{cmyk}{0,100,100,0}
@@
 \newcolumntype{R}[1]{>{\raggedleft\arraybackslash}m{#1}}
+
+\usepackage{hyperref}
+\hypersetup{linktocpage}
 
@@
 
-\tableofcontents
+\tableofcontents\thispagestyle{empty}
+
+\mainmatter
 
 \chapter{Introduction}
-Some people speak faster, others speak slower, and all people speak sometimes faster or slower as they usually do.
-This ``speed`` at which one speaks is what we call ``speech rate``.
+Some people speak faster, others speak slower, and all people speak sometimes faster and sometimes slower than they usually do.
+The tempo at which one speaks is what we call ``speech rate``.
 The fact that we understand people at all natural speech rates is certainly largely due to the quality features of sounds (phonemes) in the given language.
 However, if we have a perfect audio recording of speech, and play it faster or slower, we immediately notice how naturalness deteriorates.
-Synthetic speech shows a similar behavior when changing the speaking rate.
+Synthetic speech shows a similar behavior when changing tempo.
 In both cases phoneme quality doesn't need to change.
-It is its duration that changes, along with some other suprasegmental features like pitch, phrasal accents, etc., and that do the trick.
+It is its duration that changes, along with some other suprasegmental features like pitch, and phrasal accents that do the trick.
 
@@
 Phoneme length adjustment is the aspect we wish to adress in this work.
-In particular, we want to find an algorithm for adjusting phoneme durations according to the variation of the speech rate.
-The resulting phoneme duration should be close to the actual duration that occurs when the speaking rate alters.
+In particular, we want to find an algorithm for adjusting phoneme durations to the variation of the speech rate.
+The resulting phoneme duration should be close to the actual duration that occurs when speaking rate alters.
 This would be the same as playing an audio recording faster or slower.
@@
 There are several models for phoneme duration approximation that work pretty well for some languages, German inclusive.
-However, they consistently don't take the speech rate into consideration, their performance level being acquired rather within a specific speech rate.
+However, they consistently don't take the speech rate into consideration, their performance level being acquired rather at a specific speech rate.
 When this varies, model performance deteriorates, like van Santen states for his model \cite{Santen1994}.
-Moreover, all these models have an upper asymptote for their accuracy situated at less than 100\%.
-It has been suggested, that the influence of so-called ``macroscopic``or para-linguistic \cite{Santen1994} factors should be examined in order to move this asymptote further up.
+Moreover, all of these models have an upper asymptote for their accuracy situated at less than 100\%.
+It has been suggested, that the influence of so-called ``macroscopic`` or para-linguistic \cite{Santen1994} factors should be examined in order to move this asymptote further up.
 Such a factor is speech rate, and we took the challenge of integrating it into a speech model.
 
-The purpose is therefore not to create another duration approximation model for TTS-systems, but to improve the algorithm of phoneme duration prediction by considering the tempo of speech.
-Table \ref{tab:Question} visualizes our challenge: we want to put all available/visible information from the table in relation so that we can replace the question marks with values that are close to the real ones.
-Consequently, the modified speech unit would sound more natural at a faster or a slower speech rate as when being modified with a linear function.
+The purpose is therefore not to create another duration approximation model for TTS-systems, but to improve the algorithm of phoneme duration prediction by considering speech tempo.
+We argue therefore, that speech tempo produces a non-linear alteration of the phoneme durations.
+
+Table \ref{tab:Question} visualizes our challenge: we want to put all available, i.e.
+visible information from the table in relation so that we can replace the question marks with values that are close to the real ones.
+Consequently, the modified speech unit sounds more natural when tempo varies.
 
 \begin{table}[htbp]
-\caption {The challenge} 
 \centering
@@
 \end{tabular}
-\caption{The first column shows our inputs: Duration of each phoneme in German word ``genau``, given a speech rate of 0.07 sec, or a word duration of 0.28 sec.
-The second column visualizes our challenge: How long is the duration of each phoneme of the given word, at a speech rate of 0.11 sec, equivalent of a word length of 0.42 sec?}
+\vspace{5mm}
+\caption{Our inputs: duration of each phoneme in German word ``genau`` at a speech rate of 0.07 sec.
+corresponding to a word duration of 0.28 sec, and a new speech rate, 0.11 sec, equivalent to a word duration of 0.42 sec, for which we have the challenge of finding appropriate phoneme durations.}
 \label{tab:Question}
@@
 \ref{fig:linear_approx}.
+
+Across the paper we use the German SAMPA notations when referring to phonemes.
 \begin{figure}[htbp]
 	\centering
-	\includegraphics[scale=0.7]{../Graphen/Linear_approx.eps}
-	\caption[Linear phoneme duration approximation]{This is a linear solution to the challenge presented in table \ref{tab:Question}.}
+	\includegraphics[trim=23mm 23mm 0mm 0mm, scale=0.7]{../Graphen/Linear_approx.eps}
+	\vspace{4mm}
+	\caption[Linear phoneme duration approximation]{This is a linear solution to the challenge presented in Table \ref{tab:Question}.}
 	\label{fig:linear_approx}	
@@
 
-Across the paper we use the German SAMPA notations when referring to phonemes.
-
 \subsection*{Outline}
-We start our challenge by defining our methods in chapter 2.
-In the following chapter we examine the available resources so that we can proceed with the actual task in chapter 4, where we analyze the German phonemes.
-We dedicate chapter 5 to the challenge of finding an adequate speech rate definition for our database, and which can serve our purpose.
-Chapter 6 finally deals with phoneme duration prediction models and the challenge of finding a solution to our above presented issue.
+We start our challenge by defining our methods in Chapter \ref{chap_2}.
+In the following chapter we examine the available resources so that we can proceed with the actual task in Chapter \ref{chap_phonemes}, where we analyze the German phonemes.
+We dedicate Chapter \ref{chap_SR} to the challenge of finding an adequate speech rate definition for our database, and which can serve our purpose.
+Chapter \ref{chap_6:models} finally deals with phoneme duration prediction models and the challenge of finding a solution to our above presented issue.
 
 \chapter{Means and Methods}
+\label{chap_2}
 
@@
 We used machine learning to solve a task involving both classification and numeric prediction, and selected from its repertoire two model tree methods: M5P and REPTree.
-The reasons we chose are: trees are easy to interpret, they can perform both classification and numeric prediction tasks, we could easily generate them using Weka, and they are the closest to classification and regression (CART) trees, also used in the literature \cite{Brinckmann_2003} for phoneme duration prediction, so we have data to compare with.
+We chose these trees because: trees are easy to interpret, they can perform both classification and numeric prediction tasks, we could easily generate them using Weka, and they are the closest to classification and regression (CART) trees, also used in the literature \cite{Brinckmann_2003} for phoneme duration prediction, so we have data to compare with.
 
 \subsection{Decision trees}
+\label{dec_trees}
 Decision trees divide the attribute space into well defined clusters, identified by a class name or a numeric value.
@@
 \end{enumerate}
+Throughout this work we consistently mean the Pearson's correlation coefficient whenever we speak about correlation.
 
@@
 
-In order to extract the needed information from our database, and put it in a form ready to use with Weka, we used self-made code written in Python 3.6.
+In order to extract the needed information from our database, and put it in a form ready to use with Weka, we used self-made code written in \textbf{Python} 3.6.
 This is an easy to learn programming language with many features dedicated to statistical analysis and even machine learning.
@@
 \chapter{Corpora}
+\label{chap_3}
 Language related databases are called corpora, and for our purpose we needed one containing segmented and annotated German speech as a set of text files, so-called ``textual data``.
@@
 
-There are three corpora which we had at our disposal: (1) The Kiel Corpus of Read (PhonDat) and of Spontaneous Speech, (2) Verbmobil, and (3) spoken Wikipedia.
+There are three corpora which we had at our disposal: (1) The Kiel Corpus of Read (\textit{PhonDat} or \textit{KCoRS}) and of Spontaneous Speech (\textit{KCoSS}), (2) Verbmobil, and (3) spoken Wikipedia.
 The first one contains two equal sized corpora of > 4 hours each: KCoRS, which is a collection of read speech, and KCoSS, which is a collection of spontaneous speech, both being manually segmented, a good indicator for good segmentation quality.
@@
 Spoken Wikipedia is a new project comprising of a much larger collection of read speech.
-Unfortunately, the phoneme-level segmentation was not ready on time to evaluate in our project.
+Unfortunately, the phoneme-level segmentation was not ready on time to evaluate in our project, so we chose Verbmobil.
 
@@
 \chapter{German phoneme system}
+\label{chap_phonemes}
 The phoneme inventory used in Verbmobil includes 52 classes defined according to the German SAMPA.
 However, only 49 of these occur in the examined corpus.
-The missing ones are \texttt{/2/} as in \textit{Ökonom}, and the nasals \texttt{/E\~/} and \texttt{/O\~/} as in the loanwords \textit{Teint, Saison}.
+The missing ones are \texttt{/2/} as in \textit{Ökonom}, and the nasals \texttt{/E$\sim$/} and \texttt{/O$\sim$/} as in the loanwords \textit{Teint, Saison}.
 Using the classical phonetic conventions, we can group these phonemes using a tree structure.
@@
 Pauses may be caused by and/or be quantitatively modified both by linguistic (e.g.
-phonological) and non-linguistic factors, like technical, psychological, dialog dynamics, etc.
-Consequently pause length variation needs to be treated separately from phoneme duration variation, as a not negligible large set of other factors are to be considered for this, such as dialogue dynamics, hesitations, technical issues of the recording, which extends beyond the purpose of this work.
+phonological) and non-linguistic factors, like technical, psychological, and dialog dynamics.
+Consequently pause length variation needs to be treated separately from phoneme duration variation, as a not negligible large set of other factors are to be considered for this, such as dialogue dynamics, hesitations, or technical issues of the recording which all extend beyond the purpose of this work.
 
@@
 We investigated this assumption by comparing the vowel duration inside words containing a specific number of phonemes, as well as the variation of the word time proportion occupied by them in such words.
-The results confirmed this hypothesis, as you may see in Fig.
-\ref{fig:w_dur_vs_pho} and \ref{fig:boxplot_prop}.
+The results confirmed this hypothesis.
+We illustrate this in Figure \ref{fig:w_dur_vs_pho} on the word ``genau``, where you can see that the length of the diphthong \texttt{/aU/} changes proportionally to the word duration, while the other phonemes seem to be unaffected by word duration.
+This leads to the idea that at least some phonemes tend to occupy a specific steak (proportion) of the total word duration in words with a specific composition accounting for a linear correlation with word duration in such cases.
+Figure \ref{fig:boxplot_prop} shows the proportion distribution of diphthongs, the longest long vowel \texttt{/a:/}, its short counterpart \texttt{/a/}, and a plosive consonant \texttt{/b/} in words containing a total of three phonemes.
+As expected, the diphthongs occupy the greater steak, while the consonant the smaller one.
 This also means that it is a good idea to look for the correlation between vowels and speech rate.
@@
 \begin{figure}[htbp]
-	\noindent\makebox[\textwidth]{\includegraphics[width=\textwidth, clip]{../Graphen/Stats_for_genau.eps}}
+	\noindent\makebox[\textwidth]{\includegraphics[trim=30mm 30mm 30mm 10mm, clip, width=\textwidth]{../Graphen/Stats_for_genau.eps}}
 	\centering
-	\vspace{-10mm}
+	%\vspace{mm}
 	\caption[Word duration of ``genau`` vs.
@@
 	\label{fig:w_dur_vs_pho}
-\end{figure}
-
-\begin{figure}[htbp]
-	\includegraphics[width=.55\linewidth]{../Graphen/box3_oyauaiaab.eps}
-	\centering
-	\caption[Proportions of diphthong in 3-phoneme-words]{The time slice occupied by vowels in words of specific length  vary relatively little, indicating a possible linear correlation between phoneme duration and the variation of word duration.
-The y-axis measures time in seconds.}
-	\label{fig:boxplot_prop}
 \end{figure}
@@
 
-Table \ref{tab:vowel_stats} below presents the main duration information regarding the above mentioned classes that consolidates the discussion presented above.
+\begin{figure}[htbp]
+	\includegraphics[trim=15mm 0mm 15mm 0mm, width=.55\linewidth]{../Graphen/box3_oyauaiaab.eps}
+	\centering
+	\caption[Proportions of diphthong in 3-phoneme-words]{The time slice occupied by vowels in words of specific length  vary relatively little, indicating a possible linear correlation between their duration and the variation of word duration.
+Here you can see the difference in steak occupied by diphthongs next to the longest monophthongs \texttt{/a:/}, \texttt{/a/}, and a consonant \texttt{/b/}.
+The y-axis measures time in seconds.}
+	\label{fig:boxplot_prop}
+\end{figure}
+
+Table \ref{tab:vowel_stats} below presents the main duration information regarding the above mentioned vowel classes that consolidates the discussion presented above.
 \begin{table}[htbp]
@@
 \end{tabular}
+\vspace{5mm}
 \caption{Numeric information on the duration of the main vowel classes. The unit of the values presented in all columns except ``Count`` is second.} 
@@
 \begin{figure}[htbp]
-	\includegraphics[width=.55\linewidth]{../Graphen/vow_art.eps}
+	\includegraphics[width=.5\linewidth]{../Graphen/vow_art.eps}
 	\centering
+\vspace{5mm}
 	\caption[Vowel square]{Vowels are produced by the air flow in an open vowel tract, and they differ based on the modification we add to the shape of this cavity.
@@
 These changes are usually represented in the so-called vowel square, which may be further simplified to a vowel triangle.
-Here you can see how to place a vowel square containing the main vowels in the mouth cavity.
+Here you can see how to place a vowel square containing the main vowels in the mouth cavity and how alone schwa lies in the middle of this square - for the production of schwa we don't need to act on the air flow.
 }
@@
 \section{Glottal stop}
+\label{Q}
 The glottal stop occurs 13 348 times and only word initially in our database, marked with \texttt{/Q/}.
@@
 
-However, unlike pauses we did consider it  for analysis, as it is has an important communicative function as delimiter for words and morphemes.
+However, unlike pauses we did consider it  for analysis, as it has an important communicative function as delimiter for words and morphemes.
 It is in fact one of the important clues which help hearer understand speech \cite{Ternes2012}.
@@
 \section{Consonants}
+\label{cons}
 In general, we can say that consonants present different degrees of elasticity, which are, however, generally lower than the ones of vowels.
 
-Based on the correlation between phoneme duration and speech rate we created a subgroup of consonants containing the nasals and \texttt{/x/, /C/, /h/} with a relatively good correlation value $\geq 0.7$, and considered plosives to be a subgroup because of their consistently low correlation $\leq 0.42$.
+Based on the correlation between phoneme duration and speech rate we created a subgroup of consonants containing the nasals and \texttt{/x/, /C/, /h/} with a relatively good correlation value $\geq 0.7$, and considered plosives to be an independent subgroup because of their consistently low correlation $\leq 0.42$.
 
@@
 We chose to include them in our phoneme overview, as some of these categories proved to be relevant for the modelling of phoneme durations as well.
-Consequently, level 2 split shows for for the consonant node following categories:
+Consequently, level 2 split of the consonant node shows following categories:
 \begin{itemize}
@@
 Nasals may also constitute the nucleus of a syllable, when the vowel gets reduced.
-/m/, /n/, /N/.
+German has three nasal consonants: /m/, /n/, /N/.
 	\item Liquids, also called approximants, are produced by partial closure of the mouth produced by the thongue.
@@
 One relevant feature is that they have the greatest freedom to occur in consonant clusters, that is a group of consonants belonging to the same syllable part.
-This also means that they must be particularly short, as we identified vowels as the greater steak holders in word duration.
-Another fact we considered, is the \texttt{/r/} is relatively seldom in German, being often reduced to a schwa, and when it is articulated, then it is generally extremely short, which durations comparable to the ones of the glottal stop.
+This also means that they must be particularly short, as we identified vowels to be the greater steak holders in word duration.
+Another fact we considered, is that \texttt{/r/} is relatively seldom in German, being often reduced to a schwa, and when it is articulated, then it is generally extremely short, which durations comparable to the ones of the glottal stop.
 \end{itemize}
@@
 \end{tabular}
+\vspace{4mm}
 \caption{} 
@@
 \chapter{Speech rate}
-Speech rate is a measurement of the ``tempo`` of speech, that is how quickly someone speaks.
-This may be seen as a discrete/qualitative measure, with discrete values as \textit{slow}, \textit{normal}, and \textit{fast} or as a continuous one, with continuous numerical values.
-We opted for the continuous version, because it is quantifyable and so can be a useful input into a regression, see Section 2.1.1.
+\label{chap_SR}
+Speech rate is a measurement of the tempo of speech, that is, how quickly someone speaks.
+This may be seen as a discrete (qualitative) measure, with discrete values such as \textit{slow}, \textit{normal}, and \textit{fast} or as a continuous one, with continuous numerical values.
+We opted for the continuous version, because it is quantifiable and so can be a useful input into a regression, see Section \ref{dec_trees} on decision trees.
 
@@
 In the case of speech, distance is not counted in the spatial domain but in relevant linguistic units.
-Other than in the case of distance, the length of the speech units is not standardized (like kilometers, miles, etc.).
-Moreover it's size is also expressed in units of time: phoneme \texttt{/a/} has the length of x seconds, miliseconds, etc.
+Other than in the case of distance, the length of the speech units is not standardized, like kilometers, or miles, but have different lengths.
+Moreover its size is also expressed in units of time: phoneme \texttt{/a/} has the length of \textit{x} seconds, milliseconds, etc.
+This is why we chose to revert the speed calculating fraction, as you may see in Eq.
+\ref{eq:eq1}, and divide the total time by the number of units contained to obtain units of equal length.
+The measure of speech tempo is then given by the resulting size of the units.
+It is logical, that when someone speaks faster, the number of phonemes uttered in a given time interval is larger, and therefore their average length shorter.
+In the case of slow speech the opposite is true, so this is a good measure for speech tempo.
 
-This is fine, if one has a fixed time interval and is interested in the number of units that may fit in.
-However, this way of looking at it does not consider the fact, that the length of the unit may vary.
-We are not interested in variing the number of units, which should actually stay constant.
-We are rather interested in variing the time interval while keeping the number of units constant.
-This is what happens when someone is speaking faster or slower than a specific norm.
-Therefore we inverted the typical speed-like calculating formula, as you can see in Eq.
-\ref{eq1}, so we can investigate the length variation in the speech units, with focus on units = phonemes.
+Next, we need to choose an adequate time interval to calculate speech rate for.
+As we are interested in comparing the phoneme duration at different speech rates, one global speech rate calculated on the entire corpus would be inappropriate.
+In addition, the fact that our corpus contains only spontaneous speech suggests there is speech tempo variability also inside speech turns, and even inside prosodic phrases, as you may see in Figure \ref{fig:SR_example}, so we decided to get very local and calculate speech rate at word level.
+
+\begin{figure}[htbp]
+	\centering
+	\noindent\makebox[\textwidth]{\includegraphics[trim= 53mm 0mm 70mm 70mm, clip, width=.83\paperwidth]{../Graphen/SR_example.eps}}
+	\vspace{-25mm}
+	\caption[Example of speech rate variation inside a turn]{Transcription of speech turn no.
+22 of speaker AAJ in dialogue g001.
+The closer the points are to one another, the faster he speaks.
+Orange lines mark prosodic boundaries, delimiting prosodic phrases.
+You can see great variation in speech tempo across this speech turn, as well as inside single phrases.}
+	\label{fig:SR_example}
+\end{figure}
+
+The units to be counted inside given time interval, which is in our case the duration of a word, may be phonemes, or syllables.
+In the literature \cite{Pfitzinger1998} we also found mentions about a relative speech rate, which combines speech rate calculated for two different units: phonemes, and syllables.
+As our corpus shows a very large amount of 1- and 2-syllable words, taking phonemes as unit seems to be the most reasonable option.
+Therefore, we used speech rate defined as:
 
 \begin{equation}
-\label{eq1}
+\label{eq:eq1}
 	SR = \frac{word \quad duration}{phoneme\quad count}
@@
 
-The next thing to think about is the size and identity of the fraction elements.
-Unit size may be phoneme, syllable or word.
-Variing the scope from local to globa, the segment on which we calculate speech rate may be a syllable, a word, a phrase, a speech turn or the entire corpus.
-In the literature \cite{Pfitzinger1998} we also found mentions about a relative speech rate, which combines the speech rate calculated to two different unit sizes.
-
-From the beginning, we did not consider the variant of calculating a speech rate for the entire corpus.
-Given its structure of dialogues recorded by many speakers, it is straightfoward that speech rate may vary tremendousely not only between speakers, but also between different speech turns, and consequently be a bad measure for characterizing our corpus.
-The size of a speech turn in a dialog proved to be a bad measure as well.
-The reason for this is again the nature of spontaneous speech as ``live`` creation of the human mind.
-Hesitations, thinking, and emotions have a much greater influence as in read speech, and result in a variing tempo at which we produce speech.
-In figure \ref{fig:SR_example} we are illustrating an example of this phenomenon.
-
-\begin{figure}[htbp]
-	\centering
-	\noindent\makebox[\textwidth]{\includegraphics[width=\paperwidth, clip]{../Graphen/SR_example.eps}}
-	\vspace{-25mm}
-	\caption[Example of speech rate variation inside a turn]{This is the content of the speech turn number 22 of speaker AAJ in dialoge g001.
-The closer the points are to oneanother, the faster he speaks.
-Orange lines mark prosodic boundaries.}
-	\label{fig:SR_example}
-\end{figure}
-
-Seing this, there is no wonder that we got no good correlation values as long as we calculated speech rate on entire turns.
-This got better when we reduced the size of the segment to phrase, and then further down to words.
 An additional measure we used to evaluate how well a speech rate calculating formula may suit our purpose was the correlation value between speech rate and phoneme duration.
-As long vowels proved to be the most sensible to the variation of the speaking tempo, we present below the values we obtained for this group, together with the speech rate calculated at different segment sizes.
+You may see a listing of the Pearson correlation values in Table \ref{tab:SR_corr} as well as a scatter plot of the best suiting result in Figure \ref{fig:wsr}.
+As long vowels proved to be the most sensible to the variation of the speaking tempo, we present below the values obtained for this group, together with the speech rate calculated at different segment sizes.
 
 \begin{table}[htbp]
-\caption {Duration of long vowels vs. speech rate calculated at turn, phrase, and word level} 
-%\centering
+\centering
 \vspace{10mm}
@@
 \hline
+Turn	& $turn\_dur/phon\_count$ & 0.73 \\
 
-Phrase  & $phrase\_dur / phon\_count$ & 0.561266362391\\
 \hline
-Word  & $word\_dur / phon\_count$ &  0.794901825792\\
+Phrase  & $phrase\_dur / phon\_count$ & 0.56\\
+	\hline
+Word  & $word\_dur / phon\_count$ &  0.80\\
 	\hline
 \end{tabular}
+\vspace{4mm}
+\caption {Duration of long vowels vs. speech rate calculated at turn, phrase, and word level} 
+\label{tab:SR_corr}
 \end{table}
@@
 	\centering
-	\includegraphics[width=\textwidth]{../Graphen/SR4.eps}
-	\label{fig:}
-\caption[Speech rate vs long vowel duration]{Phoneme durations measured in msec plotted against speech rates.}
+	\includegraphics[width=.8\textwidth]{../Graphen/SR4.eps}
+	\vspace{4mm}
+	\caption[Speech rate vs long vowel duration]{Long vowel durations plotted against their respective speech rate (msec).}
 	\label{fig:wsr}
@@
 \chapter{Phoneme duration prediction models}
+\label{chap_6:models}
 There are several duration prediction methods in use: 
@@
 \section{Method}
-Our starting premise was that phoneme durations are being influenced by the speaking tempo.
-In addition to this, we assumed that this influence is not equal on all phonemes, and on all situations, and therefore also not linear.
+Our starting premise was that phoneme durations are being influenced by the speaking rate.
+In addition to this, we assumed that this influence is not equal on all phonemes, and on all situations, and therefore not linear.
 So we formulated our aim as to investigate the influence of speech rate on the duration of different phonemes and find a way of using this information to improve the performance of phoneme prediction models.
@@
 There are consonants like \texttt{/N/}, and \texttt{/m/} which may be longer than some long vowels, and long vowels which are shorter than some short vowels or even several consonants, like \texttt{/y:/}.
-If one would apply a clustering method on phonemes to group them on seven classes based on their duration interval coordinates using, say k-means method, the resulting groups would certainly differ from those listed in Chapter 4.
+If one would apply a clustering method on phonemes to group them on seven classes based on their duration interval coordinates using, say k-means method, the resulting groups would certainly differ from those listed in Chapter \ref{chap_phonemes}.
 
@@
 The relation between phoneme duration and factors other than speech rate was treated only marginally, as these have already been extensively examined in the literature on various corpora.
-For the German language, intuition aside, we have information on such factors provided by: Kohler \cite{Kohler1992}, Riedi \cite{Riedi1995}, Moebius \cite{Moebius1996}, and Brinckmann \& Trouvain \cite{Brinckmann_2003}, who applied some of the existent segment duration models to the German language.At this stage, we considered such factors only to the extent they might have helped identify groups of phonemes that may be best suited to evaluate the influence of speech rate.
+For the German language, intuition aside, we have information on such factors provided by: Kohler \cite{Kohler1992}, Riedi \cite{Riedi1995}, Moebius \cite{Moebius1996}, and Brinckmann \& Trouvain \cite{Brinckmann_2003}, who applied some of the existent segment duration models to the German language.
+At this stage, we considered such factors only to the extent they might have helped identify groups of phonemes that may be best suited to evaluate the influence of speech rate.
 Such a factor is primary stress, which creates a group of very elastic vowels, which are therefore also sensitive to changes in speech tempo.
@@
 This supposed the presence of a good correlation between speech rate and phoneme duration, at least for some phoneme groups.
-We explained our choice whith respect to this in Chapter 5.
+We explained our choice whith respect to this in Chapter \ref{chap_SR}.
 The fact, that speech rate doesn't show any correlation with some phonemes, like \texttt{/r/}, while it has a pretty good correlation with others, like \texttt{/a:/} suggests the need for classification in whichever model we choose.
@@
 	\item The first one predicted the phoneme duration as the mean of its durations across the whole corpus. 
-	\item The second one build a two level tree, which split the root on the 49 realized phonemes, and then each phoneme on two specific speech rate values.
+	\item The second one built a two level tree, which split the root on the 49 realized phonemes, and then each phoneme on two specific speech rate values.
 The leaves contained three predicted values for each phoneme.
@@
 Their principle is easy to understand and they are available in Weka.
+After pruning, the \textit{word} related features got completely eliminated, together with \textit{syllable-place-in-word} and \textit{articulation-of-realised-previous phoneme}.
 Consequently, following features were selected:
@@
 	\begin{itemize}
-		\item phoneme id: all the 52 phoneme definitions used in Verbmobil
+		\item phoneme id: all 52 phoneme definitions used in Verbmobil
 		\item phoneme type: vowel, or consonant
@@
 		\item position of phoneme in syllable: initial, final
-		\item syllable part containing current phoneme: onset, nucleus, code, single-phoneme-syllable
+		\item syllable part containing current phoneme: onset, nucleus, coda, single-phoneme-syllable
 		\item lexical stress of the syllable (using the stress type of the nucleus vowel)
@@
 		\item if following phoneme is voiced: yes, no, or none
-		\item syllable part to which following phoneme belongs: onset, nucleus, code, single-phoneme-syllable
+		\item syllable part to which following phoneme belongs: onset, nucleus, coda, single-phoneme-syllable
 	\end{itemize}
@@
 
-Note that all features consider only actually realised units in speech.
+Note that all features refer to actually realised units in speech.
 We also had to reduce the amount of data used for the model to 58\% of the total corpus, due to capacity limitations imposed by Weka on our system.
-After pruning, the \textit{word} related features got completely eliminated, together with \textit{syllable-place-in-word} and \textit{articulation-of-realised-previous phoneme}.
 
-To this collection of features we added \textit{speech rate} in the second phase and compare the results.
+To this collection of features we added \textit{speech rate} in the second phase and compared the results.
 
 \section{Results}
+\label{chap_6:results}
 We trained and evaluated both a M5P and a REPTree model.
-The results were very similar, so present here only the results obtained using the model tree M5P, as they were slightly better.
-As can be seen in Table \ref{tab:perfM} adding the speech rate feature caused a significant improvement on all our models.
+The results were very similar, so we list here only the results obtained using the model tree M5P, as they were slightly better.
+As can be seen in Table \ref{tab:perfM}, adding the speech rate feature caused a significant improvement on all our models.
 
@@
 \end{tabular}
+\vspace{4mm}
 \caption{Overview of the performance results of 3 models with (+) and without (-) speech rate (SR).
 CART is the model created by Brinckmann \& Trouvain \cite{Brinckmann_2003}.
-The two values in the CART cells represent the performance obtained for 2 different speakers.}
+The two values in the CART cells represent the performance obtained for 2 different speakers.
+RMSE was not reported for this model.}
 \label{tab:perfM}
@@
 
-However, the performance of our model tree is smaller than the performance of the CART built by Brinckmann \& Trouvain \cite{Brinckmann_2003}.
+However, the performance of our model tree is weaker than the performance of the CART model built by Brinckmann \& Trouvain \cite{Brinckmann_2003}.
 This is explainable if we consider the nature of the used datasets.
@@
 The reported results show even for PhonDat a difference in performance between speakers.
-Another aspect that can have a great impact on the model result is the quality of the input data.
+Another aspect that can have a great impact on model performance is the quality of the input data.
 While PhonDat was manually segmented and annotated, Verbmobil was processed automatically using the MAUS segmentation system.
-When evaluating outliers we found many examples of wrong segmentation, as we also discussed in Section 4.2.
+When evaluating outliers we found many examples of wrong segmentation, as we also discussed in Sections \ref{Q} and \ref{cons}.
 
 On the other hand, our model tree shows robustness with respect to factors like inter-speaker variability and speech rate.
-The results obtained for a test set containing data from a speaker not included in the training data were similar to those obtained on known speakers on terms of correlation coefficient, and showed less error rates, as can be seen in table \ref{tab:perfM5P}.
+The results obtained for a test set containing data from a speaker not included in the training data were similar to those obtained on known speakers on terms of correlation coefficient, and showed less error rates, as can be seen in Table \ref{tab:perfM5P}.
 Moreover, if the model is to be evaluated on data produced by only one speaker, it produces good results even when using less than 10\% of the amount of data used to train on the multiple speaker corpus.
@@
 \end{tabular}
+\vspace{4mm}
 \caption{Overview of the performance of our M5P model tree on different training and test sets.} 
@@
 \chapter{Conclusion and Outlook}
-Our model calculates an approximation for the phoneme duration when its corresponding speech segment changes in terms of duration.
-It does this no matter what was the cause of the duration change in the segment.
-However, when speech rate changes, there are also other speech characteristics, like prosody and phoneme quality which change, and which cannot be simulated only by modifying the phoneme duration.
-Therefore, making speech sound natural while modifying the speech rate needs to consider and adjust these other features too.
+Speech tempo is an important feature of natural speech, which produces lengthening or shortening of the speech units with reference to some ``standard`` speaking tempo.
+Our experiment shows that we can trace this influence at phoneme level, where it produces a non-linear modification in terms of duration of individual phonemes.
+Moreover, we have seen that adding a feature of tempo to a phoneme duration prediction model resulted in a significant improvement of model performance, even with a significant quantity of noise in the input data, as we stated in the discussion of our experiment results in Section \ref{chap_6:results}.
 
-Koreman showed \cite{Koreman_2006} that the perceived speech rate is also influenced by the listener's knowledge of the expected articulations for a particular utterance, therefore it makes a difference if all expected phones are also articulated or not.
-Consequently, one could compare speech rate considering only articulated vowels/syllables, as well as considering both realized and intended syllables, and test both methods with a given model.
-One possible result of such an approach would be the prediction of phonemes reaching $duration = 0$ at specific speech rates, which means predict phoneme drop phenomena.
+The main difficulty we encountered was to find an appropriate definition of speech rate to be used in our models.
+Our definition considers the structure of the corpus used and the purpose it needs to serve.
+In addition, it did produce a significant improvement of the model used, so it may be used in phoneme duration prediction models.
+However, a universal definition of speech rate, applying to different corpora, would be desirable.
+
+Future work would be to train and test such a model using the speech rate feature on data with a highly reduced amount of noise, which should result in an improved performance.
+
+A further possible improvement may achieved considering Koreman's observation \cite{Koreman_2006} on speech rate.
+He showed that perceived speech rate is influenced by the listener's knowledge of the expected articulations for a particular utterance, therefore it makes a difference if all expected phones are also articulated or not.
+Consequently, one could compare speech rate considering only articulated vowels/syllables, as well as considering both realized and intended phonemes, testing both methods with a given model.
+One possible result of such an approach would be the prediction of phonemes reaching $duration = 0$ at specific speech rates, which means predicting phoneme drop phenomena.
 In spontaneous speech, the word and syllable boundaries change, and this change influences other phonetic aspects such as stress, syllable duration, phoneme duration.
+
+More generally, we spoke about improving the quality of rendered speech under varying tempo.
+Correct phoneme length prediction is just a step when improving the speech quality retrieved by audio programs when the rendering speed is modified, which we adressed in this work.
+However, when tempo changes in natural speech, other speech characteristics are being modified as well, like prosody, pitch, and phoneme quality, which cannot be simulated only by modifying the phoneme duration.
+Therefore, making speech sound natural while modifying the rendering speed needs to consider and adjust these other features too, implying further work to investigate changes of these factors produced by speech rate modification.
 
@@
 \begin{center}\textbf{Eidesstattliche Erklärung}\end{center}
-Hiermit versichere ich an Eides statt, dass ich die vorliegende Arbeit im Bachelorstudiengang Wirtschaftsinformatik selbstständig verfasst und keine anderen als die angegebenen Hilfsmittel – insbesondere keine im Quellenverzeichnis nicht benannten Internet-Quellen – benutzt habe.
+Hiermit versichere ich an Eides statt, dass ich die vorliegende Arbeit im Bachelorstudiengang Informatik selbstständig verfasst und keine anderen als die angegebenen Hilfsmittel – insbesondere keine im Quellenverzeichnis nicht benannten Internet-Quellen – benutzt habe.
 Alle Stellen, die wörtlich oder sinngemäß aus Veröffentlichungen entnommen wurden, sind als solche kenntlich gemacht.

### 2017-05-02 Paper_v2/Paper_v2.tex
Fixed Bibliography in Paper_v2.

--- 2017-05-02 Paper/Paper_v2.tex
@@
 \usepackage[style=numeric]{biblatex}
-\bibliography{../PaperBib2}
+\bibliography{PaperBib}
 \usepackage{colortbl}	
@@
 \newgeometry{centering,left=2cm,right=2cm,top=2cm,bottom=2cm}
+
 \begin{titlepage}
@@
 \Large
+
 \begin{center} 
       {\color{uhhred}\textbf{\so{BACHELORTHESIS}}}
-%oder {\color{uhhred}\textbf{\so{MASTERTHESIS}}}
 	\vspace*{2.0cm}\\
@@
 \end{center}
+
 \vspace*{3.9cm}
@@
 Fachbereich Informatik \vspace*{0.4cm} \\ 
-%Ggf. Professur/Institut \vspace*{0.4cm} \\
 Studiengang: Informatik \vspace*{0.4cm} \\ 
@@
 \printbibliography
+
 \backmatter

### 2017-05-04 py_files/create_eval_lists.py
Added missing model (baseline + SR)

@@ -4,6 +4,7 @@ import model_wl
 import model_wl_ip
 import model_wl_ipdur
 import model_ipdur
+import model_ipdur_SR
 
 
 # List of paths to the files in our test data, to later iterate on
@@ -15,7 +16,7 @@ valid_phonemes = ["a", "a~", "e", "E", "I", "i", "O", "o", "U", "u", "Y", "y", "
 
 path_list_test = model_utilities.get_path_list('C:/Users/alexutza_a/Abschlussarbeit/DB_Verbmobil/Evaluation/Test_a')
 
-# Model type uses individual phonemes and their median, mean, or both in words having specific no. of phonemes
+# Model type uses individual phonemes and their median, mean, or both 
 # Actually creates the lists of predictions and actual values to evaluate model
 # @param model: 0, 1, or 2 meaning (median, mean, or (median + mean)/2)
 # Return: 2 lists of integers
@@ -35,7 +36,27 @@ def create_modelIPdur_lists(model):
 	#print(len(actuals_list))
 	return predictions_list, actuals_list
 
-# Model type uses individual phonemes and their median, mean, or both
+# Model type uses individual phonemes and their median, mean, or both and a split of the speech rate values
+# Actually creates the lists of predictions and actual values to evaluate model
+# @param model: 0, 1, or 2 meaning (median, mean, or (median + mean)/2)
+# Return: 2 lists of integers
+def create_modelIPdurSR_lists(model):
+	predictions_list, actuals_list = [], []
+	for datei in path_list_test:
+		work_file = open(datei)
+		for line in work_file:
+			# Restrain model to relevant phonemes
+			if re.match("MAU", line) and (str(line.split()[4]) in valid_phonemes):
+				predicted_pdur = model_ipdur_SR.pdur_prediction_value(datei, int(line.split()[3]), str(line.split()[4]), model)
+				predictions_list.append(predicted_pdur)
+
+				actuals_list.append(int(line.split()[2]))
+		work_file.close()
+	#print(len(predictions_list))
+	#print(len(actuals_list))
+	return predictions_list, actuals_list
+
+# Model type uses individual phonemes and their median, mean, or both in words having specific no. of phonemes
 # Actually creates the lists of predictions and actual values to evaluate model
 # @param model: 0, 1, or 2 meaning (median, mean, or (median + mean)/2)
 # Return: 2 lists of integers
@@ -53,6 +74,7 @@ def create_modelWLIP_lists(model, path_list_test, path_list_training):
[…]

### 2017-05-04 py_files/model_ipdur_SR.py

@@ -1,0 +1,50 @@
+import numpy as np
+from collections import defaultdict
+import model_utilities
+import model_ipdur
+import new_SR
+
+# This is a relatively static phoneme duration prediction model which only considers the median (or mean) value of a specific phoneme
+# and one split of speech rate values.
+#
+# Predicted value = median or mean of phon_dur in provided (training) dataset
+
+###################################################### START ##################################################################
+
+############## STATIC DATA STRUCTURES (not dataset specific) ##################################
+
+
+# Create list of filepaths to explore. This one uses entire data set, for exploration purposes.
+# path_list_training = verbmo_par_files('C:/Users/alexutza_a/Abschlussarbeit/DB_Verbmobil/verbmobil_par')
+path_list_training = model_utilities.get_path_list('C:/Users/alexutza_a/Abschlussarbeit/DB_Verbmobil/Evaluation/Training_a')
+
+valid_phonemes = ["a", "a~", "e", "E", "I", "i", "O", "o", "U", "u", "Y", "y", "9", "2", "a:", "a~:", "e:", "E:", "i:",
+                 "o:", "u:", "y:", "2:", "OY", "aU", "aI", "@", "6", "z", "S", "Z", "C", "x", "N", "Q", "b", "d", "f", 
+                 "g", "h", "j", "k", "l", "m", "n", "p", "r", "s", "t", "v"]
+
+vowel_list = ["Q", "a", "a~", "e", "E", "I", "i", "O", "o", "U", "u", "Y", "y", "9", "2", "@", "6", "a:", "a~:", "e:", "E:", "i:", "o:", "u:", "y:", "2:", "OY", "aU", "aI"]
+
+#################################### DYNAMIC DATA STRUCTURES (training data) ################################
+
+# Returns: (type: int) the predicted duration (samples) of given phoneme based on number of phonemes in word
+# @param model: 0, 1, or 2 meaning (median, mean, or (median + mean)/2)
+def pdur_prediction_value(datei, word_no, phoneme, model):
+	pdur_compressed_dict = model_ipdur.phon_dur_compressed_dict(model_utilities.phon_dur_dict(path_list_training))
+	word_dur, phoneme_count, mau_syl_count = model_utilities.word_statistics(datei, word_no)
+	SR = new_SR.get_word_SR(datei, word_no)[0]
+	SD = 73.91
[…]

### 2017-05-22 Paper/Paper_v1.tex
Added 2 notebooks on neat analysis of vowels and consonants.

@@
 \usepackage{geometry}	
-\usepackage[backend=bibtex,style=numeric]{biblatex}
-\usepackage{ngerman}
-\usepackage[T1]{fontenc}
+%\usepackage[backend=bibtex,style=numeric]{biblatex}
+%\usepackage{ngerman}
+%\usepackage[T1]{fontenc}
 \usepackage[utf8]{inputenc}
@@
 \usepackage{xcolor}
-\usepackage{soul}
+%\usepackage{soul}
 \usepackage{amsmath}

### 2017-05-22 py_files/model_utilities.py

@@ -48,6 +48,24 @@ def phon_dur_dict(path_list):
     #print(len(phon_dur_dict["a~:"]))
     return phon_dur_dict
 
+# Creates a flat dictionary of phonemes, their durations, and filename in the given dataset
+# Return looks like: {"a" : [1599, g001a_acn1_AAJ.par, 2348, g001a_acn1_AAJ.par, 959, ...], 
+#                     "b" : [...], ...}
+def dur_dict_filename(path_list):
+    dur_dict = dict( (i, []) for i in valid_phonemes)
+    for datei in path_list:
+        work_file = open(datei)
+        
+        for line in work_file:
+            if re.match("MAU", line) and (str(line.split()[4]) in valid_phonemes):
+                dur_dict[str(line.split()[4])].append(int(line.split()[2]))
+                dur_dict[str(line.split()[4])].append(str(datei)[-22:])
+        work_file.close()
+    #print(l)
+    #print(len(phon_dur_dict["Z"]))
+    #print(len(phon_dur_dict["a~:"]))
+    return dur_dict
+
 
 # Creates a nested dict of phoneme proportions in words of specific lengths (as no of phonemes) - w_dur in samples
 # Returns: a dictionary

### 2017-05-22 py_files/word_stats.py

@@ -1,17 +1,15 @@
 import re
-import phon_dict
-import speech_rate
+import model_utilities
 from collections import defaultdict
-
-path_list = phon_dict.verbmo_par_files()
 
 word_list = ["genau", "eineinhalb", "Arbeitstreffen", "Flughafen", "auf", "einen", "fliegen", "Restaurant", "Regul"]
 
 # Returns a dictionary of the reference words containing:
 #    filename, word_duration, list of sounds and their durations in this word
 # 	 All durations are in samples. Searches the complete verbmobil database.
-def create_word_dict():
+def create_word_dict(path_list):
 	word_dict = defaultdict(list)
+
 	for datei in path_list:
 		filename = str(datei)
 		work_file = open(datei)
@@ -20,7 +18,7 @@ def create_word_dict():
 				# Filename
 				word_dict[line.split()[2]].append(datei[-20:])
 				# Word duration
-				word_duration, phoneme_count, mau_syl_count = speech_rate.word_duration(datei, int(line.split()[1]))
+				word_duration, phoneme_count, mau_syl_count = model_utilities.word_statistics(datei, int(line.split()[1]))
 				word_dict[line.split()[2]].append(word_duration)
 				#word_dict[line.split()[2]].append(int(round(word_duration/0.0000625, 0)))
 				# List of word sounds and their respective durations
@@ -42,11 +40,11 @@ def word_sounds(datei, word_no):
 	work_file.close()
 	return sound_list
 
-word_dict = create_word_dict()
+word_dict = create_word_dict(model_utilities.get_path_list('C:/Users/alexutza_a/Abschlussarbeit/DB_Verbmobil/verbmobil_par'))
 # Create a list of durations of word "genau"
 genau_dauer = word_dict["genau"][1::3]
-#genau_dauer = [ float(x)*0.0625 for x in genau_dauer]
-print(genau_dauer)
+genau_dauer = [ float(x)*0.0625 for x in genau_dauer]
+#print(genau_dauer)
 
 sound_lists = word_dict["genau"][2::3]
 #print(sound_lists)
